### 2023-11-08 GHI_enh_02_process.tex
Commit 2023-11-08 14:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-11-08}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
Alpha * HAU is CS\_ref

\hypertarget{plot-some-days-with-strong-enhancement-cases}{%
\subsection{Plot some days with strong enhancement cases}\label{plot-some-days-with-strong-enhancement-cases}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-20} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-24} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-30} \end{center}

\textbf{END}

\begin{verbatim}
2023-11-08 09:12:54.0 athan@tyler GHI_enh_02_process.R 0.296492 mins
\end{verbatim}

\end{document}