### 2023-11-21 GHI_enh_02_process.tex
Commit 2023-11-21 12:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-11-21}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
Alpha * HAU is CS\_ref

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/unnamed-chunk-6-1} \end{center}

\hypertarget{plot-some-days-with-strong-enhancement-cases}{%
\subsection{Plot some days with strong enhancement cases}\label{plot-some-days-with-strong-enhancement-cases}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-20} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-24} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/strong_days-30} \end{center}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/enchtrendyear-1}

}

\caption{Trend of the total of enhancement cases per year.}\label{fig:enchtrendyear}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# \_ Cases per month  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}

\FunctionTok{plot}\NormalTok{(Enh\_monthly}\SpecialCharTok{$}\NormalTok{Date, Enh\_monthly}\SpecialCharTok{$}\NormalTok{N\_att ,}
     \AttributeTok{ylab =} \FunctionTok{bquote}\NormalTok{(}\StringTok{"Difference from mean [\%]"}\NormalTok{ )}
\NormalTok{)}
\CommentTok{\# title("Number of enchantments incidences", cex = 0.7)}
\NormalTok{lm1        }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{( Enh\_monthly}\SpecialCharTok{$}\NormalTok{N\_att }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Enh\_monthly}\SpecialCharTok{$}\NormalTok{Date )}
\FunctionTok{abline}\NormalTok{(lm1, }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\NormalTok{fit }\OtherTok{\textless{}{-}}\NormalTok{ lm1[[}\DecValTok{1}\NormalTok{]]}
\FunctionTok{legend}\NormalTok{(}\StringTok{\textquotesingle{}topleft\textquotesingle{}}\NormalTok{, }\AttributeTok{lty =} \DecValTok{1}\NormalTok{, }\AttributeTok{bty =} \StringTok{"n"}\NormalTok{,}
       \FunctionTok{paste}\NormalTok{(}\StringTok{\textquotesingle{}Y =\textquotesingle{}}\NormalTok{, }\FunctionTok{signif}\NormalTok{(fit[}\DecValTok{1}\NormalTok{],}\DecValTok{2}\NormalTok{),}\ControlFlowTok{if}\NormalTok{(fit[}\DecValTok{2}\NormalTok{]}\SpecialCharTok{\textgreater{}}\DecValTok{0}\NormalTok{)}\StringTok{\textquotesingle{}+\textquotesingle{}}\NormalTok{else}\StringTok{\textquotesingle{}{-}\textquotesingle{}}\NormalTok{,}\FunctionTok{signif}\NormalTok{(}\FunctionTok{abs}\NormalTok{(fit[}\DecValTok{2}\NormalTok{]}\SpecialCharTok{*}\NormalTok{Days\_of\_year),}\DecValTok{3}\NormalTok{),}\StringTok{\textquotesingle{}* year\textquotesingle{}}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/unnamed-chunk-9-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# \_ Cases per day  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/enchtrendday-1}

}

\caption{Trend of the total of enhancement cases per year.}\label{fig:enchtrendday}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/enchNtrendday-1}

}

\caption{Trend of the total of enhancement cases per year.}\label{fig:enchNtrendday}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# ignore extreme cases for now}
\CommentTok{\# plot(Enh\_sza$SZA, Enh\_sza$N\_ex)}
\end{Highlighting}
\end{Shaded}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{GHI_enh_02_process_files/figure-latex/szacases-1}

}

\caption{Enhancement cases percentage in total GHI data per SZA ($1^\circ$ bin).}\label{fig:szacases}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# consecutive data id}
\CommentTok{\# test \textless{}{-} c(F,T,F, rep(T,5),F,F,T,F,F,F, rep(T,5))}
\CommentTok{\#}
\CommentTok{\# DT \textless{}{-} data.table(test)}
\CommentTok{\# DT[, cnF := cumsum(test == FALSE)]}
\CommentTok{\# DT[, cnT := cumsum(test == TRUE) ]}
\CommentTok{\# \# the realation of diff may be a general solution}
\CommentTok{\# DT[, ddF := c(0, diff(cnF))]}
\CommentTok{\# DT[, ddT := c(0, diff(cnT))]}
\CommentTok{\#}
\CommentTok{\# DT[, G1  := test]}
\CommentTok{\# DT[, G0  := test]}
\CommentTok{\#}
\CommentTok{\#}
\CommentTok{\# allow \textless{}{-} 1}
\CommentTok{\# for (i in 1:nrow(DT)) \{}
\CommentTok{\#     p1 \textless{}{-} i {-} 1}
\CommentTok{\#     n1 \textless{}{-} i + 1}
\CommentTok{\#     if (p1 \textgreater{} 0 \& n1 \textless{}= nrow(DT)) \{}
\CommentTok{\#         if (DT$G1[p1] == TRUE  \&}
\CommentTok{\#             DT$G1[i]  == FALSE \&}
\CommentTok{\#             DT$G1[n1] == TRUE  ) \{}
\CommentTok{\#             DT$G1[i]  \textless{}{-} TRUE}
\CommentTok{\#         \}}
\CommentTok{\#     \}}
\CommentTok{\# \}}
\CommentTok{\#}
\CommentTok{\# DT[, Grp1 := rleid(c(NA,diff(cumsum(G1))))]}
\CommentTok{\# DT[G1 == FALSE, Grp1 := NA]}
\CommentTok{\#}
\CommentTok{\# DT[, Grp0 := rleid(c(NA,diff(cumsum(G0))))]}
\CommentTok{\# DT[G0 == FALSE, Grp0 := NA]}
\CommentTok{\#}
\CommentTok{\# DT}
\end{Highlighting}
\end{Shaded}

\textbf{END}

\begin{verbatim}
2023-11-21 10:00:59.8 athan@sagan GHI_enh_02_process.R 0.429637 mins
\end{verbatim}

\end{document}