### 2023-11-23 GHI_enh_01_raw_data.tex
Commit 2023-11-23 20:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-11-23}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\#\# Prepare raw data for use in the paper

This collects raw data and prepare it mainly SDR analysis.
It uses the data output of one of the ClearSky algorithms.
It caches the raw data in order to reuse it in next states of analyses.
Creates:
- Raw\_Input.Rds

\begin{verbatim}
## 
##  !! Create raw input data -> ./data/Raw_Input.Rds 
## 
##     Will read 31 input files
## 
##     Load data from Clear Sky process from original
## 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_1993.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_1994.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_1995.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_1996.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_1997.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_1998.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_1999.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2000.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2001.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2002.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2003.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2004.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2005.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2006.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2007.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2008.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2009.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2010.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2011.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2012.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2013.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2014.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2015.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2016.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2017.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2018.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2019.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2020.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2021.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2022.Rds 
## READING: /home/athan/DATA/Broad_Band/CS_id/Clear_sky_id_Reno-Hansen_apply_v14_2_2023.Rds 
## 
##   There are  0  duplicate dates remaining!
## 
##   Extra SKIP of CM-21 data!
## 
##   Select quality data only
\end{verbatim}

\begin{verbatim}
## Written:  ./figures/tbl_tsi_info.dat 
## 
## 
##   Sun Earth distance correction
## 
##   Saved raw input data: ./data/Raw_Input.Rds
\end{verbatim}

\textbf{END}

\begin{verbatim}
2023-11-23 16:53:58.9 athan@sagan GHI_enh_02_process.R 7.216420 mins
\end{verbatim}

\end{document}