### 2023-11-27 article/article.qmd
Commit 2023-11-27 18:14

---
title:    SDR enhancement by clouds
subtitle: A Short Subtitle
author:

  - name:  Athanasios Natsis
    email: natsisphysicist@gmail.com
    affiliations: 
        - id:          auth
          name:        Aristotle University of Thessaloniki
          department:  Laboratory of Atmospheric Physics
          address:     Campus Box 149 
          city:        Thessaloniki
          state:       Greece
          postal-code: 54124
    attributes:
        corresponding: true
    note: This is the first author footnote.

  - name:  Alkiviadis Bais
    email: abais@auth.gr
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.

  - name:  Charikleia Meleti
    email: meleti@auth.gr
    affiliations:
        - ref: auth
    note: Yet another author footnote.

abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography:  [bibliography.bib, Mainreferences.bib]
format:
  elsevier-pdf:
    keep-tex:     true
    journal:
      name:       Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
  docx: default
  html: default

editor_options: 
  chunk_output_type: console
---

# Intro

# Data and methodology

# Results

# Discussion and conclusions

------------------------------------------------------------------------

::: callout-important
## gcff

This is nice
:::

<!--# check this out -->

# Bibliography styles

Here are two sample references: @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. You can sets extra options with `natbiboptions` variable in YAML header. Example

```         
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded and stored locally, or the url can be used as in the example header.

@eq-lowbap

[@Bais1997]

[@Bais2013; @Bennie2008]

# Equations

Here is an equation: $$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$ {#eq-lowbap}

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {.unnumbered}