### 2024-01-15 GHI_enh_03_process.tex
Commit 2024-01-15 12:16

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-01-15}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
Alpha * HAU is CS\_ref

\hypertarget{time-series-decomposition-tests}{%
\subsection{Time series decomposition tests}\label{time-series-decomposition-tests}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_daily, Date, wattGLB.mean)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_daily, Date, wattGLB.sum)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-2} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_daily, Date, wattGLB.max)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-3} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_daily, Date, wattGLB.N)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-4} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_daily, Date, wattGLB.mean)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-5} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_daily, Date, wattGLB.sum)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-6} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_daily, Date, wattGLB.max)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-7} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_daily, Date, wattGLB.N)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-8} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_monthly, Date, wattGLB.mean)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-9} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_monthly, Date, wattGLB.sum )}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-10} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_monthly, Date, wattGLB.max )}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-11} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_monthly, Date, wattGLB.N   )}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-12} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series}\NormalTok{(ST\_E\_monthly, Date, wattGLB.N   ,}\AttributeTok{.smooth=}\NormalTok{T, }\AttributeTok{.smooth\_span =} \FloatTok{0.3}\NormalTok{, }\AttributeTok{.interactive =}\NormalTok{ F)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-13} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_acf\_diagnostics}\NormalTok{(ST\_E\_daily, Date, wattGLB.N, }\AttributeTok{.interactive =}\NormalTok{F, }\AttributeTok{.lags =}\DecValTok{1}\SpecialCharTok{:}\DecValTok{60}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-14} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_stl\_diagnostics}\NormalTok{(ST\_E\_daily, Date, wattGLB.N,}
                     \AttributeTok{.feature\_set =} \FunctionTok{c}\NormalTok{(}\StringTok{"observed"}\NormalTok{,}\StringTok{"season"}\NormalTok{, }\StringTok{"trend"}\NormalTok{, }\StringTok{"remainder"}\NormalTok{),}
                     \AttributeTok{.trend =} \DecValTok{180}\NormalTok{,}
                     \AttributeTok{.frequency =} \DecValTok{30}\NormalTok{,}
                     \AttributeTok{.interactive =}\NormalTok{ F}
\NormalTok{                     )}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
frequency = 30 observations
\end{verbatim}

\begin{verbatim}
trend = 180 observations
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-15} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_stl\_diagnostics}\NormalTok{(ST\_E\_monthly, Date, wattGLB.N,}
                     \AttributeTok{.feature\_set =} \FunctionTok{c}\NormalTok{(}\StringTok{"observed"}\NormalTok{,}\StringTok{"season"}\NormalTok{, }\StringTok{"trend"}\NormalTok{, }\StringTok{"remainder"}\NormalTok{),}
                     \AttributeTok{.trend =} \DecValTok{180}\NormalTok{,}
                     \AttributeTok{.frequency =} \DecValTok{30}\NormalTok{,}
                     \AttributeTok{.interactive =}\NormalTok{ F}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
frequency = 30 observations
trend = 180 observations
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-16} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_seasonal\_diagnostics}\NormalTok{(ST\_E\_daily, Date, wattGLB.N, }\AttributeTok{.interactive =}\NormalTok{ F,}
                          \AttributeTok{.feature\_set =} \FunctionTok{c}\NormalTok{(}\StringTok{"week"}\NormalTok{, }\StringTok{"month.lbl"}\NormalTok{, }\StringTok{"quarter"}\NormalTok{, }\StringTok{"year"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-17} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_seasonal\_diagnostics}\NormalTok{(ST\_E\_monthly, Date, wattGLB.N, }\AttributeTok{.interactive =}\NormalTok{ F)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-18} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# https://business{-}science.github.io/timetk/articles/TK08\_Automatic\_Anomaly\_Detection.html}

\FunctionTok{plot\_time\_series\_regression}\NormalTok{(}
    \AttributeTok{.data         =}\NormalTok{ ST\_E\_monthly,}
    \AttributeTok{.date\_var     =}\NormalTok{ Date,}
    \AttributeTok{.formula      =}\NormalTok{ wattGLB.N }\SpecialCharTok{\textasciitilde{}} \FunctionTok{as.numeric}\NormalTok{(Date)  ,}
    \AttributeTok{.facet\_ncol   =} \DecValTok{2}\NormalTok{,}
    \CommentTok{\# .interactive  = FALSE,}
    \AttributeTok{.show\_summary =} \ConstantTok{TRUE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

Call:
stats::lm(formula = .formula, data = df)

Residuals:
   Min     1Q Median     3Q    Max 
-404.5 -198.5  -82.3  139.7 1090.3

Coefficients:
                   Estimate Std. Error t value Pr(>|t|)    
(Intercept)      -7.444e+01  6.622e+01  -1.124    0.262    
as.numeric(Date)  3.317e-07  5.344e-08   6.207 1.51e-09 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 276.5 on 353 degrees of freedom
Multiple R-squared:  0.09841,   Adjusted R-squared:  0.09585 
F-statistic: 38.53 on 1 and 353 DF,  p-value: 1.513e-09
\end{verbatim}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-19} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lm1 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(ST\_E\_monthly}\SpecialCharTok{$}\NormalTok{wattGLB.N }\SpecialCharTok{\textasciitilde{}}\NormalTok{ ST\_E\_monthly}\SpecialCharTok{$}\NormalTok{Date)}
\FunctionTok{summary}\NormalTok{(lm1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

Call:
lm(formula = ST_E_monthly$wattGLB.N ~ ST_E_monthly$Date)

Residuals:
   Min     1Q Median     3Q    Max 
-404.5 -198.5  -82.3  139.7 1090.3

Coefficients:
                    Estimate Std. Error t value Pr(>|t|)    
(Intercept)       -7.444e+01  6.622e+01  -1.124    0.262    
ST_E_monthly$Date  3.317e-07  5.344e-08   6.207 1.51e-09 ***
---
Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1

Residual standard error: 276.5 on 353 degrees of freedom
Multiple R-squared:  0.09841,   Adjusted R-squared:  0.09585 
F-statistic: 38.53 on 1 and 353 DF,  p-value: 1.513e-09
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot\_time\_series\_regression}\NormalTok{(}
    \AttributeTok{.data         =}\NormalTok{ ST\_E\_daily,}
    \AttributeTok{.date\_var     =}\NormalTok{ Date,}
    \AttributeTok{.formula      =}\NormalTok{ wattGLB.N }\SpecialCharTok{\textasciitilde{}} \FunctionTok{as.numeric}\NormalTok{(Date)  ,}
    \AttributeTok{.facet\_ncol   =} \DecValTok{2}\NormalTok{,}
    \CommentTok{\# .interactive  = FALSE,}
    \AttributeTok{.show\_summary =} \ConstantTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
TypeError: Attempting to change the setter of an unconfigurable property.
TypeError: Attempting to change the setter of an unconfigurable property.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_03_process_files/figure-latex/unnamed-chunk-6-20} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#  Save from environment  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\DocumentationTok{\#\# Variables}
\NormalTok{objects }\OtherTok{\textless{}{-}} \FunctionTok{grep}\NormalTok{(}\StringTok{"\^{}tic$|\^{}tac$|\^{}Script.Name$|\^{}tag$"}\NormalTok{, }\FunctionTok{ls}\NormalTok{(), }\AttributeTok{value =}\NormalTok{ T, }\AttributeTok{invert =}\NormalTok{ T)}
\NormalTok{objects }\OtherTok{\textless{}{-}}\NormalTok{ objects[}\FunctionTok{sapply}\NormalTok{(objects, }\ControlFlowTok{function}\NormalTok{(x)}
    \FunctionTok{is.numeric}\NormalTok{(}\FunctionTok{get}\NormalTok{(x)) }\SpecialCharTok{|}
        \FunctionTok{is.character}\NormalTok{(}\FunctionTok{get}\NormalTok{(x)) }\SpecialCharTok{\&}
        \FunctionTok{object.size}\NormalTok{(}\FunctionTok{get}\NormalTok{(x)) }\SpecialCharTok{\textless{}} \DecValTok{1009} \SpecialCharTok{\&}
\NormalTok{        (}\SpecialCharTok{!}\FunctionTok{is.vector}\NormalTok{(}\FunctionTok{get}\NormalTok{(x)) }\SpecialCharTok{|}
             \SpecialCharTok{!}\FunctionTok{is.function}\NormalTok{(}\FunctionTok{get}\NormalTok{(x))), }\AttributeTok{simplify =}\NormalTok{ T)]}
\DocumentationTok{\#\# Data}
\NormalTok{objects }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}
\NormalTok{    objects,}
    \FunctionTok{grep}\NormalTok{(}\StringTok{"\^{}ST\_"}\NormalTok{, }\FunctionTok{ls}\NormalTok{(), }\AttributeTok{value =} \ConstantTok{TRUE}\NormalTok{)}
\NormalTok{)}

\FunctionTok{save}\NormalTok{(}\AttributeTok{file =} \FunctionTok{paste0}\NormalTok{(}\StringTok{"./data/"}\NormalTok{, }\FunctionTok{basename}\NormalTok{(}\FunctionTok{sub}\NormalTok{(}\StringTok{"}\SpecialCharTok{\textbackslash{}\textbackslash{}}\StringTok{.R"}\NormalTok{, }\StringTok{".Rda"}\NormalTok{, Script.Name))),}
     \AttributeTok{list =}\NormalTok{ objects,}
     \AttributeTok{compress =} \StringTok{"xz"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\textbf{END}

\begin{verbatim}
2024-01-15 09:05:38.4 athan@sagan GHI_enh_03_process.R 1.420055 mins
\end{verbatim}

\end{document}