### 2024-02-09 Article.Rmd
Commit 2024-02-09 09:10

---
title:         "Trends of SDR in Thessaloniki"
# header-includes:
author:
  - name:         Athanasios N. Natsis 
    email:        natsisphysicist@gmail.com
    affiliation:  Laboratory of Atmospheric Physics
#    footnote:     1
    correspondingauthor: true
  - name:         Alkiviadis Bais
    email:        abais@auth.gr 
    affiliation:  Laboratory of Atmospheric Physics
    correspondingauthor: false
address:
  - code:         Laboratory of Atmospheric Physics
    organization: Aristotle University of Thessaloniki
    addressline:  Campus Box 149
    city:         Thessaloniki
    # state:        State
    postcode:     54124
    country:      Greece
#  - code:         Another University
#    organization: Department
#    addressline:  A street 29
#    postcode:     2054 NX
#    city:         Manchester,
#    country:      The Netherlands
# footnote:
#   - code:         1
#     text:         "This is the first author footnote."
#   - code:         2
#     text:         "Another author footnote."

abstract: |
  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
  Observations of SDR at different locations around the world with different environmental

# back matter
keywords: |
  GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols; clouds.

# acknowledgement: |
#   All sources of funding of the study should be disclosed. Please clearly
#   indicate grants that you have received in support of your research work.
#   Clearly state if you received funds for covering the costs to publish in open
#   access.

# authorcontributions: |
#   For research articles with several authors, a short paragraph specifying their
#   individual contributions must be provided. The following statements should be
#   used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the
#   experiments; X.X. and Y.Y. analyzed the data; W.W. contributed
#   reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
#   limited to those who have contributed substantially to the work reported.

funding: |
  This research received no external funding.

journal:         "An awesome journal"
date:            "`r Sys.Date()`"
classoption:     preprint, 3p, authoryear
bibliography:    Mainreferences.bib
linenumbers:     false
numbersections:  true
abbreviations:
  - short: DNI
    long: Direct beam/normal irradiance
  - short: CSid
    long: Clear sky identification algorithm
  - short: CUSUM
    long: Cumulative sum
  - short: SDR
    long: Solar downward radiation
  - short: SZA
    long: Solar zenith angle

header-includes:
   # - \usepackage{subfig}
   - \usepackage{subcaption}
   - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
   - \usepackage{placeins}

output: 
  rticles::mdpi_article:
    extra_dependencies:  longtable
    keep_tex:            yes
  bookdown::word_document2: default
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              no
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
---

```{r echo=F, include=T, warning=FALSE, message=FALSE}
suppressMessages(
    res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
)
## document configurations
knitr::opts_chunk$set(comment    = ""      )
# knitr::opts_chunk$set(dev        = "pdf"   )
knitr::opts_chunk$set(out.width  = "70%"   )
knitr::opts_chunk$set(fig.align  = "center")
knitr::opts_chunk$set(fig.pos    = 'h!'    )
options(knitr.graphics.auto_pdf = FALSE)
library(pander)
library(data.table)
library(kableExtra)
library(dplyr)
## load data from trends analysis
source("./DHI_GHI_00_raw_data.R")
load(common_data)
load("./data/Input_1_longterm_trends.Rda")
## number of measurements for each dataset
SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N])
SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])
## TSI data span info
TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
                    strip.white = TRUE,
                    sep = ";",
                    header = TRUE))
## variables we use in the analysis
source("./DHI_GHI_0_variables.R")

## create local citations file and links here
source("~/CODE/FUNCTIONS/R/gather_citations.R")
suppressMessages(
    gather_citations(
        sourcefile   = "./Article.Rmd",
        mainreposit  = "./Mainreferences.bib",
        localreposit = "./references.bib",
        folderlink   = "./References_links",
        workingdir   = "./"
    )
)
```

# Introduction.

# Observational data and methodology.

# Results

# Conclusions

# References {-}

<div id="refs"></div>

# Appendix {-}