### 2024-02-22 article/article.qmd
Commit 2024-02-22 11:14

---
title:    SDR enhancement by clouds
subtitle: A Short Subtitle
author:

  - name:  Athanasios N. Natsis
    email: natsisphysicist@gmail.com
    affiliations: 
        - id:          auth
          name:        Aristotle University of Thessaloniki
          department:  Laboratory of Atmospheric Physics
          address:     Campus Box 149 
          city:        Thessaloniki
          state:       Greece
          postal-code: 54124
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name:  Alkiviadis Bais
    email: abais@auth.gr
    affiliations:
        - id: another-u
          name: Another University
          department: Department Name
          address: Street Address
          city: City
          state: State
          postal-code: Postal Code
    note: |
      Another author footnote, this is a very long footnote and it should be a really long footnote. But this footnote is not yet sufficiently long enough to make two lines of footnote text.
  - name:  Charikleia Meleti
    email: meleti@auth.gr
    affiliations:
        - ref: auth
    note: Yet another author footnote.
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date:          last-modified
bibliography:  [bibliography.bib]
format:
  elsevier-pdf:
    keep-tex:     true
    journal:
      name:       Journal Name
      formatting: preprint
      model:      3p
      cite-style: authoryear
  docx: default
  html: default
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=F, include=T, warning=FALSE, message=FALSE}
suppressMessages(
    res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
)
## document configurations
knitr::opts_chunk$set(comment    = ""      )
# knitr::opts_chunk$set(dev        = "pdf"   )
knitr::opts_chunk$set(out.width  = "70%"   )
knitr::opts_chunk$set(fig.align  = "center")
knitr::opts_chunk$set(fig.pos    = 'h!'    )
options(knitr.graphics.auto_pdf = FALSE)
library(pander)
library(data.table)
library(kableExtra)
library(dplyr)
## load data from trends analysis
source("~/MANUSCRIPTS/02_enhancement/GHI_enh_00_variables.R")
TSIinfo <- data.table(read.csv("~/MANUSCRIPTS/02_enhancement/figures/tbl_tsi_info.dat",
                    strip.white = TRUE,
                    sep = ";",
                    header = TRUE))
# ## create local citations file and links here
# source("~/CODE/FUNCTIONS/R/gather_citations.R")
# suppressMessages(
#     gather_citations(
#         sourcefile   = "./Article.Rmd",
#         mainreposit  = "./Mainreferences.bib",
#         localreposit = "./references.bib",
#         folderlink   = "./References_links",
#         workingdir   = "./"
#     )
# )
```

CE: cloud enhancement cases one minute

CEG: continuous group of cloud enhancement cases

# Intro

# Data and methodology

The recording of the signal is the one minute average.
Thus we will refer to this time scale as a cloud enhancement event (CE).
Due to the instruments response time.

## Data quality control

## Data selection

To preserve an unbiased representation of the data we applied a constrain similar the one used by @CastillejoCuberos2020. Where, for each valid hour of day, there must exist at list 45 minutes of valid measurements, including night time near sunrise and sunset. Days with less than 5 valid hours are rejected completely. <!--# CastillejoCuberos2020 There must be at least 5 h of valid data (Tests 1–7), otherwise all data for this day is flagged as invalid. There must be at least This work valid data 45 min of valid data (Tests 1–7) for an hour to be flagged as valid, otherwise all minute data for this hour is flagged as invalid. --> Furthermore, due to a systematic obstruction of the horizon and the significant measurement uncertainty on large SZAs, we excluded all data with $SZA > `r 90 - BIO_ELEVA`^\circ$.

## Clear sky identification

Use it to find some interesting days to manual check the CE thresholds. No other usage of clear/cloud characterization in CE.

## Criteria for enhancement based on TSI

Many have used models or some statistical method to determine then CE thresholds.

....... cite ...... Descriptions

We used as reference the .... extraterrestrial horizontal irradiance .....

$$
\text{ETH} = \cos(\theta) \times \text{TSI}_\text{TOA}
$$

$$
\text{GHI} > 999 \text{ETH}
$$

$$
f(x)=\begin{cases}
1, & \text{if $x<0$}.\\
0, & \text{otherwise}.
\end{cases}
$$

-   some other were tested

Thresholds by inspecting a selection of days with some specs

# Results

-   max
-   seasonal
-   trends
-   duration
-   occurrence
-   SZA

# Discussion and conclusions

------------------------------------------------------------------------

::: callout-important
## gcff

This is nice
:::

<!--# check this out -->

# Bibliography styles

Here are two sample references: @Feynman1963118 @Dirac1953888.

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. You can sets extra options with `natbiboptions` variable in YAML header. Example

```         
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL

If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded and stored locally, or the url can be used as in the example header.

@eq-lowbap

[@Bais1997]

# Equations

Here is an equation: $$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$ {#eq-lowbap}

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {.unnumbered}