### 2024-04-02 Article_B4/Article_B4.Rmd
Commit 2024-04-02 13:13

---
title:    SDR enhancement by clouds

subtitle: A Short Subtitle

author:
  - Athanasios N. Natsis^[Laboratory of Atmospheric Physics _natsisphysicist@gmail.com_] 
  - Alkiviadis Bais^[Laboratory of Atmospheric Physics _abais@auth.gr_] 
  - Charikleia Meleti^[Laboratory of Atmospheric Physics _meleti@auth.gr_]

# abstract: |
#   This is the abstract. Lorem ipsum dolor sit amet, eu efficitur justo rutrum at. Sed at tempus elit.

keywords:
  - keyword1
  - keyword2

bibliography:  [bibliography.bib]

date: "`r format(Sys.time(), '%F')`"

header-includes:
  - \usepackage{caption}
  - \usepackage{placeins}
  - \captionsetup{font=small}

output:
  # redoc::redoc: default
 #  rticles::mdpi_article:
 #    extra_dependencies:  longtable
 #    keep_tex:            yes
  bookdown::pdf_document2:
    number_sections:  yes
    fig_caption:      yes
    keep_tex:         yes
    latex_engine:     xelatex
    toc:              yes
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
    bookdown::word_document2: default
---

```{r setup, echo=F, include=FALSE}
rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
renv::load("~/MANUSCRIPTS/02_enhancement/", quiet = TRUE)
## document configurations
knitr::opts_chunk$set(comment    = ""      )
# knitr::opts_chunk$set(dev        = "pdf"   )
knitr::opts_chunk$set(out.width  = "60%"   )
knitr::opts_chunk$set(fig.align  = "center")
knitr::opts_chunk$set(fig.pos    = 'h!'    )
options(knitr.graphics.auto_pdf = FALSE)
library(pander)
library(data.table)
library(kableExtra)
library(dplyr)
## load data from trends analysis
source("~/MANUSCRIPTS/02_enhancement/GHI_enh_00_variables.R")
load("~/MANUSCRIPTS/02_enhancement/data/GHI_enh_03_process.Rda")
TSIinfo <- data.table(read.csv("~/MANUSCRIPTS/02_enhancement/figures/tbl_tsi_info.dat",
                               strip.white = TRUE,
                               sep = ";",
                               header = TRUE))
```

# Abstract {.unnumbered}

- Phenomenon
- Effects
- bibliography
- our approach
- our results

CE: cloud enhancement cases one minute

ECE: extreme cloud enhancement cases one minute over TSI on horizontal plane

CEG: cloud enhancement groups, cases events with consecutive CE

$\text{GHI}_\text{i}$: measured one minute global horizontal irradiance

$\text{GHI}_\text{mCSi}$: modeled clear sky one minute global horizontal irradiance

# Intro

The shortwave solar irradiance, reaching Earth's surface, is the main energy source
of the atmosphere and biosphere and drives and governs the climate [@Gray2010].  It
has direct practical application, in different industries, like energy production and
agriculture method.  The variability of its intensity, can cause difficulties in
predicting the yield, and designing the specifications of the appropriate equipment.
A lot of research has been focused on predicting the renewable energy production in a
fine timescale and in near real-time (for a review see @Inman2013; @Graabak2016).

A big aspect of this variability is the interaction with the clouds. In general,
clouds absorb part of the solar irradiance, but under certain conditions, can enhance
the total shortwave irradiance reaching the ground. This effect, can locally increase
the observed total shortwave irradiance higher than the expected clear sky irradiance
[see references therein].

Some of the proposed underling mechanisms of those events, have been summarized by
@Gueymard2017, and include scattering on the edge of cumulus clouds or through thin
cirrus.  Further investigation with radiation transfer modeling methods and
observations have pointed as the prevailing mechanism, the forward Mie scattering
[@Pecenak2016; @Thuillier2013; @Yordanov2013; @Yordanov2015], through the clouds.
The overall phenomenon depends on different interactive factors, that include cloud
thickness, constitution and type; and the relative position of the sun, the clouds,
and the observer [@Gueymard2017; @Veerman2022].  As such, there are multiple
contributing mechanism that are responsible for the observed irradiance enhancements.

Cloud enhancements have been reported, to be able to exceed in intensity even the
value of the solar constant, resulting of clear indexes above unit. A summary of
extreme enchantment cases have been compiled by @Martins2022. There are also some
practical implication of the cloud enchantments.  The intensity and duration of
enhancements can effect the efficiency and stability of photovoltaic power production
[@Lappalainen2020; @Jaervelae2020], and extreme enhancements cases, have the
potential to compromise the integrity of photovoltaic plants infrastructure
[@DoNascimento2019].  It has also been demonstrated, that these events can interfere
in the comparison of ground data and satellite observations [@Damiani2018]

<!--
Study of the mechanism has also been done with modeling of different cloud types and
atmospheric conditions.  @Gueymard2017; @Veerman2022
-->

<!--
Instruments respose ....
@Martins2022
-->

Methods of identification cloud enchantment events in the literature, usually include
the use of a simulated clear sky radiation as baseline, that is combined with an
appropriate threshold or some other statistical characteristics, and in some cases,
with visual methods with a sky cam [@Vamvakas2020; @Mol2023 and references therein].

<!-- our work -->
In this study, we evaluated the effects cloud enhancements on the total downward
radiation by studding the occurrences, their intensity, and their duration in a
thirty-year period at the city of Thessaloniki.  We used modeled clear sky
irradiance, as a baseline to identify cloud enhancements.  We were able to determine
some trends of the phenomenon, it's Climatology, and some of their  general
characteristics. We weren't able to find a comparable study that provides trends for
similar long term dataset, as ours.  The recording of the radiation signal is the
mean of one-minute. Thus, the minimum resolution of cloud enhancement events (CE), in
this study is one minute.

<!-- Define CE and ECE -->
In the relative bibliography different definition have been used for these events,
some of the are summarized by @Gueymard2017.  Here, we defined as cloud enhancement
events (CE) the cases when the measured global horizontal irradiance (GHI) at ground
level, exceeds the expected value under clear-sky conditions.  Similar, we define as
extreme cloud enhancement events (ECE), the cases when GHI, exceeds the Total Solar
Irradiance on horizontal plane at ground level.  Although the duration of these
bursts varies, from instantaneous to several minutes, here we are constrained by the
recorded data, to one-minute steps.

<!--
Identification methods from bib

- clearness index
- modeled approaches
- algorithmic id
- AI
- Visual

Constrains of the site

- data
- instrument
- location
-->

# Data and methodology

## GHI data

The monitoring site is operating in the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m\ a.s.l.).
In this study we present data from the period
`r strftime(min(ST_daily$Date), "%d\\ %B %Y")` to
`r strftime(max(ST_daily$Date), "%d\\ %B %Y")`.
<!-- Instrument -->
The GHI data were measured with a Kipp\ & Zonen CM-21 pyranometer.  During the study
period, the pyranometer was independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, verifying that the stability of the
instrument’s sensitivity was better than $0.7\,\%$ relative to the initial
calibration by the manufacturer.
<!-- Measurements -->
For the acquisition of radiometric data, the signal of the pyranometer was sampled at
a rate of $1\,\text{Hz}$.  The mean and the standard deviation of these samples were
calculated and recorded for every minute.  The measurements were corrected for the
zero offset ("dark signal" in volts), which was calculated by averaging all
measurements recorded for a period of $3\,\text{h}$, before (morning) or after
(evening) the Sun reaches an elevation angle of $-10^\circ$.  The signal was
converted to irradiance using the ramped value of the instrument’s sensitivity
between subsequent calibrations.

<!-- Radiation data quality check -->
To further improve the quality of the irradiance data, a manual screening was
performed to remove inconsistent and erroneous recordings that can occur
stochastically or systematically during the continuous operation of the instruments.
The manual screening was aided by a radiation data quality assurance procedure,
adjusted for the site, which was based on the methods of Long and Shi\ [@Long2006;
@Long2008a].  Thus, problematic recordings have been excluded from further
processing.  Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality of the
radiation measurements used in this study.
<!-- Data selection -->
To preserve an unbiased representation of the data we applied a constraint, similar
the one used by @CastillejoCuberos2020. Where, for each valid hour of day, there must
exist at least 45 minutes of valid measurements, including nighttime near sunrise and
sunset. Days with less than 5 valid hours are rejected completely.
<!--# CastillejoCuberos2020 There must be at least 5 h of valid data (Tests 1–7),
otherwise all data for this day is flagged as invalid. There must be at least This
work valid data 45 min of valid data (Tests 1–7) for an hour to be flagged as valid,
otherwise all minute data for this hour is flagged as invalid. -->
Furthermore, due to the significant measurement uncertainty when the Sun is near the
horizon, and due to some systematic obstructions by nearby buildings, we have
excluded all measurements with solar zenith angle (SZA) greater than 
$`r 90 - BIO_ELEVA`^\circ$.

## Cloud enhancement detection

To be able to detect the CE cases, we had to establish a baseline, above which we can
characterize each data point as an enhancement event, by estimating the occurring
over irradiance (OIR). The OIR, here is defined as the irradiance difference of the
measured one-minute $\text{GHI}_i$ from the CE identification criterion in
Equation\nobreakspace\ref{eq:CE4}
($\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}$).
To have an estimation, and a first insight on the phenomenon, we experimented with two
simple approaches for the reference.  The Haurwitz's model [@Haurwitz1945], which is
as simple clear sky model, and we had already adjusted and had good fit with our data
[@Natsis2023], and the TSI at the top of the atmosphere.  We have tested both cases
by using an appropriate relative threshold and/or an additional constant offset.  The
initial results, showed that we can detect a big portion of the CE events. These
results were helpful, and helped to establish some criteria to further improve the CE
identification. It was evident, by inspecting the daily plot of irradiance, that
changes on the atmospheric conditions introduced numerous false positive and false
negative results.  To produce a more accurate reference, we had to take into account
more factors that effect the clear sky radiation. So we used a radiation transfer
model.

## Modeled clear Sky Irradiance

We approximated the expected clear sky GHI by using a model.  Because of the lack of
observational data for the whole period, we used some long term climatological data
for the main factors responsible for the attenuation of the broadband downward solar
radiation in the atmosphere. Which are mainly, the aerosols and the water vapors.
Fortunately, our site participates in the Aerosol Robotic Network (AERONET)
[@Giles2019; @Buis1998], as we operate a Cimel photometer since 2003, collocated with
the CM-21 pyranometer.  The mean monthly aerosol optical depth (AOD) on different
wavelengths is provided by AERONET, along with the equivalent water column height in
the atmosphere.  The simulation of the clear sky radiation was performed by the well
established radiation transfer model Libradtran [@Emde2016], a similar
approach, was also used by @Vamvakas2020 for creating a clear sky reference.

<!-- Libradtran options -->
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated for combinations of AOD at
$500\,\text{nm}$ ($t_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w_h$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].

<!-- Lookup table -->
To create a look-up table that aligns with our dataset, we applied some adjustments.
To account for the Sun's variability in our one-minute GHI measurements, we corrected
the model's input spectrum integral, to the one of the total solar irradiance (TSI)
provided by NOAA [@Coddington2005].  Also, we applied the effect of the Earth -- Sun
distance on the irradiance, by using the distance calculated by the Astropy
[@AstropyCollaboration2022] software library. As needed, we interpolate the resulting
irradiances to the exact SZA of our measurements.  For each period of the year, we
used the appropriate atmospheric profile (afglms or afglmw).  Finally, we calculated
the clear sky irradiance value at the horizontal plane.  Thus, we were able to
emulate different atmospheric condition and levels of atmospheric clearness for the
climatological conditions of the site.  With this method, the modeled clear sky
irradiances can be directly compared to each measured one-minute value of GHI, for
different conditions of atmospheric clearness.

## CE Criteria investigation

The use of the actual modeled values of clear sky GHI alone, can not provide us with
a robust method to distinguish the CE cases, due to the limited accuracy of the input
data. Our main focus was to positively identify over irradiance events from CEs, thus
we used a relative factor, to create an upper envelope of the clear sky irradiance,
above which, any GHI value can safely attributed to CE.  We evaluated the performance
of the modeled clear radiation for each of atmospheric level of clearness, as
reference, in order to conclude which is the most appropriate.

To select the exact values of these thresholds factors (Equation \ref{eq:CE4}), we
implemented an empirical method, by manual inspection of the CE identification, on
specially selected days from the whole dataset.  We used seven sets of selected days,
with characteristics relevant to the efficiency of the identification threshold.
These sets were random groups of about 20 to 30 days with the following
characteristics:
(a) the largest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) without clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) without clouds and without EC events,
(e) with at least $60\,\%$ of the day length without clouds and some EC events,
(h) random days and
(i) some manual selected days that were included during the manual inspection.
Where it was needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

After evaluating the modeled clear radiation for the different atmospheric
conditions, in relation to the measured GHI data, we choose as a representative of
the clear sky radiation, the case where
$t_{\text{cs}} = t_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{mCSi}$).
These values represent a typical atmosphere in Thessaloniki with low load of aerosols
and humidity, which are the main factor that attenuate the GHI, excluding clouds.  To
create the limit of CE identification, we created a two branched threshold, as a
function of SZA. A constant factor for low SZAs, and a higher ramped factor for
higher SZAs (Equation\nobreakspace\ref{eq:CE4}).
This is the criterion of our CE identification.
<!--
\begin{equation}
\text{CE}_\text{Threshold} = \begin{cases}
`r C4_cs_ref_ratio`, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) + `r C4_lowcut_ratio` , & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4a}
\end{equation}
where: $\theta$ solar zenith angle.
-->
<!--
\begin{equation}
\text{CE} : \begin{cases}
 \text{GHI}_\text{i} > `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{i} > \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4b}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modeled clear sky irradiance.
-->
\begin{equation}
\text{CE} : \text{GHI}_\text{i} > \text{GHI}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{GHI}_\text{CSlim,i} = `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{CSlim,i} = \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modeled clear sky irradiance.

<!-- reasoning for cretria -->
We have to note, that the differentiation of the threshold factor was needed, because
of the high irradiance values we observed, early in the morning and late in the
afternoon. We have confirmed, by inspecting images from the sky cam, that we have
duration of elevated irradiance, either due to the clearness of the atmosphere, or some
interferences by reflections on nearby bright surfaces. Although, these cases may
mask some of the CE events. The actual CE events, produce higher irradiance and thus
are identified as such. As a side effect, the reported OIR will be slight
underestimated for those SZAs, but the overall contribution of those cases to the
total daily energy, is minimal due the low occurrences and the lower irradiances.

<!--
- can follow for clear days
- this selection has some subjectivity
- the most important are the thresholds
- model have to be uniform on clear days
- the level is adjusted by the threshold 
- the effect of the sza uniform
-->

.............

These selections have some subjectivity, as the definition of clear sky, is depended
on the intended usage. Our main focus is to be able to identify the OIR created by the
clouds. Although, the aerosols and water vapor are always present and can not be
completely removed. 
The effect of the different clearness levels

.............

<!-- extreme cases criterion -->
Another aspect of the CE events, that is often reported in the relative bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and satisfy the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times \text{TSI}_\text{iTOA}
\label{eq:ECE}
\end{equation}

.......

- Include Example of days plot in the Appendix?
... this is clearer to understand and describe ....

.......

<!-- ```{r echo=F, include=T, warning=FALSE, message=FALSE} -->
<!-- ST_total -->
<!-- cat("\n - - - \n") -->
<!-- ST_daily[which.max(wattGLB.max), wattGLB.max] -->
<!-- round(ST_E_daily[which.max(GLB_diff.max), GLB_diff.max], 2) -->
<!-- round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1) -->
<!-- plot(ST_daily$Date, ST_daily$GLB_diff.max) -->
<!-- ``` -->

\FloatBarrier

# Results

Our dataset, after the data selection processing, consists of
`r ST_total$wattGLB.N` records of GHI, of which
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are CE and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are ECE events.
The highest GHI recorder was 
$`r round(ST_daily[which.max(wattGLB.max), wattGLB.max],1)`\,W/m^2$
on `r strftime(ST_daily[which.max(wattGLB.max), Date], "%d\\ %B %Y")`.
The absolute stronger CE event had an OIR of 
$`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 2)`\,W/m^2$ on
`r strftime(ST_daily[which.max(GLB_diff.max), Date], "%d\\ %B %Y")`.
The relative stronger CE event was 
$`r round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1)`\,\%$ above the 
clear sky threshold, on 
`r strftime(ST_daily[which.max(GLB_ench.max), Date], "%d\\ %B %Y")`.

## Trends

```{r loadtrends, echo=F, include=T, warning=FALSE, message=FALSE}
dailytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear.csv"))
dailyCE <- dailytrends[DATA == "ST_E_daily"]
yearlytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear_Proper.csv"))
yearlyCE <- yearlytrends[DATA == "ST_E_yearly"]

```

We computed the daily trend of the mean OIR of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order autoregressive
model with lag of 1 day, using the 'maximum likelihood' fitting method [@Gardner1980;
@Jones1980] by implementing the function 'arima' from the library 'stats' of the R
programming language [@RCT2023]. The trends are reported together with the $2\sigma$
error.  We observe an increase of
$`r round(dailyCE[var == "GLB_diff.mean", Tmod_Estimate], 3)`\pm
`r round(2 * dailyCE[var == "GLB_diff.mean", Tmod_Std..Error], 3)`\,W/m^2$
on the mean OIR.

```{r CEmeanDaily, echo=FALSE, fig.cap="Daily mean values and trend of the CE over irradiance."}
knitr::include_graphics("../images/P_daily_trend-1.png")
```

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements.  The annual
mean OIR is
$`r round(yearlyCE[var == "GLB_diff.mean", slope], 2)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 2)`\,W/m^2$
(Figure\nobreakspace{}\ref{fig:P-energy-mean}),
but this value alone is not very useful due the intrinsic high variability of this
metric.

<!-- ```{r CEmeanDaily2, echo=FALSE, fig.cap="Trend of the daily excess energy due to CE"} -->
<!-- knitr::include_graphics("../images/daily-12.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Treds of the mean over irradiance per CE"} -->
<!-- knitr::include_graphics("../images/energy-3.png") -->
<!-- ``` -->

```{r P-energy-mean, echo=FALSE, fig.cap="Trends of the mean OIR per CE."}
knitr::include_graphics("../images/P_energy-7.png")
```

A better indicator of changes on the characteristic of CE would be the number of CE
occurrences and the total energy of the CE over irradiance. The annual number of CE
occurrences, shows a steady increase of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,/y$
(Figure\nobreakspace{}\ref{fig:P-energy-N}).
We have to note that the energy related to the CE events can not be directly linked
with the total energy balance on the atmosphere. The net sun radiation of the region
is not increased, but rather redistributed through the CE. Although, the
instantaneous values, can exceed the equivalent clear sky irradiance to a
considerable level.

<!-- ```{r echo=FALSE, fig.cap="Trend of number of CE cases"} -->
<!-- knitr::include_graphics("../images/energy-2.png") -->
<!-- ``` -->

```{r P-energy-N, echo=FALSE, fig.cap="Trend of yearly CE number of occurancies."}
knitr::include_graphics("../images/P_energy-6.png")
```
Subsequently, the total annual energy observed during CE events shows an increase of
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,kJ/y$
(Figure\ \ref{fig:P-energy-sum}), witch follows the trend of the number of
occurrences.

<!-- ```{r echo=FALSE, fig.cap="Trend of the yearly excess energy due to CE over irradiance"} -->
<!-- knitr::include_graphics("../images/energy-1.png") -->
<!-- ``` -->

```{r P-energy-sum, echo=FALSE, fig.cap="Trend of the yearly excess energy due to CE over irradiance"}
knitr::include_graphics("../images/P_energy-5.png")
```

```{r P-energy-median, echo=FALSE, fig.cap="Trend of the yearly median over irradiance due to CE over irradiance"}
knitr::include_graphics("../images/P_energy-8.png")
```

\FloatBarrier

## Climatology

Another interesting aspect of the CE cases, is their seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurancies}, we have the box plot
(whisker plot), where the values have been normalized by the highest median value,
that occurs in May.  Although the number of occurrences has a wide spread throughout
the study period, the most active period of CE occurrences is during May and June.
During the Winter (December -- February) the CE cases are about $25\,\%$ of the maximum.
The rest of the months the occurrences ramp between the maximum and minimum.

```{r relative-month-occurancies, echo=FALSE, fig.cap="Statistics of the number of CE occurancies for each month. The box represents the values of the low $25\\,\\%$ percentile to $75\\,\\%$ percentile, where the thick horizontal line inside is the mean, the verical lines extend to the macimum and minimum vales, the dots are outlier values, and the rhombus is the mean."}
knitr::include_graphics("../images/clim_CE_month_norm_MAX_median_N-2.png")
```
<!--
```{r echo=FALSE, fig.cap="Distribution of the number of CE cases"}
knitr::include_graphics("../images/climCEmonth-1.png")
```
--->

<!-- ```{r echo=FALSE, fig.cap="Distribution of CE above reference"} -->
<!-- knitr::include_graphics("../images/relative_distributions-2.png") -->
<!-- ``` -->
<!--
```{r echo=FALSE, fig.cap="Distribution of CE above reference"}
knitr::include_graphics("../images/relative_distributions-4.png")
```
-->

The distribution of the CE over irradiance spreads uniformly
Figure\nobreakspace{}\ref{fig:ovir-distribution}. Where there is an inverse relation
between the events frequency and events intensity. This is expected as, the stronger
the CE events become, the rarer the particular atmospheric and sun conditions
occur.

```{r ovir-distribution, echo=FALSE, fig.cap="Distribution of CE over irradiance"}
knitr::include_graphics("../images/P-relative-distribution-diff-1.png")
```

<!-- whisky plots relative to GLB -->

<!-- set of yearly plots -->

\FloatBarrier

## Groups stats

In order to further study the characteristics of the CE events, we grouped the single
minute CE events, to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can have a duration of multiple of minutes.  We have
identified `r ST_G0[, .N]` groups, where the group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude that
longer durations are getting increasingly rare, with durations above 10 minutes very
rare.

```{r ceg-duration-distribution, echo=FALSE, fig.cap="Distribution of CE groups"}
knitr::include_graphics("../images/groups-1.png")
```

... @Zhang2018 ....

```{r echo=FALSE, fig.cap="Relation of mean over irradiance and CE group duration"}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

<!-- ```{r echo=FALSE, fig.cap="Relation of mean over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-4.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Relation of maximun over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-8.png") -->
<!-- ``` -->

\FloatBarrier

## Extreme CE above TSI

Another aspect of the CE events are the cases when the irradiance is above the
expected irradiance on top of the atmosphere, we have defined this as ECE
(Equation\nobreakspace{}\ref{eq:ECE}).

-  max value ECE
  relative to other works max values

Analogous to Figure\nobreakspace{}\ref{fig:relative-month-occurancies} we have
computed the distribution of the number of occurrences of ECE events by month in
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}. Here the most active
period is in the spring (March -- May), followed by the months of June, September and
October.

.......

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Distribution of "}
knitr::include_graphics("../images/clim_ECE_month_norm_MAX_median_N-2.png")
```

<!-- ```{r echo=FALSE, fig.cap="Distribution of the number of ECE cases"} -->
<!-- knitr::include_graphics("../images/climECEmonth-1.png") -->
<!-- ``` -->

```{r echo=FALSE, fig.cap="Distribution of ECE above 'clear sky' reference"}
knitr::include_graphics("../images/extremedistributions-2.png")
```

<!-- ```{r echo=FALSE, fig.cap="Distribution of ECE Irradiance above 'TSI'"} -->
<!-- knitr::include_graphics("../images/extremedistributions-4.png") -->
<!-- ``` -->

-   SZA

<!-- ### Yearly plots relative to all GHI data. -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-1.png") -->
<!-- ``` -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-2.png") -->
<!-- ``` -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-3.png") -->
<!-- ``` -->

\FloatBarrier

# Discussion and conclusions

Compare to other location stats.
max, ECE, distributions ....

Climatology results

# Appendix {.unnumbered}

# References {.unnumbered}