### 2024-04-10 high_aod/GHI_enh_02_ID_CE.tex
Commit 2024-04-10 12:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-04-09}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
[1] "/home/athan/MANUSCRIPTS/02_enhancement/high_aod"
\end{verbatim}

\begin{verbatim}
[1] "Cloud" "Clear"
\end{verbatim}

\begin{verbatim}
 [1] "Exact_B.High_W.edir"  "Exact_B.High_W.edn"   "Exact_B.High_W.eup"  
 [4] "Exact_B.Low_2_W.edir" "Exact_B.Low_2_W.edn"  "Exact_B.Low_2_W.eup" 
 [7] "Exact_B.Exact_W.edir" "Exact_B.Exact_W.edn"  "Exact_B.Exact_W.eup" 
[10] "Exact_B.Low_W.edir"   "Exact_B.Low_W.edn"    "Exact_B.Low_W.eup"   
\end{verbatim}

\begin{verbatim}
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  1.031   1.034   1.035   1.035   1.035   1.036 
\end{verbatim}

Alpha * HAU is CS\_ref

\hypertarget{using-criteria-enhanc_c_4-for-final-application}{%
\subsection{\texorpdfstring{Using criteria \textbf{Enhanc\_C\_4} for final application}{Using criteria Enhanc\_C\_4 for final application}}\label{using-criteria-enhanc_c_4-for-final-application}}

\hypertarget{use-tsi-as-reference-for-clear-sky.}{%
\subsection{1. Use TSI as reference for Clear sky.}\label{use-tsi-as-reference-for-clear-sky.}}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# \_\_ 1. My criteria  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\CommentTok{\# C1\_GLB\_ench\_THRES     \textless{}{-}  1.10 \#\# enchantment relative to HAU}
\NormalTok{C1\_GLB\_diff\_THRES     }\OtherTok{\textless{}{-}} \DecValTok{20}    \DocumentationTok{\#\# enchantment absolute diff to HAU}
\NormalTok{C1\_Clearness\_Kt\_THRES }\OtherTok{\textless{}{-}}  \FloatTok{0.8}  \DocumentationTok{\#\# enchantment threshold}
\CommentTok{\# C1\_wattGLB\_THRES      \textless{}{-} 20    \#\# minimum value to consider}

\NormalTok{DATA[, Enhanc\_C\_1 }\SpecialCharTok{:=} \ConstantTok{FALSE}\NormalTok{]}
\CommentTok{\# DATA[wattGLB           \textgreater{} CS\_ref * C1\_GLB\_ench\_THRES + C1\_GLB\_diff\_THRES \&}
\CommentTok{\#      ClearnessIndex\_kt \textgreater{} C1\_Clearness\_Kt\_THRES,}
\CommentTok{\#      Enhanc\_C\_1 := TRUE]}

\NormalTok{DATA[, Enhanc\_C\_1\_ref }\SpecialCharTok{:=}\NormalTok{ ETH }\SpecialCharTok{*}\NormalTok{ C1\_Clearness\_Kt\_THRES }\SpecialCharTok{+}\NormalTok{ C1\_GLB\_diff\_THRES]}
\NormalTok{DATA[wattGLB }\SpecialCharTok{\textgreater{}}\NormalTok{ Enhanc\_C\_1\_ref,}
\NormalTok{     Enhanc\_C\_1 }\SpecialCharTok{:=} \ConstantTok{TRUE}\NormalTok{]}

\ControlFlowTok{if}\NormalTok{ (SelEnhanc }\SpecialCharTok{==} \StringTok{"Enhanc\_C\_1"}\NormalTok{) \{}
\NormalTok{    DATA[ , GLB\_diff }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_1\_ref                    ] }\DocumentationTok{\#\# enhancement}
\NormalTok{    DATA[ , GLB\_ench }\SpecialCharTok{:=}\NormalTok{ ( wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_1\_ref ) }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_1\_ref ] }\DocumentationTok{\#\# relative enhancement}
\NormalTok{    DATA[ , GLB\_rati }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_1\_ref                    ]}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{use-tsi-as-reference-for-clear-sky.-1}{%
\subsection{2. Use TSI as reference for Clear sky.}\label{use-tsi-as-reference-for-clear-sky.-1}}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# \_\_ 2. Gueymard2017 Criteria  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\DocumentationTok{\#\# Clearness index \textgreater{} 0.8 / 1}
\NormalTok{C2\_Clearness\_Kt\_THRES }\OtherTok{\textless{}{-}} \FloatTok{0.8}
\NormalTok{DATA[, Enhanc\_C\_2 }\SpecialCharTok{:=} \ConstantTok{FALSE}\NormalTok{]}
\NormalTok{DATA[, Enhanc\_C\_2\_ref }\SpecialCharTok{:=}\NormalTok{ ETH }\SpecialCharTok{*}\NormalTok{ C2\_Clearness\_Kt\_THRES]}
\CommentTok{\# DATA[ClearnessIndex\_kt \textgreater{} C1\_Clearness\_Kt\_THRES,}
\CommentTok{\#      Enhanc\_C\_2 := TRUE]}
\NormalTok{DATA[wattGLB }\SpecialCharTok{\textgreater{}}\NormalTok{ Enhanc\_C\_2\_ref,}
\NormalTok{     Enhanc\_C\_2 }\SpecialCharTok{:=} \ConstantTok{TRUE}\NormalTok{]}

\ControlFlowTok{if}\NormalTok{ (SelEnhanc }\SpecialCharTok{==} \StringTok{"Enhanc\_C\_2"}\NormalTok{) \{}
\NormalTok{    DATA[ , GLB\_diff }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_2\_ref                    ] }\DocumentationTok{\#\# enhancement}
\NormalTok{    DATA[ , GLB\_ench }\SpecialCharTok{:=}\NormalTok{ ( wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_2\_ref ) }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_2\_ref ] }\DocumentationTok{\#\# relative enhancement}
\NormalTok{    DATA[ , GLB\_rati }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_2\_ref                    ]}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{use-haurwitz-as-reference-for-clear-sky.}{%
\subsection{3. Use Haurwitz as reference for Clear sky.}\label{use-haurwitz-as-reference-for-clear-sky.}}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# \_\_ 3. Vamvakas2020  Criteria  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\DocumentationTok{\#\# +5\% from model =\textgreater{} enhancements above 15 Wm\^{}2 the instrument uncertainty}
\NormalTok{C3\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.05}
\NormalTok{DATA[, Enhanc\_C\_3 }\SpecialCharTok{:=} \ConstantTok{FALSE}\NormalTok{]}
\NormalTok{DATA[, Enhanc\_C\_3\_ref }\SpecialCharTok{:=}\NormalTok{ CS\_ref }\SpecialCharTok{*}\NormalTok{ C3\_cs\_ref\_ratio]}
\NormalTok{DATA[wattGLB }\SpecialCharTok{\textgreater{}}\NormalTok{ Enhanc\_C\_3\_ref,}
\NormalTok{     Enhanc\_C\_3 }\SpecialCharTok{:=} \ConstantTok{TRUE}\NormalTok{]}

\ControlFlowTok{if}\NormalTok{ (SelEnhanc }\SpecialCharTok{==} \StringTok{"Enhanc\_C\_3"}\NormalTok{) \{}
\NormalTok{    DATA[ , GLB\_diff }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_3\_ref                    ] }\DocumentationTok{\#\# enhancement}
\NormalTok{    DATA[ , GLB\_ench }\SpecialCharTok{:=}\NormalTok{ ( wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_3\_ref ) }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_3\_ref ] }\DocumentationTok{\#\# relative enhancement}
\NormalTok{    DATA[ , GLB\_rati }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_3\_ref                    ]}
\NormalTok{\}}

\DocumentationTok{\#\# \_\_ 4. my Criteria  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}

\DocumentationTok{\#\# set values base}
\NormalTok{csmodel }\OtherTok{\textless{}{-}} \StringTok{"High\_B.Low\_W"}
\end{Highlighting}
\end{Shaded}

\hypertarget{use-libradtran-high_b.low_w-as-reference-for-clear-sky.}{%
\subsection{\texorpdfstring{4. Use libradtran \textbf{High\_B.Low\_W} as reference for Clear sky.}{4. Use libradtran High\_B.Low\_W as reference for Clear sky.}}\label{use-libradtran-high_b.low_w-as-reference-for-clear-sky.}}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cat}\NormalTok{(}\StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{ USING CSMODE:"}\NormalTok{, csmodel, }\StringTok{"}\SpecialCharTok{\textbackslash{}n\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

 USING CSMODE: High_B.Low_W 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\ControlFlowTok{switch}\NormalTok{(csmodel,}
       \AttributeTok{Exact\_B.Exact\_W =}\NormalTok{ \{ C4\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.02}\NormalTok{; C4\_GLB\_diff\_THRES }\OtherTok{\textless{}{-}} \DecValTok{55}\NormalTok{; C4\_lowcut\_sza }\OtherTok{\textless{}{-}} \DecValTok{60}\NormalTok{; C4\_lowcut\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.12}\NormalTok{\},}
       \AttributeTok{Low\_2\_B.Low\_2\_W =}\NormalTok{ \{ C4\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.03}\NormalTok{; C4\_GLB\_diff\_THRES }\OtherTok{\textless{}{-}}  \DecValTok{5}\NormalTok{; C4\_lowcut\_sza }\OtherTok{\textless{}{-}} \DecValTok{60}\NormalTok{; C4\_lowcut\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.12}\NormalTok{\},}
       \AttributeTok{Low\_B.Exact\_W   =}\NormalTok{ \{ C4\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.04}\NormalTok{; C4\_GLB\_diff\_THRES }\OtherTok{\textless{}{-}} \DecValTok{20}\NormalTok{; C4\_lowcut\_sza }\OtherTok{\textless{}{-}} \DecValTok{60}\NormalTok{; C4\_lowcut\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.12}\NormalTok{\},}
       \AttributeTok{Low\_B.High\_W    =}\NormalTok{ \{ C4\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.05}\NormalTok{; C4\_GLB\_diff\_THRES }\OtherTok{\textless{}{-}} \DecValTok{20}\NormalTok{; C4\_lowcut\_sza }\OtherTok{\textless{}{-}} \DecValTok{60}\NormalTok{; C4\_lowcut\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.12}\NormalTok{\},}
       \AttributeTok{Low\_B.Low\_W     =}\NormalTok{ \{ C4\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.05}\NormalTok{; C4\_GLB\_diff\_THRES }\OtherTok{\textless{}{-}}  \DecValTok{0}\NormalTok{; C4\_lowcut\_sza }\OtherTok{\textless{}{-}} \DecValTok{60}\NormalTok{; C4\_lowcut\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.18}\NormalTok{\},}
       \AttributeTok{High\_B.Low\_W    =}\NormalTok{ \{ C4\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.05}\NormalTok{; C4\_GLB\_diff\_THRES }\OtherTok{\textless{}{-}}  \DecValTok{0}\NormalTok{; C4\_lowcut\_sza }\OtherTok{\textless{}{-}} \DecValTok{60}\NormalTok{; C4\_lowcut\_ratio }\OtherTok{\textless{}{-}} \FloatTok{1.18}\NormalTok{\}, }\DocumentationTok{\#\# same as the used criterio}
\NormalTok{                         \{ C4\_cs\_ref\_ratio }\OtherTok{\textless{}{-}} \DecValTok{1}\NormalTok{   ; C4\_GLB\_diff\_THRES }\OtherTok{\textless{}{-}}  \DecValTok{0}\NormalTok{; C4\_lowcut\_sza }\OtherTok{\textless{}{-}}  \DecValTok{0}\NormalTok{; C4\_lowcut\_ratio }\OtherTok{\textless{}{-}} \DecValTok{1}\NormalTok{   \})}
\DocumentationTok{\#\# init flag}
\NormalTok{DATA[, Enhanc\_C\_4 }\SpecialCharTok{:=} \ConstantTok{FALSE}\NormalTok{]}

\CommentTok{\# DATA[, max(SZA)]}

\NormalTok{smo }\OtherTok{\textless{}{-}} \FunctionTok{approxfun}\NormalTok{(}
    \AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\DecValTok{90} \SpecialCharTok{{-}}\NormalTok{ BIO\_ELEVA, C4\_lowcut\_sza  ),}
    \AttributeTok{y =} \FunctionTok{c}\NormalTok{(C4\_lowcut\_ratio,  C4\_cs\_ref\_ratio)}
\NormalTok{    )}

\FunctionTok{smo}\NormalTok{(}\DecValTok{80}\SpecialCharTok{:}\DecValTok{70}\NormalTok{) }\SpecialCharTok{*}\NormalTok{ (}\DecValTok{1}\SpecialCharTok{/}\FunctionTok{cosd}\NormalTok{(}\DecValTok{80}\SpecialCharTok{:}\DecValTok{70}\NormalTok{) }\SpecialCharTok{/} \FunctionTok{max}\NormalTok{(}\DecValTok{1}\SpecialCharTok{/}\FunctionTok{cosd}\NormalTok{(}\DecValTok{80}\SpecialCharTok{:}\DecValTok{70}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
 [1]        NA        NA 0.9855379 0.9053113 0.8366196 0.7771548 0.7251860 0.6793893
 [9] 0.6387354 0.6024114 0.5697672
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cat}\NormalTok{(}\StringTok{"C4 factor:"}\NormalTok{, C4\_cs\_ref\_ratio,   }\StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
C4 factor: 1.05 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cat}\NormalTok{(}\StringTok{"C4 offset:"}\NormalTok{, C4\_GLB\_diff\_THRES, }\StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
C4 offset: 0 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# \_\_\_\_ Create global irradiance W/m\^{}2  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\NormalTok{DATA[, }\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{) }\SpecialCharTok{:=}\NormalTok{ (}\FunctionTok{get}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".edir"}\NormalTok{)) }\SpecialCharTok{+} \FunctionTok{get}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".edn"}\NormalTok{))) }\SpecialCharTok{/} \DecValTok{1000}\NormalTok{ ]}

\DocumentationTok{\#\# \_\_\_\_ Apply sun{-}earth distance correction  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\NormalTok{DATA[, }\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{) }\SpecialCharTok{:=} \FunctionTok{get}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{)) }\SpecialCharTok{/}\NormalTok{ sun\_dist}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{ ]}

\DocumentationTok{\#\# \_\_\_\_ Apply adjustment to Kurudz spectrum  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\NormalTok{DATA[, }\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{) }\SpecialCharTok{:=} \FunctionTok{get}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{)) }\SpecialCharTok{*}\NormalTok{ TSI\_Kurudz\_factor ]}

\DocumentationTok{\#\# \_\_\_\_ Calculate clearness index  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\NormalTok{DATA[, ClearnessIndex\_C\_4 }\SpecialCharTok{:=}\NormalTok{ wattGLB }\SpecialCharTok{/} \FunctionTok{get}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{)) ]}

\FunctionTok{hist}\NormalTok{(DATA}\SpecialCharTok{$}\NormalTok{ClearnessIndex\_C\_4, }\AttributeTok{breaks =} \DecValTok{100}\NormalTok{)}
\FunctionTok{abline}\NormalTok{(}\AttributeTok{v =}\NormalTok{ C4\_cs\_ref\_ratio, }\AttributeTok{col =} \StringTok{"red"}\NormalTok{ )}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-10-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# \_\_\_\_ Calculate reference and mark data  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
\DocumentationTok{\#\# for most of the data}
\NormalTok{DATA[SZA }\SpecialCharTok{\textless{}}\NormalTok{ C4\_lowcut\_sza, Enhanc\_C\_4\_ref }\SpecialCharTok{:=}\NormalTok{ (}\FunctionTok{get}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{)) }\SpecialCharTok{*}\NormalTok{ C4\_cs\_ref\_ratio) }\SpecialCharTok{+}\NormalTok{ C4\_GLB\_diff\_THRES ]}
\DocumentationTok{\#\# fol low sun angles}
\CommentTok{\# DATA[SZA \textgreater{} C4\_lowcut\_sza, Enhanc\_C\_4\_ref := (get(paste0(csmodel,".glo")) * C4\_lowcut\_ratio) ]}
\NormalTok{DATA[SZA }\SpecialCharTok{\textgreater{}}\NormalTok{ C4\_lowcut\_sza, Enhanc\_C\_4\_ref }\SpecialCharTok{:=}\NormalTok{ (}\FunctionTok{get}\NormalTok{(}\FunctionTok{paste0}\NormalTok{(csmodel,}\StringTok{".glo"}\NormalTok{)) }\SpecialCharTok{*} \FunctionTok{smo}\NormalTok{(SZA)) ]}

\NormalTok{DATA[wattGLB }\SpecialCharTok{\textgreater{}}\NormalTok{ Enhanc\_C\_4\_ref ,}
\NormalTok{     Enhanc\_C\_4 }\SpecialCharTok{:=} \ConstantTok{TRUE}\NormalTok{]}
\DocumentationTok{\#\# use threshold to compute values}
\ControlFlowTok{if}\NormalTok{ (SelEnhanc }\SpecialCharTok{==} \StringTok{"Enhanc\_C\_4"}\NormalTok{) \{}
\NormalTok{    DATA[ , GLB\_diff }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_4\_ref                    ] }\DocumentationTok{\#\# enhancement}
\NormalTok{    DATA[ , GLB\_ench }\SpecialCharTok{:=}\NormalTok{ ( wattGLB }\SpecialCharTok{{-}}\NormalTok{ Enhanc\_C\_4\_ref ) }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_4\_ref ] }\DocumentationTok{\#\# relative enhancement}
\NormalTok{    DATA[ , GLB\_rati }\SpecialCharTok{:=}\NormalTok{   wattGLB }\SpecialCharTok{/}\NormalTok{ Enhanc\_C\_4\_ref                    ]}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\FloatBarrier

\hypertarget{using-creteria-enhanc_c_4}{%
\section{Using creteria: Enhanc\_C\_4}\label{using-creteria-enhanc_c_4}}

\hypertarget{distribution-of-different-metrics}{%
\section{Distribution of different metrics}\label{distribution-of-different-metrics}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-3} \end{center}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}@{}}
\caption{Enhanc\_C\_1}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
6082440 & 119727 \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}@{}}
\caption{Enhanc\_C\_2}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
6012875 & 189292 \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}@{}}
\caption{Enhanc\_C\_3}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
5455724 & 746443 \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1389}}@{}}
\caption{Enhanc\_C\_4}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
FALSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
TRUE
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
5785494 & 416673 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-20} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-24} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-31} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-32} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/unnamed-chunk-13-33} \end{center}

\FloatBarrier

\hypertarget{plot-some-days-with-strong-enhancement-cases}{%
\section{Plot some days with strong enhancement cases}\label{plot-some-days-with-strong-enhancement-cases}}

\FloatBarrier

\hypertarget{days-with-extreme-cases-day}{%
\subsection{Days with extreme cases day}\label{days-with-extreme-cases-day}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-20} \end{center}

\FloatBarrier

\hypertarget{days-with-strong-enhancement-day}{%
\subsection{Days with strong enhancement day}\label{days-with-strong-enhancement-day}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-24} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-31} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-32} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-33} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-34} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-35} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-36} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-37} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-38} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-39} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-40} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-41} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-42} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-43} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-44} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-45} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-46} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-47} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-48} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-49} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-50} \end{center}

\FloatBarrier

\hypertarget{days-with-sunny-day}{%
\subsection{Days with sunny day}\label{days-with-sunny-day}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-51} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-52} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-53} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-54} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-55} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-56} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-57} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-58} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-59} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-60} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-61} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-62} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-63} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-64} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-65} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-66} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-67} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-68} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-69} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-70} \end{center}

\FloatBarrier

\hypertarget{days-with-sunny-enhancement-day}{%
\subsection{Days with sunny enhancement day}\label{days-with-sunny-enhancement-day}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-71} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-72} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-73} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-74} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-75} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-76} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-77} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-78} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-79} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-80} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-81} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-82} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-83} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-84} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-85} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-86} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-87} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-88} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-89} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-90} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-91} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-92} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-93} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-94} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-95} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-96} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-97} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-98} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-99} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-100} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-101} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-102} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-103} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-104} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-105} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-106} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-107} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-108} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-109} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-110} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-111} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-112} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-113} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-114} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-115} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-116} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-117} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-118} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-119} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-120} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-121} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-122} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-123} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-124} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-125} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-126} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-127} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-128} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-129} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-130} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-131} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-132} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-133} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-134} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-135} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-136} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-137} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-138} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-139} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-140} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-141} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-142} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-143} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-144} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-145} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-146} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-147} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-148} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-149} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-150} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-151} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-152} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-153} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-154} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-155} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-156} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-157} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-158} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-159} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-160} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-161} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-162} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-163} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-164} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-165} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-166} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-167} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-168} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-169} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-170} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-171} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-172} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-173} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-174} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-175} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-176} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-177} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-178} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-179} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-180} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-181} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-182} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-183} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-184} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-185} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-186} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-187} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-188} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-189} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-190} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-191} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-192} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-193} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-194} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-195} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-196} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-197} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-198} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-199} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-200} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-201} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-202} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-203} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-204} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-205} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-206} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-207} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-208} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-209} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-210} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-211} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-212} \end{center}

\FloatBarrier

\hypertarget{days-with-cloudy-day}{%
\subsection{Days with cloudy day}\label{days-with-cloudy-day}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-213} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-214} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-215} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-216} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-217} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-218} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-219} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-220} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-221} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-222} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-223} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-224} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-225} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-226} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-227} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-228} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-229} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-230} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-231} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-232} \end{center}

\FloatBarrier

\hypertarget{days-with-random-day}{%
\subsection{Days with random day}\label{days-with-random-day}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-233} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-234} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-235} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-236} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-237} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-238} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-239} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-240} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-241} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-242} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-243} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-244} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-245} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-246} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-247} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-248} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-249} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-250} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-251} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-252} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-253} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-254} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-255} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-256} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-257} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-258} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-259} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-260} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-261} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-262} \end{center}

\FloatBarrier

\hypertarget{days-with-manual-test-days}{%
\subsection{Days with manual test days}\label{days-with-manual-test-days}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-263} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-264} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-265} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_days-266} \end{center}

\newpage
\FloatBarrier

\hypertarget{plot-years-with-enhancement-cases}{%
\section{Plot years with enhancement cases}\label{plot-years-with-enhancement-cases}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-20} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-24} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_02_ID_CE_files/figure-latex/example_years-31} \end{center}

\textbf{END}

\begin{verbatim}
2024-04-09 19:11:13.0 athan@tyler GHI_enh_02_ID_CE.R 62.482203 mins
\end{verbatim}

\end{document}