### 2024-04-13 high_aod/GHI_enh_04_investigate.tex
Commit 2024-04-13 14:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-04-13}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}

Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:data.table':

    yearmon, yearqtr
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}
Registered S3 methods overwritten by 'ggpp':
  method                  from   
  heightDetails.titleGrob ggplot2
  widthDetails.titleGrob  ggplot2
\end{verbatim}

\begin{verbatim}

Attaching package: 'ggpp'
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggh4x':

    stat_centroid
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggplot2':

    annotate
\end{verbatim}

\begin{verbatim}
Registered S3 method overwritten by 'ggpmisc':
  method                  from   
  as.character.polynomial polynom
\end{verbatim}

\begin{verbatim}
ST_daily ST_E_daily ST_E_daily_seas ST_extreme_daily
\end{verbatim}

\newpage
\FloatBarrier

\hypertarget{daily-trends}{%
\subsubsection{Daily Trends}\label{daily-trends}}

\begin{verbatim}
Warning in ans * length(l) + if1: longer object length is not a multiple of
shorter object length
\end{verbatim}

\newpage

\begin{verbatim}
Loading required package: stringr
\end{verbatim}

\hypertarget{trends-on-daily-observations-data}{%
\paragraph{Trends on Daily Observations data}\label{trends-on-daily-observations-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 1.517841 +/- 0.099133

lmY: 1.517857 +/- 0.099134

Arima: 1.518153 +/- 0.131133

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-1} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.986143 +/- 0.155127

lmY: 0.986168 +/- 0.155129

Arima: 1.050369 +/- 0.236468

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-2} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 1.152913 +/- 0.165733

lmY: 1.152941 +/- 0.165736

Arima: 1.217146 +/- 0.245786

Units for GLB\_diff.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
\end{verbatim}

\begin{verbatim}
Warning in sqrt(diag(se)): NaNs produced
Warning in sqrt(diag(se)): NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-3} \end{center}

Arima: -0.013183 +/- 6.904159

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-4} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: 1.06442 +/- 0.24189

lmY: 1.06443 +/- 0.241895

Arima: 1.064477 +/- 0.369173

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-5} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: 0.965697 +/- 0.257186

lmY: 0.965706 +/- 0.25719

Arima: 0.965757 +/- 0.397302

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-6} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 1.086689 +/- 0.074568

lmY: 1.086704 +/- 0.07457

Arima: 1.086709 +/- 0.111711

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-7} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 1.121477 +/- 0.069651

lmY: 1.121491 +/- 0.069652

Arima: 1.1215 +/- 0.100263

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-8} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 34.75674 +/- 5.785425

lmY: 34.75733 +/- 5.785525

Arima: 34.809263 +/- 9.232706

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-9} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.00359 +/- 0.000275

lmY: 0.00359 +/- 0.000275

Arima: 0.003611 +/- 0.000363

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-10} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.001821 +/- 0.000281

lmY: 0.001821 +/- 0.000281

Arima: 0.001976 +/- 0.000461

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-11} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.001769 +/- 0.000314

lmY: 0.00177 +/- 0.000314

Arima: 0.001923 +/- 0.000501

Units for GLB\_ench.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-12} \end{center}

Arima: -0.013183 +/- 6.904159

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-13} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 0.001905 +/- 0.000317

lmY: 0.001905 +/- 0.000317

Arima: 0.001905 +/- 0.000497

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-14} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 0.001871 +/- 0.000311

lmY: 0.001871 +/- 0.000311

Arima: 0.001871 +/- 0.00048

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-15} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.002627 +/- 0.000234

lmY: 0.002627 +/- 0.000234

Arima: 0.002627 +/- 0.000357

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-16} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.002857 +/- 0.00023

lmY: 0.002857 +/- 0.00023

Arima: 0.002857 +/- 0.000337

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-17} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 1.027493 +/- 0.146835

lmY: 1.027517 +/- 0.146837

Arima: 1.088887 +/- 0.223551

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-18} \end{center}

\newpage

\hypertarget{trends-on-daily-enhancements-data}{%
\paragraph{Trends on Daily Enhancements data}\label{trends-on-daily-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.821483 +/- 0.093059

lmY: 0.821496 +/- 0.093061

Arima: 0.821491 +/- 0.106018

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-19} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.231635 +/- 0.035413

lmY: 0.231639 +/- 0.035414

Arima: 0.231639 +/- 0.038541

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-20} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 0.185627 +/- 0.035397

lmY: 0.18563 +/- 0.035398

Arima: 0.185631 +/- 0.037388

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-21} \end{center}

Units for GLB\_diff.N :

lmD: 0.484056 +/- 0.045622

lmY: 0.484065 +/- 0.045623

Arima: 0.484067 +/- 0.052795

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-22} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: 0.004522 +/- 0.01809

lmY: 0.004522 +/- 0.018091

Arima: 0.004522 +/- 0.018546

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-23} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -0.01638 +/- 0.016701

lmY: -0.01638 +/- 0.016702

Arima: -0.01638 +/- 0.016967

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-24} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.513947 +/- 0.066925

lmY: 0.513956 +/- 0.066926

Arima: 0.513955 +/- 0.073889

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-25} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.62079 +/- 0.075699

lmY: 0.620801 +/- 0.075701

Arima: 0.620798 +/- 0.084973

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-26} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 1.376585 +/- 0.138013

lmY: 1.376613 +/- 0.138016

Arima: 1.376617 +/- 0.159843

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-27} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.00152 +/- 0.000147

lmY: 0.00152 +/- 0.000147

Arima: 0.00152 +/- 0.000172

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-28} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.00047 +/- 6.5e-05

lmY: 0.00047 +/- 6.5e-05

Arima: 0.00047 +/- 7.4e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-29} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.000371 +/- 6.8e-05

lmY: 0.000371 +/- 6.8e-05

Arima: 0.000371 +/- 7.5e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-30} \end{center}

Units for GLB\_ench.N :

lmD: 0.484056 +/- 0.045622

lmY: 0.484065 +/- 0.045623

Arima: 0.484067 +/- 0.052795

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-31} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 2.4e-05 +/- 3.2e-05

lmY: 2.4e-05 +/- 3.2e-05

Arima: 2.4e-05 +/- 3.6e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-32} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: -9e-06 +/- 2.8e-05

lmY: -9e-06 +/- 2.8e-05

Arima: -9e-06 +/- 3.3e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-33} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.001043 +/- 0.000117

lmY: 0.001043 +/- 0.000117

Arima: 0.001043 +/- 0.000134

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-34} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.00124 +/- 0.000128

lmY: 0.00124 +/- 0.000128

Arima: 0.00124 +/- 0.000149

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-35} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.040357 +/- 0.004191

lmY: 0.040358 +/- 0.004191

Arima: 0.040358 +/- 0.004746

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-36} \end{center}

\newpage

\hypertarget{trends-on-daily-extreme-enhancements-data}{%
\paragraph{Trends on Daily Extreme Enhancements data}\label{trends-on-daily-extreme-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.194018 +/- 0.214287

lmY: 0.194046 +/- 0.21429

Arima: 0.186379 +/- 0.278973

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-37} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.05596 +/- 0.153614

lmY: 0.055976 +/- 0.153616

Arima: 0.03062 +/- 0.246509

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-38} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 0.096383 +/- 0.153736

lmY: 0.096398 +/- 0.153738

Arima: 0.071517 +/- 0.241036

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-39} \end{center}

Units for GLB\_diff.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-40} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: -0.06688 +/- 0.139832

lmY: -0.066875 +/- 0.139835

Arima: -0.104892 +/- 0.240811

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-41} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -0.088196 +/- 0.139661

lmY: -0.088192 +/- 0.139663

Arima: -0.125787 +/- 0.23978

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-42} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.176855 +/- 0.193744

lmY: 0.176884 +/- 0.193747

Arima: 0.166242 +/- 0.263579

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-43} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.185582 +/- 0.202704

lmY: 0.18561 +/- 0.202708

Arima: 0.17647 +/- 0.269872

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-44} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 0.354016 +/- 0.243235

lmY: 0.35404 +/- 0.243239

Arima: 0.354135 +/- 0.247045

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-45} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.000568 +/- 0.000406

lmY: 0.000568 +/- 0.000406

Arima: 0.000566 +/- 0.000539

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-46} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.000409 +/- 0.00035

lmY: 0.000409 +/- 0.00035

Arima: 4e-04 +/- 0.000503

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-47} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.000433 +/- 0.000359

lmY: 0.000433 +/- 0.000359

Arima: 0.000424 +/- 0.000507

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-48} \end{center}

Units for GLB\_ench.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-49} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 0.000213 +/- 0.000334

lmY: 0.000213 +/- 0.000334

Arima: 0.000198 +/- 0.000495

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-50} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 2e-04 +/- 0.000333

lmY: 2e-04 +/- 0.000333

Arima: 0.000184 +/- 0.000494

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-51} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.000567 +/- 0.000389

lmY: 0.000567 +/- 0.000389

Arima: 0.000564 +/- 0.000526

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-52} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.000568 +/- 0.000396

lmY: 0.000568 +/- 0.000397

Arima: 0.000565 +/- 0.000532

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-53} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.008854 +/- 0.005655

lmY: 0.008855 +/- 0.005655

Arima: 0.008895 +/- 0.005992

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-54} \end{center}

Arima: 0.231639 +/- 0.038541

\begin{verbatim}
Warning: The S3 guide system was deprecated in ggplot2 3.5.0.
i It has been replaced by a ggproto system that can be extended.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_daily_trend-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{group-stats}{%
\subsubsection{Group stats}\label{group-stats}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-1} \end{center}

\begin{verbatim}
Warning in scale_y_log10(): log-10 transformation introduced infinite values.
\end{verbatim}

\begin{verbatim}
Warning: Removed 15 rows containing missing values or values outside the scale
range (`geom_bar()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-2} \end{center}

\begin{verbatim}
Warning: A numeric `legend.position` argument in `theme()` was deprecated in ggplot2 3.5.0.
i Please use the `legend.position.inside` argument of `theme()` instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-6} \end{center}

\hypertarget{zhang2018}{%
\subsection{@Zhang2018}\label{zhang2018}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-8} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 36 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{verbatim}
Warning: Removed 3 rows containing missing values or values outside the scale
range (`geom_tile()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-1} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 36 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{verbatim}
Warning: Removed 4 rows containing missing values or values outside the scale
range (`geom_tile()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-2} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 36 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{verbatim}
Warning: Removed 3 rows containing missing values or values outside the scale
range (`geom_tile()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-3} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 36 rows containing non-finite outside the scale range (`stat_bin2d()`).
Removed 3 rows containing missing values or values outside the scale range (`geom_tile()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_N_doy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-1} \end{center}

\begin{verbatim}
Warning: The `fun.y` argument of `stat_summary()` is deprecated as of ggplot2 3.3.0.
i Please use the `fun` argument instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climECEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_daily-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_week-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{energy-contribution-of-enhancements}{%
\subsubsection{Energy contribution of enhancements}\label{energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-8} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-9} \end{center}

lmY: 346.8852 +/- 33.74613

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-1} \end{center}

lmY: 133.6263 +/- 11.25784

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-2} \end{center}

lmY: 0.232888 +/- 0.050234

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-3} \end{center}

lmY: 0.202917 +/- 0.037683

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_multi-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-3} \end{center}

\newpage
\FloatBarrier

\hypertarget{monthly-energy-contribution-of-enhancements}{%
\subsubsection{Monthly Energy contribution of enhancements}\label{monthly-energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-6} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-7} \end{center}

\textbf{END}

\begin{verbatim}
2024-04-13 09:28:09.9 athan@tyler GHI_enh_04_investigate.R 3.383829 mins
\end{verbatim}

\end{document}