### 2024-04-26 GHI_enh_Aeronet.tex
Commit 2024-04-27 02:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-04-26}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
Cross over point: 2014.75 
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-5} \end{center}

\begin{verbatim}
[1] 1997
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-9} \end{center}

\begin{verbatim}
[1] 2005.667
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-13} \end{center}

SZA min : 0.00133216113167738

SZA mean : 0.00192971268218197

SZA median : 0.00178690927379885

\newpage

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0972}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
year
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
typer
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1997 & BR SZA median & 449.5 \\
2005 & SZA median & 392.1 \\
2005 & BR SZA median & 459.3 \\
2006 & SZA median & 493.1 \\
2007 & SZA median & 470.8 \\
2008 & SZA median & 326.2 \\
2011 & SZA median & 505.1 \\
2012 & SZA median & 500.8 \\
2013 & SZA median & 498.6 \\
2014 & SZA median & 477.7 \\
2015 & SZA median & 504.8 \\
2016 & SZA median & 510.9 \\
2017 & SZA median & 551.4 \\
2018 & SZA median & 499.7 \\
2019 & SZA median & 476 \\
2020 & SZA median & 483.7 \\
2021 & SZA median & 506 \\
2022 & SZA median & 554.9 \\
2023 & SZA median & 522.8 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-1} \end{center}

\newpage

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0972}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
year
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
typer
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1997 & BR SZA min & 1055 \\
2005 & SZA min & 668.3 \\
2005 & BR SZA min & 1064 \\
2006 & SZA min & 817.1 \\
2007 & SZA min & 785.1 \\
2008 & SZA min & 557.2 \\
2011 & SZA min & 831.1 \\
2012 & SZA min & 824.3 \\
2013 & SZA min & 814 \\
2014 & SZA min & 822.1 \\
2015 & SZA min & 828.6 \\
2016 & SZA min & 832.3 \\
2017 & SZA min & 899.3 \\
2018 & SZA min & 822.3 \\
2019 & SZA min & 789.7 \\
2020 & SZA min & 820.3 \\
2021 & SZA min & 829.7 \\
2022 & SZA min & 902.4 \\
2023 & SZA min & 859.5 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-2} \end{center}

\newpage

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0972}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1944}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
year
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
typer
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1997 & BR SZA mean & 497.3 \\
2005 & SZA mean & 353.9 \\
2005 & BR SZA mean & 507.1 \\
2006 & SZA mean & 460.7 \\
2007 & SZA mean & 434.8 \\
2008 & SZA mean & 288.6 \\
2011 & SZA mean & 476.1 \\
2012 & SZA mean & 468.4 \\
2013 & SZA mean & 468.5 \\
2014 & SZA mean & 444.8 \\
2015 & SZA mean & 472.3 \\
2016 & SZA mean & 480.3 \\
2017 & SZA mean & 520.9 \\
2018 & SZA mean & 467.3 \\
2019 & SZA mean & 440.6 \\
2020 & SZA mean & 450.8 \\
2021 & SZA mean & 473.5 \\
2022 & SZA mean & 525.5 \\
2023 & SZA mean & 490.7 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-3} \end{center}

null device
1

\textbf{END}

\end{document}