### 2024-04-27 GHI_enh_CSid.tex
Commit 2024-04-27 20:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-04-27}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\texttt{Low\_B.Low\_W.glo} has compensation for sun distance, TSI variation and Kurudz difference

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# keep only clear}
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[TYPE }\SpecialCharTok{==} \StringTok{"Clear"}\NormalTok{]}

\FunctionTok{table}\NormalTok{(DATA}\SpecialCharTok{$}\NormalTok{TYPE)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

  Clear 
2577597 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wattGLB)]}

\FunctionTok{hist}\NormalTok{(DATA}\SpecialCharTok{$}\NormalTok{ClearnessIndex\_C\_4)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# Hard limits}
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[ClearnessIndex\_C\_4 }\SpecialCharTok{\textgreater{}} \FloatTok{0.8}\NormalTok{ ]}
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[ClearnessIndex\_C\_4 }\SpecialCharTok{\textless{}} \FloatTok{1.3}\NormalTok{ ]}

\NormalTok{DTdaily }\OtherTok{\textless{}{-}}\NormalTok{ DATA[,}
\NormalTok{                .(}
                  \AttributeTok{glo\_Sum   =} \FunctionTok{sum}\NormalTok{(Low\_B.Low\_W.glo),}
                  \AttributeTok{GLB\_Sum   =} \FunctionTok{sum}\NormalTok{(wattGLB),}
                  \AttributeTok{GLB\_N     =} \FunctionTok{sum}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wattGLB)),}
                  \AttributeTok{DayLength =} \FunctionTok{max}\NormalTok{(DayLength),}
\NormalTok{                  .N}
\NormalTok{                ),}
\NormalTok{                by }\OtherTok{=}\NormalTok{ Day]}

\NormalTok{Day\_points }\OtherTok{\textless{}{-}} \DecValTok{30}

\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[ Day }\SpecialCharTok{\%in\%}\NormalTok{ DTdaily[GLB\_N }\SpecialCharTok{\textgreater{}}\NormalTok{ Day\_points, Day]]}

\FunctionTok{hist}\NormalTok{(DTdaily}\SpecialCharTok{$}\NormalTok{GLB\_N)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-2} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{hist}\NormalTok{(DTdaily[, GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength])}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-3} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# plot(DTdaily[, GLB\_N/DayLength, Day])}
\FunctionTok{summary}\NormalTok{(DTdaily[, GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
0.001152 0.181870 0.422018 0.420691 0.657944 0.838020 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# remove days with few relative data}
\NormalTok{Daytime\_Ratio }\OtherTok{\textless{}{-}} \FloatTok{0.3}

\NormalTok{DTdaily[GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength }\SpecialCharTok{\textless{}}\NormalTok{ Daytime\_Ratio, .N]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 2932
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[Day }\SpecialCharTok{\%in\%}\NormalTok{ DTdaily[GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength }\SpecialCharTok{\textgreater{}}\NormalTok{ Daytime\_Ratio, Day]]}

\NormalTok{DTdaily }\OtherTok{\textless{}{-}}\NormalTok{ DATA[,}
\NormalTok{                .(}
                  \AttributeTok{glo\_Sum   =} \FunctionTok{sum}\NormalTok{(Low\_B.Low\_W.glo),}
                  \AttributeTok{GLB\_Sum   =} \FunctionTok{sum}\NormalTok{(wattGLB),}
                  \AttributeTok{GLB\_N     =} \FunctionTok{sum}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wattGLB)),}
                  \AttributeTok{DayLength =} \FunctionTok{max}\NormalTok{(DayLength),}
\NormalTok{                  .N}
\NormalTok{                ),}
\NormalTok{                by }\OtherTok{=}\NormalTok{ Day]}

\CommentTok{\# plot(DATA[, ClearnessIndex\_C\_4, Date ])}

\CommentTok{\# table(DATA$BAD\_h)}

\NormalTok{DTyear }\OtherTok{\textless{}{-}}\NormalTok{ DATA[, .(}\AttributeTok{glo =} \FunctionTok{mean}\NormalTok{(Low\_B.Low\_W.glo),}
                   \AttributeTok{GLB =} \FunctionTok{mean}\NormalTok{(wattGLB),}
\NormalTok{                   .N),}
\NormalTok{               by }\OtherTok{=}\NormalTok{ .(}\FunctionTok{year}\NormalTok{(Date))]}

\NormalTok{ylim }\OtherTok{\textless{}{-}} \FunctionTok{range}\NormalTok{(DTyear[, glo, GLB])}

\CommentTok{\# plot(  DTyear[, glo, year],}
\CommentTok{\#        ylim = ylim)}
\CommentTok{\# points(DTyear[, GLB, year], col = "green")}

\FunctionTok{plot}\NormalTok{(DTyear[, N, year], }\AttributeTok{col =} \StringTok{"blue"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-4} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(DTyear[, GLB}\SpecialCharTok{/}\NormalTok{glo, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{title}\NormalTok{(}\StringTok{"Clear GLB / CS libratran by year"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-5} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DTmonth }\OtherTok{\textless{}{-}}\NormalTok{ DATA[, .(}\AttributeTok{glo =} \FunctionTok{mean}\NormalTok{(Low\_B.Low\_W.glo),}
                    \AttributeTok{GLB =} \FunctionTok{mean}\NormalTok{(wattGLB),}
\NormalTok{                    .N),}
\NormalTok{                by }\OtherTok{=}\NormalTok{ .(}\FunctionTok{year}\NormalTok{(Date), }\FunctionTok{month}\NormalTok{(Date))]}
\NormalTok{DTmonth[, tsy }\SpecialCharTok{:=}\NormalTok{ year }\SpecialCharTok{+}\NormalTok{ (month }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{)}\SpecialCharTok{/}\DecValTok{12}\NormalTok{ ]}

\NormalTok{ylim }\OtherTok{\textless{}{-}} \FunctionTok{range}\NormalTok{(DTmonth[, glo, GLB])}

\CommentTok{\# plot(  DTmonth[, glo, tsy],}
\CommentTok{\#        ylim = ylim)}
\CommentTok{\# points(DTmonth[, GLB, tsy], col = "green")}

\CommentTok{\# plot(  DTmonth[, .N, tsy], col = "blue")}
\FunctionTok{plot}\NormalTok{(  DTmonth[, GLB}\SpecialCharTok{/}\NormalTok{glo, tsy], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{title}\NormalTok{(}\StringTok{"Clear GLB / CS libratran by month"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-6} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\SpecialCharTok{::}\NormalTok{opts\_chunk}\SpecialCharTok{$}\FunctionTok{set}\NormalTok{(}\AttributeTok{out.height =} \StringTok{"32\%"}\NormalTok{   )}

\NormalTok{Ratio\_lim }\OtherTok{\textless{}{-}} \FloatTok{0.85}

\NormalTok{DTdaily[GLB\_Sum}\SpecialCharTok{/}\NormalTok{glo\_Sum }\SpecialCharTok{\textless{}}\NormalTok{ Ratio\_lim, .N]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 134
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{days }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(}\FunctionTok{c}\NormalTok{(DTdaily[GLB\_Sum}\SpecialCharTok{/}\NormalTok{glo\_Sum }\SpecialCharTok{\textless{}}\NormalTok{ Ratio\_lim, Day]))}

\NormalTok{days }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(DTdaily[, Day], }\DecValTok{40}\NormalTok{)}

\ControlFlowTok{for}\NormalTok{ (ad }\ControlFlowTok{in} \FunctionTok{sort}\NormalTok{(days)) \{}
\NormalTok{  temp }\OtherTok{\textless{}{-}}\NormalTok{ DATA[Day }\SpecialCharTok{==}\NormalTok{ ad,]}

\NormalTok{  ylim }\OtherTok{\textless{}{-}} \FunctionTok{range}\NormalTok{(temp[, Low\_B.Low\_W.glo, wattGLB])}

  \FunctionTok{plot}\NormalTok{(  temp[, Low\_B.Low\_W.glo, Date],}
         \AttributeTok{col =} \StringTok{"blue"}\NormalTok{,}
         \AttributeTok{ylim =}\NormalTok{ ylim)}
  \FunctionTok{points}\NormalTok{(temp[, wattGLB, Date], }\AttributeTok{col =} \StringTok{"green"}\NormalTok{)}
  \FunctionTok{title}\NormalTok{(}\FunctionTok{as.Date}\NormalTok{(ad, }\AttributeTok{origin =} \StringTok{"1970{-}01{-}01"}\NormalTok{))}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-20} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-24} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-31} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-32} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-33} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-34} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-35} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-36} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-37} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-38} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-39} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-40} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-41} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-42} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-43} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-44} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-45} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-46} \end{center}

\textbf{END}

\end{document}