### 2024-05-08 GHI_enh_CSid.tex
Commit 2024-05-08 22:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-05-08}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\texttt{Low\_B.Low\_W.glo} has compensation for sun distance, TSI variation and Kurudz difference only

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# keep only clear}
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[TYPE }\SpecialCharTok{==} \StringTok{"Clear"}\NormalTok{]}

\FunctionTok{table}\NormalTok{(DATA}\SpecialCharTok{$}\NormalTok{TYPE)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

  Clear 
2577597 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wattGLB)]}

\FunctionTok{hist}\NormalTok{(DATA}\SpecialCharTok{$}\NormalTok{ClearnessIndex\_C\_4)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-1} \end{center}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# Hard limits}
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[ClearnessIndex\_C\_4 }\SpecialCharTok{\textgreater{}} \FloatTok{0.8}\NormalTok{ ]}
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[ClearnessIndex\_C\_4 }\SpecialCharTok{\textless{}} \FloatTok{1.3}\NormalTok{ ]}

\FunctionTok{source}\NormalTok{(}\StringTok{"./GHI\_enh\_Aeronet.R"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Cross over point: 2005 
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-20} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-24} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-26} \end{center}

\begin{verbatim}
SZA min : 0.00137165912186399 
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-31} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-32} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-33} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-34} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-35} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-36} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-37} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-38} \end{center}

\begin{verbatim}
SZA mean : 0.0020024477410843 
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-39} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-40} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-41} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-42} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-43} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-44} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-45} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-46} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-47} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-48} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-49} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-50} \end{center}

\begin{verbatim}
SZA median : 0.00184853525463023

\newpage
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-51} \end{center}

\begin{verbatim}

\newpage
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_CSid_files/figure-latex/unnamed-chunk-5-52} \end{center}

\begin{verbatim}

\newpage
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA[, tsy }\SpecialCharTok{:=} \FunctionTok{year}\NormalTok{(Date) }\SpecialCharTok{+}\NormalTok{ (}\FunctionTok{month}\NormalTok{(Date) }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{)}\SpecialCharTok{/}\DecValTok{12}\NormalTok{ ]}

\DocumentationTok{\#\# apply AOD trend}
\NormalTok{trendsf }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(trend\_median, trend\_mean, trend\_min)}

\NormalTok{trend\_apply  }\OtherTok{\textless{}{-}}\NormalTok{ trend\_median}
\NormalTok{trend\_apply  }\OtherTok{\textless{}{-}}\NormalTok{ trend\_mean}

\NormalTok{trend\_factor }\OtherTok{\textless{}{-}} \FloatTok{0.86}

\NormalTok{DATA[, Thresh\_test }\SpecialCharTok{:=}\NormalTok{ Enhanc\_C\_4\_ref }\SpecialCharTok{*}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{trend\_apply}\NormalTok{(tsy) }\SpecialCharTok{*}\NormalTok{ trend\_factor)]}

\FunctionTok{rmserr}\NormalTok{(DATA}\SpecialCharTok{$}\NormalTok{Enhanc\_C\_4\_ref, DATA}\SpecialCharTok{$}\NormalTok{wattGLB)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
$mae
[1] 94.57364

$mse
[1] 10035.43

$rmse
[1] 100.177

$mape
[1] 0.1485271

$nmse
[1] 0.1642926

$rstd
[1] 0.1495398
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{f }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x) \{}

\NormalTok{  test }\OtherTok{\textless{}{-}}\NormalTok{ DATA[, .(}
    \FunctionTok{mean}\NormalTok{(Enhanc\_C\_4\_ref }\SpecialCharTok{*}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+} \FunctionTok{trend\_apply}\NormalTok{(tsy) }\SpecialCharTok{*}\NormalTok{ x)),}
    \FunctionTok{mean}\NormalTok{(wattGLB)),}
\NormalTok{       by }\OtherTok{=}\NormalTok{ .(}\FunctionTok{year}\NormalTok{(Date), }\FunctionTok{month}\NormalTok{(Date))]}

  \CommentTok{\# DATA[, sum( abs((Enhanc\_C\_4\_ref * (1 + trend\_apply(tsy) * x)) {-} wattGLB)) ]}

  \CommentTok{\# DATA[, sum(1 {-} (Enhanc\_C\_4\_ref * (1 + trend\_apply(tsy) * x) / wattGLB)) ]}

  \FunctionTok{rmserr}\NormalTok{(test}\SpecialCharTok{$}\NormalTok{V1, test}\SpecialCharTok{$}\NormalTok{V2)}\SpecialCharTok{$}\NormalTok{mse}
  \CommentTok{\# test[, sum(V1 {-} V2)]}
\NormalTok{\}}

\FunctionTok{optimise}\NormalTok{(f, }\AttributeTok{interval =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{1.8}\NormalTok{), }\AttributeTok{maximum =}\NormalTok{ F, }\AttributeTok{tol =} \FloatTok{0.0001}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
$minimum
[1] 0.2000654

$objective
[1] 8644.06
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# optim(0.8, f, method = "Brent", lower = 0.2, upper = 1.5)}

\CommentTok{\# plot(DATA[, .(tsy, (1 + trend\_apply(tsy) * trend\_factor))])}

\NormalTok{DTdaily }\OtherTok{\textless{}{-}}\NormalTok{ DATA[,}
\NormalTok{                .(}
                  \AttributeTok{glo\_Sum    =} \FunctionTok{sum}\NormalTok{(Low\_B.Low\_W.glo),}
                  \AttributeTok{GLB\_Sum    =} \FunctionTok{sum}\NormalTok{(wattGLB),}
                  \AttributeTok{GLB\_N      =} \FunctionTok{sum}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wattGLB)),}
                  \AttributeTok{DayLength  =} \FunctionTok{max}\NormalTok{(DayLength),}
                  \AttributeTok{En\_Ref\_sum =} \FunctionTok{sum}\NormalTok{(Enhanc\_C\_4\_ref),}
                  \AttributeTok{Thresh\_sum =} \FunctionTok{sum}\NormalTok{(Thresh\_test),}
\NormalTok{                  .N}
\NormalTok{                ),}
\NormalTok{                by }\OtherTok{=}\NormalTok{ Day]}

\NormalTok{Day\_points }\OtherTok{\textless{}{-}} \DecValTok{30}

\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[ Day }\SpecialCharTok{\%in\%}\NormalTok{ DTdaily[GLB\_N }\SpecialCharTok{\textgreater{}}\NormalTok{ Day\_points, Day]]}

\FunctionTok{hist}\NormalTok{(DTdaily}\SpecialCharTok{$}\NormalTok{GLB\_N)}
\FunctionTok{hist}\NormalTok{(DTdaily[, GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength])}
\CommentTok{\# plot(DTdaily[, GLB\_N/DayLength, Day])}
\FunctionTok{summary}\NormalTok{(DTdaily[, GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength])}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
    Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
0.001152 0.181870 0.422018 0.420691 0.657944 0.838020 
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(DTdaily[, En\_Ref\_sum, Day])}
\FunctionTok{plot}\NormalTok{(DTdaily[, glo\_Sum, Day])}
\FunctionTok{plot}\NormalTok{(DTdaily[, GLB\_Sum, Day])}

\DocumentationTok{\#\# remove days with few relative data}
\NormalTok{Daytime\_Ratio }\OtherTok{\textless{}{-}} \FloatTok{0.3}

\NormalTok{DTdaily[GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength }\SpecialCharTok{\textless{}}\NormalTok{ Daytime\_Ratio, .N]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 2932
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\OtherTok{\textless{}{-}}\NormalTok{ DATA[Day }\SpecialCharTok{\%in\%}\NormalTok{ DTdaily[GLB\_N}\SpecialCharTok{/}\NormalTok{DayLength }\SpecialCharTok{\textgreater{}}\NormalTok{ Daytime\_Ratio, Day]]}

\NormalTok{DTdaily }\OtherTok{\textless{}{-}}\NormalTok{ DATA[,}
\NormalTok{                .(}
                  \AttributeTok{glo\_Sum    =} \FunctionTok{sum}\NormalTok{(Low\_B.Low\_W.glo),}
                  \AttributeTok{GLB\_Sum    =} \FunctionTok{sum}\NormalTok{(wattGLB),}
                  \AttributeTok{GLB\_N      =} \FunctionTok{sum}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(wattGLB)),}
                  \AttributeTok{DayLength  =} \FunctionTok{max}\NormalTok{(DayLength),}
                  \AttributeTok{En\_Ref\_sum =} \FunctionTok{sum}\NormalTok{(Enhanc\_C\_4\_ref),}
                  \AttributeTok{Thresh\_sum =} \FunctionTok{sum}\NormalTok{(Thresh\_test),}
\NormalTok{                  .N}
\NormalTok{                ),}
\NormalTok{                by }\OtherTok{=}\NormalTok{ Day]}

\CommentTok{\# plot(DATA[, ClearnessIndex\_C\_4, Date ])}

\CommentTok{\# table(DATA$BAD\_h)}

\NormalTok{DTyear }\OtherTok{\textless{}{-}}\NormalTok{ DATA[, .(}\AttributeTok{glo =} \FunctionTok{mean}\NormalTok{(Low\_B.Low\_W.glo),}
                   \AttributeTok{GLB =} \FunctionTok{mean}\NormalTok{(wattGLB),}
                   \AttributeTok{Ref =} \FunctionTok{mean}\NormalTok{(Enhanc\_C\_4\_ref),}
                   \AttributeTok{Thr =} \FunctionTok{mean}\NormalTok{(Thresh\_test),}
\NormalTok{                   .N),}
\NormalTok{               by }\OtherTok{=}\NormalTok{ .(}\FunctionTok{year}\NormalTok{(Date))]}

\NormalTok{ylim }\OtherTok{\textless{}{-}} \FunctionTok{range}\NormalTok{(DTyear[, glo, GLB])}

\CommentTok{\# plot(  DTyear[, glo, year],}
\CommentTok{\#        ylim = ylim)}
\CommentTok{\# points(DTyear[, GLB, year], col = "green")}

\FunctionTok{plot}\NormalTok{(DTyear[, N, year], }\AttributeTok{col =} \StringTok{"blue"}\NormalTok{)}

\FunctionTok{plot}\NormalTok{(DTyear[, GLB}\SpecialCharTok{/}\NormalTok{glo, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{title}\NormalTok{(}\StringTok{"Clear GLB / CS libratran yearly means"}\NormalTok{)}

\FunctionTok{plot}\NormalTok{(DTyear[, Ref}\SpecialCharTok{/}\NormalTok{GLB, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{title}\NormalTok{(}\StringTok{"Threshold ref / Clear Global yearly means"}\NormalTok{)}

\FunctionTok{plot}\NormalTok{(DTyear[, Ref}\SpecialCharTok{/}\NormalTok{glo, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{title}\NormalTok{(}\StringTok{"Threshold ref / CS libratran yearly means"}\NormalTok{)}

\FunctionTok{plot}\NormalTok{(DTyear[, Thr}\SpecialCharTok{/}\NormalTok{GLB, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{title}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"Threshold / global yearly means"}\NormalTok{, trend\_factor))}

\NormalTok{ff }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x) \{}
    \FunctionTok{plot}\NormalTok{(DTyear[, }\DecValTok{1} \SpecialCharTok{{-}}\NormalTok{ x }\SpecialCharTok{*}\NormalTok{ Thr}\SpecialCharTok{/}\NormalTok{GLB, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{, }\AttributeTok{main =}\NormalTok{ x)}
\NormalTok{    DTyear[, }\FunctionTok{sum}\NormalTok{(}\FunctionTok{abs}\NormalTok{(}\DecValTok{1} \SpecialCharTok{{-}}\NormalTok{ x }\SpecialCharTok{*}\NormalTok{ Thr}\SpecialCharTok{/}\NormalTok{GLB))]}
\NormalTok{\}}

\FunctionTok{optimize}\NormalTok{(ff, }\FunctionTok{c}\NormalTok{(.}\DecValTok{5}\NormalTok{, }\FloatTok{1.5}\NormalTok{), }\AttributeTok{tol =} \FloatTok{0.001}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
$minimum
[1] 0.8590668

$objective
[1] 0.6771877
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{ff}\NormalTok{(}\FloatTok{0.85}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 0.7037762
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{plot}\NormalTok{(DTyear[, GLB, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{plot}\NormalTok{(DTyear[, Thr, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{plot}\NormalTok{(DTyear[, glo, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{plot}\NormalTok{(DTyear[, Ref, year], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}

\FunctionTok{rmserr}\NormalTok{(DTyear}\SpecialCharTok{$}\NormalTok{GLB, DTyear}\SpecialCharTok{$}\NormalTok{Thr)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
$mae
[1] 97.55895

$mse
[1] 9799.437

$rmse
[1] 98.9921

$mape
[1] 0.1680595

$nmse
[1] 46.16276

$rstd
[1] 0.1700543
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DTmonth }\OtherTok{\textless{}{-}}\NormalTok{ DATA[, .(}\AttributeTok{glo =} \FunctionTok{mean}\NormalTok{(Low\_B.Low\_W.glo),}
                    \AttributeTok{GLB =} \FunctionTok{mean}\NormalTok{(wattGLB),}
                    \AttributeTok{Thr =} \FunctionTok{mean}\NormalTok{(Thresh\_test),}
\NormalTok{                    .N),}
\NormalTok{                by }\OtherTok{=}\NormalTok{ .(}\FunctionTok{year}\NormalTok{(Date), }\FunctionTok{month}\NormalTok{(Date))]}
\NormalTok{DTmonth[, tsy }\SpecialCharTok{:=}\NormalTok{ year }\SpecialCharTok{+}\NormalTok{ (month }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{)}\SpecialCharTok{/}\DecValTok{12}\NormalTok{ ]}

\NormalTok{ylim }\OtherTok{\textless{}{-}} \FunctionTok{range}\NormalTok{(DTmonth[, glo, GLB])}

\CommentTok{\# plot(  DTmonth[, glo, tsy],}
\CommentTok{\#        ylim = ylim)}
\CommentTok{\# points(DTmonth[, GLB, tsy], col = "green")}

\CommentTok{\# plot(  DTmonth[, .N, tsy], col = "blue")}
\FunctionTok{plot}\NormalTok{(  DTmonth[, GLB}\SpecialCharTok{/}\NormalTok{glo, tsy], }\AttributeTok{col =} \StringTok{"red"}\NormalTok{)}
\FunctionTok{title}\NormalTok{(}\StringTok{"Clear GLB / CS libratran by month"}\NormalTok{)}

\NormalTok{knitr}\SpecialCharTok{::}\NormalTok{opts\_chunk}\SpecialCharTok{$}\FunctionTok{set}\NormalTok{(}\AttributeTok{out.height =} \StringTok{"32\%"}\NormalTok{   )}

\NormalTok{Ratio\_lim }\OtherTok{\textless{}{-}} \FloatTok{0.85}

\NormalTok{DTdaily[GLB\_Sum}\SpecialCharTok{/}\NormalTok{glo\_Sum }\SpecialCharTok{\textless{}}\NormalTok{ Ratio\_lim, .N]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] 134
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{days }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(}\FunctionTok{c}\NormalTok{(DTdaily[GLB\_Sum}\SpecialCharTok{/}\NormalTok{glo\_Sum }\SpecialCharTok{\textless{}}\NormalTok{ Ratio\_lim, Day]))}

\NormalTok{days }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(DTdaily[, Day], }\DecValTok{40}\NormalTok{)}

\CommentTok{\# for (ad in sort(days)) \{}
\CommentTok{\#   temp \textless{}{-} DATA[Day == ad,]}
\CommentTok{\#}
\CommentTok{\#   ylim \textless{}{-} range(temp[, Low\_B.Low\_W.glo, wattGLB])}
\CommentTok{\#}
\CommentTok{\#   plot(  temp[, Low\_B.Low\_W.glo, Date],}
\CommentTok{\#          col = "blue",}
\CommentTok{\#          ylim = ylim)}
\CommentTok{\#   points(temp[, wattGLB, Date], col = "green")}
\CommentTok{\#   title(as.Date(ad, origin = "1970{-}01{-}01"))}
\CommentTok{\# \}}
\end{Highlighting}
\end{Shaded}

\textbf{END}

\end{document}