### 2024-05-09 GHI_enh_04_investigate.tex
Commit 2024-05-10 00:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-05-09}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}

Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:data.table':

    yearmon, yearqtr
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}
Registered S3 methods overwritten by 'ggpp':
  method                  from   
  heightDetails.titleGrob ggplot2
  widthDetails.titleGrob  ggplot2
\end{verbatim}

\begin{verbatim}

Attaching package: 'ggpp'
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggh4x':

    stat_centroid
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggplot2':

    annotate
\end{verbatim}

\begin{verbatim}
Registered S3 method overwritten by 'ggpmisc':
  method                  from   
  as.character.polynomial polynom
\end{verbatim}

\begin{verbatim}
ST_daily ST_E_daily ST_E_daily_seas ST_extreme_daily
\end{verbatim}

\newpage
\FloatBarrier

\hypertarget{daily-trends}{%
\subsubsection{Daily Trends}\label{daily-trends}}

\begin{verbatim}
Warning in ans * length(l) + if1: longer object length is not a multiple of
shorter object length
\end{verbatim}

\newpage

\begin{verbatim}
Loading required package: stringr
\end{verbatim}

\hypertarget{trends-on-daily-observations-data}{%
\paragraph{Trends on Daily Observations data}\label{trends-on-daily-observations-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.56672 +/- 0.105912

lmY: 0.566719 +/- 0.105914

Arima: 0.567051 +/- 0.141823

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-1} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: -0.300529 +/- 0.154885

lmY: -0.300528 +/- 0.154888

Arima: -0.2336 +/- 0.235583

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-2} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: -0.247556 +/- 0.167713

lmY: -0.247553 +/- 0.167716

Arima: -0.179333 +/- 0.249282

Units for GLB\_diff.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
\end{verbatim}

\begin{verbatim}
Warning in sqrt(diag(se)): NaNs produced
Warning in sqrt(diag(se)): NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-3} \end{center}

Arima: -0.013183 +/- 6.898051

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-4} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: -0.450424 +/- 0.239956

lmY: -0.450441 +/- 0.23996

Arima: -0.450398 +/- 0.364663

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-5} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -0.558109 +/- 0.253332

lmY: -0.558126 +/- 0.253337

Arima: -0.55808 +/- 0.388026

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-6} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.234369 +/- 0.074603

lmY: 0.234366 +/- 0.074604

Arima: 0.234368 +/- 0.110217

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-7} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.302033 +/- 0.07107

lmY: 0.302031 +/- 0.071072

Arima: 0.302035 +/- 0.101344

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-8} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: -12.24511 +/- 5.698072

lmY: -12.24538 +/- 5.698171

Arima: -12.195036 +/- 8.900074

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-9} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.00134 +/- 0.000272

lmY: 0.00134 +/- 0.000272

Arima: 0.001363 +/- 0.00036

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-10} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.000403 +/- 0.000279

lmY: 0.000403 +/- 0.000279

Arima: 0.000564 +/- 0.000462

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-11} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.000343 +/- 0.000314

lmY: 0.000343 +/- 0.000314

Arima: 0.000504 +/- 0.000507

Units for GLB\_ench.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-12} \end{center}

Arima: -0.013183 +/- 6.898051

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-13} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 0.000941 +/- 0.000309

lmY: 0.000941 +/- 0.000309

Arima: 0.000941 +/- 0.000482

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-14} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 0.000981 +/- 0.000301

lmY: 0.000982 +/- 0.000301

Arima: 0.000982 +/- 0.000462

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-15} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.000734 +/- 0.000235

lmY: 0.000734 +/- 0.000235

Arima: 0.000734 +/- 0.000366

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-16} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.000872 +/- 0.000229

lmY: 0.000872 +/- 0.000229

Arima: 0.000872 +/- 0.000342

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-17} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.180964 +/- 0.145233

lmY: 0.180972 +/- 0.145235

Arima: 0.238782 +/- 0.22021

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-18} \end{center}

\newpage

\hypertarget{trends-on-daily-enhancements-data}{%
\paragraph{Trends on Daily Enhancements data}\label{trends-on-daily-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.127247 +/- 0.099723

lmY: 0.127245 +/- 0.099725

Arima: 0.127249 +/- 0.115059

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-19} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.065571 +/- 0.036505

lmY: 0.065571 +/- 0.036506

Arima: 0.065573 +/- 0.04

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-20} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 0.085245 +/- 0.036172

lmY: 0.085246 +/- 0.036173

Arima: 0.085247 +/- 0.038483

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-21} \end{center}

Units for GLB\_diff.N :

lmD: 0.015901 +/- 0.044784

lmY: 0.015898 +/- 0.044784

Arima: 0.015898 +/- 0.052937

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-22} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: 0.033836 +/- 0.018543

lmY: 0.033837 +/- 0.018543

Arima: 0.033838 +/- 0.01905

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-23} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: 0.03242 +/- 0.017152

lmY: 0.032421 +/- 0.017152

Arima: 0.032421 +/- 0.017569

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-24} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.074226 +/- 0.070335

lmY: 0.074225 +/- 0.070336

Arima: 0.074226 +/- 0.077768

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-25} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.087786 +/- 0.080412

lmY: 0.087785 +/- 0.080413

Arima: 0.087788 +/- 0.090783

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-26} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 0.071192 +/- 0.144022

lmY: 0.071187 +/- 0.144025

Arima: 0.071193 +/- 0.173121

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-27} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.000213 +/- 0.000135

lmY: 0.000213 +/- 0.000135

Arima: 0.000213 +/- 0.000155

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-28} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.000117 +/- 5.9e-05

lmY: 0.000117 +/- 5.9e-05

Arima: 0.000117 +/- 6.8e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-29} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.000135 +/- 6.2e-05

lmY: 0.000135 +/- 6.2e-05

Arima: 0.000135 +/- 6.9e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-30} \end{center}

Units for GLB\_ench.N :

lmD: 0.015901 +/- 0.044784

lmY: 0.015898 +/- 0.044784

Arima: 0.015898 +/- 0.052937

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-31} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 5.8e-05 +/- 3.1e-05

lmY: 5.8e-05 +/- 3.1e-05

Arima: 5.8e-05 +/- 3.6e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-32} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 5.3e-05 +/- 2.8e-05

lmY: 5.3e-05 +/- 2.8e-05

Arima: 5.3e-05 +/- 3.2e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-33} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.000174 +/- 0.000106

lmY: 0.000174 +/- 0.000106

Arima: 0.000174 +/- 0.000121

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-34} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.000194 +/- 0.000116

lmY: 0.000194 +/- 0.000116

Arima: 0.000194 +/- 0.000134

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-35} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.001754 +/- 0.003555

lmY: 0.001754 +/- 0.003556

Arima: 0.001754 +/- 0.00409

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-36} \end{center}

\newpage

\hypertarget{trends-on-daily-extreme-enhancements-data}{%
\paragraph{Trends on Daily Extreme Enhancements data}\label{trends-on-daily-extreme-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: -1.406706 +/- 0.24506

lmY: -1.406695 +/- 0.245065

Arima: -1.417205 +/- 0.351582

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-37} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: -1.527566 +/- 0.186094

lmY: -1.527568 +/- 0.186097

Arima: -1.558334 +/- 0.32687

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-38} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: -1.488694 +/- 0.184965

lmY: -1.488697 +/- 0.184968

Arima: -1.520192 +/- 0.31545

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-39} \end{center}

Units for GLB\_diff.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-40} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: -1.626611 +/- 0.17175

lmY: -1.626625 +/- 0.171753

Arima: -1.666679 +/- 0.318198

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-41} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -1.642527 +/- 0.171263

lmY: -1.642542 +/- 0.171266

Arima: -1.682227 +/- 0.316737

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-42} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: -1.425903 +/- 0.224826

lmY: -1.425893 +/- 0.22483

Arima: -1.440441 +/- 0.339186

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-43} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: -1.416548 +/- 0.23349

lmY: -1.416538 +/- 0.233494

Arima: -1.429067 +/- 0.34436

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-44} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 0.057623 +/- 0.247504

lmY: 0.057642 +/- 0.247508

Arima: 0.05775 +/- 0.252236

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-45} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: -0.001993 +/- 0.000291

lmY: -0.001994 +/- 0.000291

Arima: -0.001994 +/- 0.000321

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-46} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: -0.002127 +/- 0.000226

lmY: -0.002127 +/- 0.000226

Arima: -0.002131 +/- 0.000268

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-47} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: -0.002106 +/- 0.000234

lmY: -0.002106 +/- 0.000234

Arima: -0.00211 +/- 0.000274

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-48} \end{center}

Units for GLB\_ench.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-49} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: -0.002288 +/- 0.000208

lmY: -0.002288 +/- 0.000208

Arima: -0.002296 +/- 0.000261

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-50} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: -0.0023 +/- 0.000208

lmY: -0.0023 +/- 0.000208

Arima: -0.002309 +/- 0.000261

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-51} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: -0.001986 +/- 0.000271

lmY: -0.001986 +/- 0.000271

Arima: -0.001986 +/- 0.000303

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-52} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: -0.001985 +/- 0.00028

lmY: -0.001985 +/- 0.00028

Arima: -0.001985 +/- 0.000311

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-53} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.000428 +/- 0.005409

lmY: 0.000428 +/- 0.005409

Arima: 0.000453 +/- 0.005642

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-54} \end{center}

Arima: 0.065573 +/- 0.04

\begin{verbatim}
Warning: The S3 guide system was deprecated in ggplot2 3.5.0.
i It has been replaced by a ggproto system that can be extended.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_daily_trend-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{group-stats}{%
\subsubsection{Group stats}\label{group-stats}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-1} \end{center}

\begin{verbatim}
Warning in scale_y_log10(): log-10 transformation introduced infinite values.
\end{verbatim}

\begin{verbatim}
Warning: Removed 23 rows containing missing values or values outside the scale
range (`geom_bar()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-2} \end{center}

\begin{verbatim}
Warning: A numeric `legend.position` argument in `theme()` was deprecated in ggplot2 3.5.0.
i Please use the `legend.position.inside` argument of `theme()` instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-6} \end{center}

\hypertarget{zhang2018}{%
\subsection{@Zhang2018}\label{zhang2018}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-8} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 18 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-1} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 18 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{verbatim}
Warning: Removed 1 row containing missing values or values outside the scale
range (`geom_tile()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-2} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 18 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-3} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 18 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_N_doy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-1} \end{center}

\begin{verbatim}
Warning: The `fun.y` argument of `stat_summary()` is deprecated as of ggplot2 3.3.0.
i Please use the `fun` argument instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climECEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_daily-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_week-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{energy-contribution-of-enhancements}{%
\subsubsection{Energy contribution of enhancements}\label{energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-8} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-9} \end{center}

lmY: 50.70791 +/- 35.58152

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-1} \end{center}

lmY: 20.14994 +/- 11.43911

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-2} \end{center}

lmY: 0.032227 +/- 0.053277

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-3} \end{center}

lmY: 0.036732 +/- 0.039744

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_multi-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-3} \end{center}

\newpage
\FloatBarrier

\hypertarget{monthly-energy-contribution-of-enhancements}{%
\subsubsection{Monthly Energy contribution of enhancements}\label{monthly-energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-6} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-7} \end{center}

\textbf{END}

\begin{verbatim}
2024-05-09 20:12:01.6 athan@sagan GHI_enh_04_investigate.R 0.834317 mins
\end{verbatim}

\end{document}