### 2024-05-14 GHI_enh_04_investigate.tex
Commit 2024-05-14 11:35

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-05-14}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:data.table':

    yearmon, yearqtr
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}
Registered S3 methods overwritten by 'ggpp':
  method                  from   
  heightDetails.titleGrob ggplot2
  widthDetails.titleGrob  ggplot2
\end{verbatim}

\begin{verbatim}
Attaching package: 'ggpp'
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggh4x':

    stat_centroid
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggplot2':

    annotate
\end{verbatim}

\begin{verbatim}
Registered S3 method overwritten by 'ggpmisc':
  method                  from   
  as.character.polynomial polynom
\end{verbatim}

\begin{verbatim}
ST_daily ST_E_daily ST_E_daily_seas ST_extreme_daily
\end{verbatim}

\newpage
\FloatBarrier

\hypertarget{daily-trends}{%
\subsubsection{Daily Trends}\label{daily-trends}}

\begin{verbatim}
Warning in ans * length(l) + if1: longer object length is not a multiple of
shorter object length
\end{verbatim}

\newpage

\begin{verbatim}
Loading required package: stringr
\end{verbatim}

\hypertarget{trends-on-daily-observations-data}{%
\paragraph{Trends on Daily Observations data}\label{trends-on-daily-observations-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.80888 +/- 0.10628

lmY: 0.808883 +/- 0.106282

Arima: 0.809228 +/- 0.142401

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-1} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.002077 +/- 0.154863

lmY: 0.002083 +/- 0.154866

Arima: 0.069972 +/- 0.235505

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-2} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 0.0726 +/- 0.167744

lmY: 0.072609 +/- 0.167747

Arima: 0.141941 +/- 0.249338

Units for GLB\_diff.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
\end{verbatim}

\begin{verbatim}
Warning in sqrt(diag(se)): NaNs produced
Warning in sqrt(diag(se)): NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-3} \end{center}

Arima: -0.013183 +/- 6.904159

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-4} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: -0.087078 +/- 0.239888

lmY: -0.087089 +/- 0.239892

Arima: -0.087046 +/- 0.364497

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-5} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -0.191061 +/- 0.253235

lmY: -0.191072 +/- 0.253239

Arima: -0.191026 +/- 0.387751

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-6} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.442171 +/- 0.074674

lmY: 0.442172 +/- 0.074676

Arima: 0.442175 +/- 0.110249

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-7} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.506855 +/- 0.071215

lmY: 0.506856 +/- 0.071216

Arima: 0.50686 +/- 0.101481

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-8} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: -1.191149 +/- 5.69247

lmY: -1.191222 +/- 5.692569

Arima: -1.138138 +/- 8.879593

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-9} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.00182 +/- 0.000272

lmY: 0.00182 +/- 0.000272

Arima: 0.001843 +/- 0.000361

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-10} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.000729 +/- 0.00028

lmY: 0.00073 +/- 0.00028

Arima: 0.000892 +/- 0.000462

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-11} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.000678 +/- 0.000314

lmY: 0.000678 +/- 0.000314

Arima: 0.000841 +/- 0.000507

Units for GLB\_ench.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-12} \end{center}

Arima: -0.013183 +/- 6.904159

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-13} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 0.001176 +/- 0.000309

lmY: 0.001176 +/- 0.000309

Arima: 0.001176 +/- 0.000483

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-14} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 0.001189 +/- 0.000301

lmY: 0.001189 +/- 0.000301

Arima: 0.001189 +/- 0.000463

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-15} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.001144 +/- 0.000235

lmY: 0.001144 +/- 0.000235

Arima: 0.001144 +/- 0.000366

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-16} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.0013 +/- 0.00023

lmY: 0.001301 +/- 0.00023

Arima: 0.001301 +/- 0.000342

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-17} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.376507 +/- 0.145413

lmY: 0.376518 +/- 0.145415

Arima: 0.436011 +/- 0.220423

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-18} \end{center}

\newpage

\hypertarget{trends-on-daily-enhancements-data}{%
\paragraph{Trends on Daily Enhancements data}\label{trends-on-daily-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.319795 +/- 0.10019

lmY: 0.319795 +/- 0.100192

Arima: 0.3198 +/- 0.116665

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-19} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.10702 +/- 0.036516

lmY: 0.107021 +/- 0.036517

Arima: 0.107023 +/- 0.040195

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-20} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 0.108459 +/- 0.036231

lmY: 0.108461 +/- 0.036232

Arima: 0.108462 +/- 0.03848

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-21} \end{center}

Units for GLB\_diff.N :

lmD: 0.144267 +/- 0.045486

lmY: 0.144265 +/- 0.045487

Arima: 0.144267 +/- 0.054102

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-22} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: 0.022726 +/- 0.0184

lmY: 0.022726 +/- 0.0184

Arima: 0.022726 +/- 0.018993

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-23} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: 0.018838 +/- 0.016889

lmY: 0.018838 +/- 0.016889

Arima: 0.018838 +/- 0.017339

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-24} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.194707 +/- 0.070475

lmY: 0.194706 +/- 0.070476

Arima: 0.194713 +/- 0.07858

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-25} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.227625 +/- 0.080608

lmY: 0.227625 +/- 0.080609

Arima: 0.22763 +/- 0.091764

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-26} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 0.447759 +/- 0.147315

lmY: 0.447759 +/- 0.147317

Arima: 0.447766 +/- 0.177904

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-27} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.000537 +/- 0.000135

lmY: 0.000537 +/- 0.000135

Arima: 0.000537 +/- 0.000157

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-28} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.000206 +/- 6e-05

lmY: 0.000206 +/- 6e-05

Arima: 0.000206 +/- 6.8e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-29} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.000195 +/- 6.2e-05

lmY: 0.000195 +/- 6.2e-05

Arima: 0.000195 +/- 7e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-30} \end{center}

Units for GLB\_ench.N :

lmD: 0.144267 +/- 0.045486

lmY: 0.144265 +/- 0.045487

Arima: 0.144267 +/- 0.054102

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-31} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 4.8e-05 +/- 3.1e-05

lmY: 4.8e-05 +/- 3.1e-05

Arima: 4.8e-05 +/- 3.6e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-32} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 3.7e-05 +/- 2.8e-05

lmY: 3.7e-05 +/- 2.8e-05

Arima: 3.7e-05 +/- 3.3e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-33} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.000394 +/- 0.000106

lmY: 0.000394 +/- 0.000106

Arima: 0.000394 +/- 0.000122

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-34} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.00045 +/- 0.000117

lmY: 0.00045 +/- 0.000117

Arima: 0.00045 +/- 0.000135

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-35} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.011422 +/- 0.003629

lmY: 0.011422 +/- 0.003629

Arima: 0.011422 +/- 0.004188

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-36} \end{center}

\newpage

\hypertarget{trends-on-daily-extreme-enhancements-data}{%
\paragraph{Trends on Daily Extreme Enhancements data}\label{trends-on-daily-extreme-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: -1.014557 +/- 0.247144

lmY: -1.014539 +/- 0.247148

Arima: -1.025903 +/- 0.356761

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-37} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: -1.13974 +/- 0.188386

lmY: -1.139735 +/- 0.188389

Arima: -1.172188 +/- 0.333189

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-38} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: -1.100642 +/- 0.18719

lmY: -1.100638 +/- 0.187194

Arima: -1.133859 +/- 0.321558

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-39} \end{center}

Units for GLB\_diff.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-40} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: -1.242646 +/- 0.174098

lmY: -1.242654 +/- 0.174101

Arima: -1.284884 +/- 0.324863

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-41} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -1.259775 +/- 0.173629

lmY: -1.259783 +/- 0.173632

Arima: -1.301601 +/- 0.323343

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-42} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: -1.034796 +/- 0.226955

lmY: -1.034779 +/- 0.226959

Arima: -1.050406 +/- 0.344658

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-43} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: -1.02486 +/- 0.235592

lmY: -1.024843 +/- 0.235596

Arima: -1.038346 +/- 0.349707

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-44} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 0.138738 +/- 0.251457

lmY: 0.138759 +/- 0.251461

Arima: 0.138867 +/- 0.256234

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-45} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: -0.001372 +/- 0.000292

lmY: -0.001372 +/- 0.000292

Arima: -0.001372 +/- 0.000322

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-46} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: -0.001515 +/- 0.000226

lmY: -0.001515 +/- 0.000226

Arima: -0.001519 +/- 0.000269

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-47} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: -0.001494 +/- 0.000234

lmY: -0.001494 +/- 0.000234

Arima: -0.001498 +/- 0.000275

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-48} \end{center}

Units for GLB\_ench.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-49} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: -0.001682 +/- 0.000209

lmY: -0.001683 +/- 0.000209

Arima: -0.00169 +/- 0.000262

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-50} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: -0.001696 +/- 0.000208

lmY: -0.001696 +/- 0.000208

Arima: -0.001704 +/- 0.000262

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-51} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: -0.001367 +/- 0.000272

lmY: -0.001367 +/- 0.000272

Arima: -0.001367 +/- 0.000304

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-52} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: -0.001364 +/- 0.000281

lmY: -0.001364 +/- 0.000281

Arima: -0.001365 +/- 0.000312

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-53} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.002612 +/- 0.005505

lmY: 0.002612 +/- 0.005505

Arima: 0.002638 +/- 0.005739

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-54} \end{center}

Arima: 0.107023 +/- 0.040196

\begin{verbatim}
Warning: The S3 guide system was deprecated in ggplot2 3.5.0.
i It has been replaced by a ggproto system that can be extended.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_daily_trend-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{group-stats}{%
\subsubsection{Group stats}\label{group-stats}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-1} \end{center}

\begin{verbatim}
Warning in scale_y_log10(): log-10 transformation introduced infinite values.
\end{verbatim}

\begin{verbatim}
Warning: Removed 23 rows containing missing values or values outside the scale
range (`geom_bar()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-2} \end{center}

\begin{verbatim}
Warning: A numeric `legend.position` argument in `theme()` was deprecated in ggplot2 3.5.0.
i Please use the `legend.position.inside` argument of `theme()` instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-6} \end{center}

\hypertarget{zhang2018}{%
\subsection{@Zhang2018}\label{zhang2018}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-8} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 19 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-1} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 19 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{verbatim}
Warning: Removed 2 rows containing missing values or values outside the scale
range (`geom_tile()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-2} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 19 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-3} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 19 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_N_doy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-1} \end{center}

\begin{verbatim}
Warning: The `fun.y` argument of `stat_summary()` is deprecated as of ggplot2 3.3.0.
i Please use the `fun` argument instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climECEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_daily-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_week-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{energy-contribution-of-enhancements}{%
\subsubsection{Energy contribution of enhancements}\label{energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-8} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-9} \end{center}

lmY: 129.677 +/- 37.08333

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-1} \end{center}

lmY: 48.08231 +/- 11.89152

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-2} \end{center}

lmY: 0.080779 +/- 0.053347

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-3} \end{center}

lmY: 0.065639 +/- 0.040043

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_multi-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-3} \end{center}

\newpage
\FloatBarrier

\hypertarget{monthly-energy-contribution-of-enhancements}{%
\subsubsection{Monthly Energy contribution of enhancements}\label{monthly-energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-6} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-7} \end{center}

\textbf{END}

\begin{verbatim}
2024-05-14 07:45:19.2 athan@tyler GHI_enh_04_investigate.R 2.165741 mins
\end{verbatim}

\end{document}