### 2024-05-14 GHI_enh_Aeronet.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-05-14}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
Cross over point: 2005 
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-5} \end{center}

\begin{verbatim}
[1] 1997
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-12} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-4-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-12} \end{center}

SZA min : 0.00137165912186399

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-20} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-24} \end{center}

SZA mean : 0.0020024477410843

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-28} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-31} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-32} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-33} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-34} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-35} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-36} \end{center}

SZA median : 0.00184853525463023

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-37} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-38} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-39} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-40} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-41} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-42} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-43} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-44} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-45} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-46} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-47} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-48} \end{center}

SZA 55 : 0.00137369825595582

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1579}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1184}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1053}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1316}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1711}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1579}}
  >{\centering\arraybackslash}p{(\columnwidth - 12\tabcolsep) * \real{0.1579}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Intercept
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Slope
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Month
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Type
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Zeropoint
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
SlopeRela
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
-1102 & 0.8309 & 1 & 565.8 & SZA min & 2014 & 0.001469 \\
-810.5 & 0.7648 & 2 & 721.8 & SZA min & 2015 & 0.00106 \\
-6892 & 3.872 & 3 & 888.2 & SZA min & 2015 & 0.00436 \\
-1277 & 1.124 & 4 & 974.5 & SZA min & 2015 & 0.001153 \\
-982.7 & 0.9904 & 5 & 1004 & SZA min & 2015 & 0.0009866 \\
375.1 & 0.3116 & 6 & 993.2 & SZA min & 2015 & 0.0003138 \\
-140.7 & 0.5624 & 7 & 984 & SZA min & 2015 & 0.0005715 \\
444.2 & 0.2576 & 8 & 967.6 & SZA min & 2015 & 0.0002663 \\
-3.347 & 0.4436 & 9 & 872.7 & SZA min & 2015 & 0.0005082 \\
-1081 & 0.9159 & 10 & 758.3 & SZA min & 2015 & 0.001208 \\
-2109 & 1.345 & 11 & 565.3 & SZA min & 2015 & 0.002379 \\
-1598 & 1.044 & 12 & 477.6 & SZA min & 2014 & 0.002186 \\
-1221 & 0.756 & 1 & 293.9 & SZA mean & 2014 & 0.002573 \\
-1063 & 0.7194 & 2 & 377.8 & SZA mean & 2015 & 0.001904 \\
-1776 & 1.125 & 3 & 485.6 & SZA mean & 2015 & 0.002317 \\
-1680 & 1.118 & 4 & 561.4 & SZA mean & 2015 & 0.001992 \\
-1628 & 1.111 & 5 & 599.9 & SZA mean & 2015 & 0.001852 \\
-262.2 & 0.4374 & 6 & 606.9 & SZA mean & 2015 & 0.0007207 \\
-745.5 & 0.6689 & 7 & 592.1 & SZA mean & 2015 & 0.00113 \\
-122.7 & 0.3402 & 8 & 566.4 & SZA mean & 2015 & 0.0006007 \\
-474.5 & 0.4816 & 9 & 477.8 & SZA mean & 2015 & 0.001008 \\
-1443 & 0.9151 & 10 & 394.4 & SZA mean & 2015 & 0.00232 \\
-2438 & 1.364 & 11 & 277.9 & SZA mean & 2015 & 0.004907 \\
-1179 & 0.7187 & 12 & 265.7 & SZA mean & 2014 & 0.002705 \\
-1216 & 0.772 & 1 & 331.6 & SZA median & 2014 & 0.002328 \\
-1041 & 0.7293 & 2 & 419.7 & SZA median & 2015 & 0.001738 \\
-1743 & 1.129 & 3 & 527.8 & SZA median & 2015 & 0.00214 \\
-1654 & 1.123 & 4 & 595.7 & SZA median & 2015 & 0.001884 \\
-1595 & 1.105 & 5 & 620.6 & SZA median & 2015 & 0.00178 \\
-243.2 & 0.4343 & 6 & 619.6 & SZA median & 2015 & 0.0007009 \\
-708.1 & 0.6584 & 7 & 608.4 & SZA median & 2015 & 0.001082 \\
-109.2 & 0.3477 & 8 & 595.2 & SZA median & 2015 & 0.0005842 \\
-441.5 & 0.4846 & 9 & 516.9 & SZA median & 2015 & 0.0009376 \\
-1433 & 0.9306 & 10 & 435.1 & SZA median & 2015 & 0.002139 \\
-2464 & 1.395 & 11 & 314.5 & SZA median & 2015 & 0.004436 \\
-1176 & 0.7354 & 12 & 302.4 & SZA median & 2014 & 0.002432 \\
-847.8 & 0.7326 & 1 & 621.4 & SZA 55 & 2014 & 0.001179 \\
-787.2 & 0.6968 & 2 & 608.5 & SZA 55 & 2015 & 0.001145 \\
-1620 & 1.102 & 3 & 596.7 & SZA 55 & 2015 & 0.001848 \\
-1575 & 1.071 & 4 & 570.5 & SZA 55 & 2015 & 0.001877 \\
-1615 & 1.079 & 5 & 549.6 & SZA 55 & 2015 & 0.001964 \\
-288.8 & 0.4157 & 6 & 536.3 & SZA 55 & 2015 & 0.000775 \\
-483.7 & 0.5087 & 7 & 531.2 & SZA 55 & 2015 & 0.0009577 \\
-140.4 & 0.3406 & 8 & 549.3 & SZA 55 & 2015 & 0.0006201 \\
-415.5 & 0.4853 & 9 & 544.7 & SZA 55 & 2015 & 0.000891 \\
-1327 & 0.9475 & 10 & 576 & SZA 55 & 2015 & 0.001645 \\
-2204 & 1.393 & 11 & 567.8 & SZA 55 & 2015 & 0.002453 \\
-785.3 & 0.6983 & 12 & 618.3 & SZA 55 & 2014 & 0.001129 \\
\end{longtable}

\newpage

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0972}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
year
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
typer
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1997 & BR SZA median & 449.5 \\
2005 & SZA median & 392.1 \\
2005 & BR SZA median & 459.3 \\
2006 & SZA median & 493.1 \\
2007 & SZA median & 470.8 \\
2008 & SZA median & 326.2 \\
2011 & SZA median & 505.1 \\
2012 & SZA median & 500.8 \\
2013 & SZA median & 498.6 \\
2014 & SZA median & 477.7 \\
2015 & SZA median & 504.8 \\
2016 & SZA median & 510.9 \\
2017 & SZA median & 551.4 \\
2018 & SZA median & 499.7 \\
2019 & SZA median & 476 \\
2020 & SZA median & 483.7 \\
2021 & SZA median & 506 \\
2022 & SZA median & 554.9 \\
2023 & SZA median & 522.8 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-49} \end{center}

\newpage

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0972}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
year
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
typer
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1997 & BR SZA min & 1055 \\
2005 & SZA min & 668.3 \\
2005 & BR SZA min & 1064 \\
2006 & SZA min & 817.1 \\
2007 & SZA min & 785.1 \\
2008 & SZA min & 557.2 \\
2011 & SZA min & 831.1 \\
2012 & SZA min & 824.3 \\
2013 & SZA min & 814 \\
2014 & SZA min & 822.1 \\
2015 & SZA min & 828.6 \\
2016 & SZA min & 832.3 \\
2017 & SZA min & 899.3 \\
2018 & SZA min & 822.3 \\
2019 & SZA min & 789.7 \\
2020 & SZA min & 820.3 \\
2021 & SZA min & 829.7 \\
2022 & SZA min & 902.4 \\
2023 & SZA min & 859.5 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-50} \end{center}

\newpage

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0972}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1944}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1389}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
year
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
typer
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1997 & BR SZA mean & 497.3 \\
2005 & SZA mean & 353.9 \\
2005 & BR SZA mean & 507.1 \\
2006 & SZA mean & 460.7 \\
2007 & SZA mean & 434.8 \\
2008 & SZA mean & 288.6 \\
2011 & SZA mean & 476.1 \\
2012 & SZA mean & 468.4 \\
2013 & SZA mean & 468.5 \\
2014 & SZA mean & 444.8 \\
2015 & SZA mean & 472.3 \\
2016 & SZA mean & 480.3 \\
2017 & SZA mean & 520.9 \\
2018 & SZA mean & 467.3 \\
2019 & SZA mean & 440.6 \\
2020 & SZA mean & 450.8 \\
2021 & SZA mean & 473.5 \\
2022 & SZA mean & 525.5 \\
2023 & SZA mean & 490.7 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-51} \end{center}

\newpage

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0972}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1667}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1667}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
year
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
typer
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MeanGlo
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1997 & BR SZA 55 & 578.9 \\
2005 & SZA 55 & 576.4 \\
2005 & BR SZA 55 & 588.6 \\
2006 & SZA 55 & 574.9 \\
2007 & SZA 55 & 584.7 \\
2008 & SZA 55 & 615 \\
2011 & SZA 55 & 562.4 \\
2012 & SZA 55 & 582.9 \\
2013 & SZA 55 & 582.1 \\
2014 & SZA 55 & 594.3 \\
2015 & SZA 55 & 586.7 \\
2016 & SZA 55 & 583.9 \\
2017 & SZA 55 & 581.3 \\
2018 & SZA 55 & 581.2 \\
2019 & SZA 55 & 584.4 \\
2020 & SZA 55 & 600.3 \\
2021 & SZA 55 & 587.5 \\
2022 & SZA 55 & 587.4 \\
2023 & SZA 55 & 583 \\
\end{longtable}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_Aeronet_files/figure-latex/unnamed-chunk-5-52} \end{center}

null device
1

\textbf{END}

\end{document}