### 2024-05-21 GHI_enh_04_investigate.tex
Commit 2024-05-21 10:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-05-20}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:data.table':

    yearmon, yearqtr
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}
Registered S3 methods overwritten by 'ggpp':
  method                  from   
  heightDetails.titleGrob ggplot2
  widthDetails.titleGrob  ggplot2
\end{verbatim}

\begin{verbatim}
Attaching package: 'ggpp'
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggh4x':

    stat_centroid
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:ggplot2':

    annotate
\end{verbatim}

\begin{verbatim}
Registered S3 method overwritten by 'ggpmisc':
  method                  from   
  as.character.polynomial polynom
\end{verbatim}

\begin{verbatim}
ST_daily ST_E_daily ST_E_daily_seas ST_extreme_daily
\end{verbatim}

\newpage
\FloatBarrier

\hypertarget{daily-trends}{%
\subsubsection{Daily Trends}\label{daily-trends}}

\begin{verbatim}
Warning in ans * length(l) + if1: longer object length is not a multiple of
shorter object length
\end{verbatim}

\newpage

\begin{verbatim}
Loading required package: stringr
\end{verbatim}

\hypertarget{trends-on-daily-observations-data}{%
\paragraph{Trends on Daily Observations data}\label{trends-on-daily-observations-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.804776 +/- 0.104681

lmY: 0.804779 +/- 0.104683

Arima: 0.805121 +/- 0.139935

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-1} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.002645 +/- 0.154816

lmY: 0.002651 +/- 0.154818

Arima: 0.069946 +/- 0.23533

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-2} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 0.075523 +/- 0.167494

lmY: 0.075532 +/- 0.167497

Arima: 0.144037 +/- 0.248629

Units for GLB\_diff.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
\end{verbatim}

\begin{verbatim}
Warning in sqrt(diag(se)): NaNs produced
Warning in sqrt(diag(se)): NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-3} \end{center}

Arima: -0.013183 +/- 6.904159

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-4} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: -0.093964 +/- 0.240209

lmY: -0.093975 +/- 0.240213

Arima: -0.093932 +/- 0.364937

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-5} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -0.197599 +/- 0.2537

lmY: -0.197611 +/- 0.253704

Arima: -0.197564 +/- 0.388597

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-6} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.446578 +/- 0.074214

lmY: 0.446579 +/- 0.074216

Arima: 0.446582 +/- 0.10962

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-7} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.508145 +/- 0.070493

lmY: 0.508147 +/- 0.070494

Arima: 0.508152 +/- 0.100518

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-8} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: -1.087297 +/- 5.707626

lmY: -1.087374 +/- 5.707725

Arima: -1.035804 +/- 8.936135

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-9} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.001815 +/- 0.00027

lmY: 0.001815 +/- 0.00027

Arima: 0.001839 +/- 0.000358

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-10} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.00073 +/- 0.000278

lmY: 0.00073 +/- 0.000278

Arima: 0.000891 +/- 0.000459

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-11} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.000679 +/- 0.000312

lmY: 0.000679 +/- 0.000312

Arima: 0.00084 +/- 0.000504

Units for GLB\_ench.N :

lmD: -0.012345 +/- 0.138879

lmY: -0.01233 +/- 0.138882

\begin{verbatim}
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
possible convergence problem: optim gave code = 1
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
Warning in arima(x = dataset[[avar]], order = c(1, 0, 0), xreg = dataset$yts, :
NaNs produced
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-12} \end{center}

Arima: -0.013183 +/- 6.904159

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-13} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 0.001171 +/- 0.000307

lmY: 0.001171 +/- 0.000307

Arima: 0.001171 +/- 0.000479

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-14} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 0.001184 +/- 0.000299

lmY: 0.001184 +/- 0.000299

Arima: 0.001184 +/- 0.00046

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-15} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.001143 +/- 0.000233

lmY: 0.001143 +/- 0.000233

Arima: 0.001143 +/- 0.000363

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-16} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.001299 +/- 0.000228

lmY: 0.001299 +/- 0.000228

Arima: 0.001299 +/- 0.00034

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-17} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.379481 +/- 0.144397

lmY: 0.379493 +/- 0.1444

Arima: 0.438045 +/- 0.218822

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-18} \end{center}

\newpage

\hypertarget{trends-on-daily-enhancements-data}{%
\paragraph{Trends on Daily Enhancements data}\label{trends-on-daily-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: 0.278784 +/- 0.099757

lmY: 0.278785 +/- 0.099759

Arima: 0.278791 +/- 0.11495

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-19} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: 0.086758 +/- 0.036893

lmY: 0.086759 +/- 0.036894

Arima: 0.086761 +/- 0.03998

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-20} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: 0.089859 +/- 0.036694

lmY: 0.089861 +/- 0.036694

Arima: 0.089862 +/- 0.038687

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-21} \end{center}

Units for GLB\_diff.N :

lmD: 0.134313 +/- 0.043256

lmY: 0.134312 +/- 0.043256

Arima: 0.134314 +/- 0.05079

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-22} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: 0.010492 +/- 0.018772

lmY: 0.010493 +/- 0.018773

Arima: 0.010494 +/- 0.019022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-23} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: 0.006132 +/- 0.017414

lmY: 0.006133 +/- 0.017414

Arima: 0.006133 +/- 0.017662

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-24} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: 0.160601 +/- 0.070705

lmY: 0.160602 +/- 0.070706

Arima: 0.160604 +/- 0.07797

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-25} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: 0.198524 +/- 0.080714

lmY: 0.198524 +/- 0.080715

Arima: 0.198532 +/- 0.091006

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-26} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 0.386176 +/- 0.138337

lmY: 0.386177 +/- 0.13834

Arima: 0.386184 +/- 0.16586

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-27} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: 0.000476 +/- 0.000135

lmY: 0.000476 +/- 0.000135

Arima: 0.000476 +/- 0.000155

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-28} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: 0.000172 +/- 6e-05

lmY: 0.000172 +/- 6e-05

Arima: 0.000172 +/- 6.8e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-29} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: 0.00016 +/- 6.3e-05

lmY: 0.00016 +/- 6.3e-05

Arima: 0.00016 +/- 7e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-30} \end{center}

Units for GLB\_ench.N :

lmD: 0.134313 +/- 0.043256

lmY: 0.134312 +/- 0.043256

Arima: 0.134314 +/- 0.05079

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-31} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: 2.8e-05 +/- 3.2e-05

lmY: 2.8e-05 +/- 3.2e-05

Arima: 2.8e-05 +/- 3.6e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-32} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: 1.7e-05 +/- 2.9e-05

lmY: 1.7e-05 +/- 2.9e-05

Arima: 1.7e-05 +/- 3.3e-05

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-33} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: 0.000347 +/- 0.000106

lmY: 0.000347 +/- 0.000106

Arima: 0.000347 +/- 0.000121

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-34} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: 0.000402 +/- 0.000116

lmY: 0.000402 +/- 0.000116

Arima: 0.000402 +/- 0.000134

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-35} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.009848 +/- 0.003421

lmY: 0.009848 +/- 0.003422

Arima: 0.009849 +/- 0.003932

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-36} \end{center}

\newpage

\hypertarget{trends-on-daily-extreme-enhancements-data}{%
\paragraph{Trends on Daily Extreme Enhancements data}\label{trends-on-daily-extreme-enhancements-data}}

Units for GLB\_diff.max : Watt/m\^{}2

lmD: -1.014326 +/- 0.242383

lmY: -1.01431 +/- 0.242387

Arima: -1.024849 +/- 0.345654

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-37} \end{center}

Units for GLB\_diff.mean : Watt/m\^{}2

lmD: -1.139227 +/- 0.182911

lmY: -1.139223 +/- 0.182914

Arima: -1.17037 +/- 0.320655

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-38} \end{center}

Units for GLB\_diff.median : Watt/m\^{}2

lmD: -1.099532 +/- 0.181744

lmY: -1.09953 +/- 0.181747

Arima: -1.131269 +/- 0.309288

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-39} \end{center}

Units for GLB\_diff.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-40} \end{center}

Units for GLB\_diff.p\_10 : Watt/m\^{}2

lmD: -1.242636 +/- 0.168369

lmY: -1.242644 +/- 0.168371

Arima: -1.283474 +/- 0.312719

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-41} \end{center}

Units for GLB\_diff.p\_5 : Watt/m\^{}2

lmD: -1.259539 +/- 0.167895

lmY: -1.259548 +/- 0.167897

Arima: -1.300028 +/- 0.311399

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-42} \end{center}

Units for GLB\_diff.p\_90 : Watt/m\^{}2

lmD: -1.033796 +/- 0.221995

lmY: -1.03378 +/- 0.221999

Arima: -1.048334 +/- 0.332949

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-43} \end{center}

Units for GLB\_diff.p\_95 : Watt/m\^{}2

lmD: -1.02427 +/- 0.230733

lmY: -1.024254 +/- 0.230737

Arima: -1.036798 +/- 0.338243

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-44} \end{center}

Units for GLB\_diff.sum : kJ/m\^{}2

lmD: 0.128338 +/- 0.244436

lmY: 0.128359 +/- 0.24444

Arima: 0.128462 +/- 0.249026

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-45} \end{center}

Units for GLB\_ench.max : Watt/m\^{}2

lmD: -0.001358 +/- 0.00029

lmY: -0.001358 +/- 0.00029

Arima: -0.001358 +/- 0.00032

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-46} \end{center}

Units for GLB\_ench.mean : Watt/m\^{}2

lmD: -0.0015 +/- 0.000225

lmY: -0.0015 +/- 0.000225

Arima: -0.001504 +/- 0.000267

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-47} \end{center}

Units for GLB\_ench.median : Watt/m\^{}2

lmD: -0.001479 +/- 0.000233

lmY: -0.001479 +/- 0.000233

Arima: -0.001483 +/- 0.000273

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-48} \end{center}

Units for GLB\_ench.N :

lmD: 0.030347 +/- 0.021245

lmY: 0.030349 +/- 0.021246

Arima: 0.030378 +/- 0.022031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-49} \end{center}

Units for GLB\_ench.p\_10 : Watt/m\^{}2

lmD: -0.001666 +/- 0.000208

lmY: -0.001666 +/- 0.000208

Arima: -0.001674 +/- 0.00026

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-50} \end{center}

Units for GLB\_ench.p\_5 : Watt/m\^{}2

lmD: -0.00168 +/- 0.000207

lmY: -0.00168 +/- 0.000207

Arima: -0.001688 +/- 0.00026

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-51} \end{center}

Units for GLB\_ench.p\_90 : Watt/m\^{}2

lmD: -0.001352 +/- 0.00027

lmY: -0.001352 +/- 0.00027

Arima: -0.001353 +/- 0.000302

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-52} \end{center}

Units for GLB\_ench.p\_95 : Watt/m\^{}2

lmD: -0.00135 +/- 0.000279

lmY: -0.00135 +/- 0.000279

Arima: -0.001351 +/- 0.00031

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-53} \end{center}

Units for GLB\_ench.sum : Watt/m\^{}2

lmD: 0.002398 +/- 0.005327

lmY: 0.002399 +/- 0.005328

Arima: 0.002424 +/- 0.005555

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/daily-54} \end{center}

Arima: 0.086761 +/- 0.03998

\begin{verbatim}
Warning: The S3 guide system was deprecated in ggplot2 3.5.0.
i It has been replaced by a ggproto system that can be extended.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_daily_trend-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{group-stats}{%
\subsubsection{Group stats}\label{group-stats}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-1} \end{center}

\begin{verbatim}
Warning in scale_y_log10(): log-10 transformation introduced infinite values.
\end{verbatim}

\begin{verbatim}
Warning: Removed 23 rows containing missing values or values outside the scale
range (`geom_bar()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-2} \end{center}

\begin{verbatim}
Warning: A numeric `legend.position` argument in `theme()` was deprecated in ggplot2 3.5.0.
i Please use the `legend.position.inside` argument of `theme()` instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-6} \end{center}

\hypertarget{zhang2018}{%
\subsection{@Zhang2018}\label{zhang2018}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-8} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-1} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{verbatim}
Warning: Removed 1 row containing missing values or values outside the scale
range (`geom_tile()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-2} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-3} \end{center}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_N_doy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-1} \end{center}

\begin{verbatim}
Warning: The `fun.y` argument of `stat_summary()` is deprecated as of ggplot2 3.3.0.
i Please use the `fun` argument instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_CE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climECEmonth-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim_ECE_month_norm_MAX_median_N-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_daily-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCE_week-1} \end{center}

\newpage
\FloatBarrier

\hypertarget{energy-contribution-of-enhancements}{%
\subsubsection{Energy contribution of enhancements}\label{energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-8} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-9} \end{center}

lmY: 116.8672 +/- 33.89196

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-1} \end{center}

lmY: 45.58799 +/- 10.96835

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-2} \end{center}

lmY: 0.062157 +/- 0.054182

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-3} \end{center}

lmY: 0.053059 +/- 0.041079

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P_energy-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_multi-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel_energy-3} \end{center}

\newpage
\FloatBarrier

\hypertarget{monthly-energy-contribution-of-enhancements}{%
\subsubsection{Monthly Energy contribution of enhancements}\label{monthly-energy-contribution-of-enhancements}}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-6} \end{center}

@Martins2022

\begin{center}\includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy_monthly-7} \end{center}

\textbf{END}

\begin{verbatim}
2024-05-20 16:56:07.0 athan@tyler GHI_enh_04_investigate.R 1.334659 mins
\end{verbatim}

\end{document}