### 2024-05-26 Article_B5/Article_B5.Rmd
Commit 2024-05-27 00:14

---
title: Long term cloud enhancement events of global solar irradiance over Thessaloniki

author:
  - name:        Athanasios N. Natsis
    email:       natsisphysicist@gmail.com
    affiliation: LAP
    correspondingauthor: true
    footnote: 1
  - name:        Alkiviadis Bais
    email:       abais@auth.gr
    affiliation: LAP
  - name:        Charikleia Meleti
    email:       meleti@auth.gr
    affiliation: LAP
    footnote: 1
address:
  - code:         LAP
    organization: Laboratory of Atmospheric Physics, Physics Department, Aristotle University of Thessaloniki
    # addressline:  1 main street
    city:         Thessaloniki
    postcode:     54124
    country:      Greece
  # - code:         Another University
  #   organization: Department
  #   addressline:  A street 29
  #   postcode:     2054 NX
  #   city:         Manchester,
  #   state:        State
  #   country:      The Netherlands
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  Here will be the abstract.

header-includes:
  - \usepackage{caption}
  - \usepackage{placeins}
  - \captionsetup{font=small}

keywords: 
  - cloud enhancement
  - total solar radiation
  - global horizontal irradiance
  - over irradiance

journal:        "An awesome journal"
date:           "`r Sys.Date()`"
linenumbers:    false
numbersections: true
bibliography:   bibliography.bib

biblio-style:   elsarticle-harv # author year style for natbib - use 'elsarticle-num' or 'elsarticle-num-names' for numbered scheme

classoption:    preprint, 3p, authoryear # remove authoryear is not using `elsarticle-harv`
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard

output: 
  rticles::elsevier_article:
    keep_tex:         true
    citation_package: natbib
  bookdown::pdf_document2:
    number_sections:  yes
    fig_caption:      yes
    keep_tex:         yes
    latex_engine:     xelatex
    toc:              yes
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
  bookdown::word_document2: default
---

```{r setup, echo=F, include=FALSE}
rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
# renv::load("~/MANUSCRIPTS/02_enhancement/", quiet = TRUE)
## document configurations
knitr::opts_chunk$set(comment    = ""      )
# knitr::opts_chunk$set(dev        = "pdf"   )
knitr::opts_chunk$set(out.width  = "50%"   )
knitr::opts_chunk$set(fig.align  = "center")
# knitr::opts_chunk$set(fig.pos    = 'h!'    )
options(knitr.graphics.auto_pdf = FALSE)
library(pander)
library(data.table)
library(kableExtra)
library(dplyr)
## load data from trends analysis
source("~/MANUSCRIPTS/02_enhancement/GHI_enh_00_variables.R")
load("~/MANUSCRIPTS/02_enhancement/data/GHI_enh_03_process.Rda")
TSIinfo <- data.table(read.csv("~/MANUSCRIPTS/02_enhancement/figures/tbl_tsi_info.dat",
                               strip.white = TRUE,
                               sep = ";",
                               header = TRUE))
```

```{r loadtrends, echo=F, include=T, warning=FALSE, message=FALSE}
dailytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear.csv"))
dailyCE <- dailytrends[DATA == "ST_E_daily"]
yearlytrends <- data.table(
  read.csv("~/MANUSCRIPTS/02_enhancement/figures/Daily_trends_byYear_Proper.csv"))
yearlyCE <- yearlytrends[DATA == "ST_E_yearly"]

```

# Abstract {.unnumbered}

We studied the enhancement of the global horizontal irradiance (GHI), over Thessaloniki, 
for the period 1993 -- 2023, on one minute average records.
We identify the cloud enhancement events (CE), by creating an appropriate clear sky
irradiance reference with the use of a radiation transfer model, and aerosol optical 
depth (AOD) data, from the AERONET, and a collocated Brewer photospectrometer.
We found that there is a trend of CE cases of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
and a
mean total energy change, of the CE events by
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kj}/\text{year}$
with the peak of the CE events are observed during May and June.
An analysis of the total duration of CE events, showed that durations longer than 
5 minutes are very rare, with exceptions lasting over an hour.
We have detected enhancements over the total solar irradiance, without exceeding the
maximum values reported for other sites, with more favorable CE conditions.

\  
\  
CE: cloud enhancement cases one minute

ECE: extreme cloud enhancement cases one minute over TSI on horizontal plane

CEG: cloud enhancement groups, cases events with consecutive CE

$\text{GHI}_\text{i}$: measured one minute global horizontal irradiance

$\text{GHI}_\text{CSm}$: modelled clear sky one minute global horizontal irradiance

# Intro

The shortwave solar irradiance, reaching Earth's surface, is the main energy source
of the atmosphere and the biosphere, and drives and governs the climate [@Gray2010].  It
has direct practical application, in industries like energy and
agricultural production. The variability of its intensity, can cause difficulties in
predicting the yield, and designing the specifications of the appropriate equipment.
Significant portion of research has been focused on predicting the renewable energy production in a
fine timescale and in near real-time (for a review see @Inman2013; @Graabak2016).

A big aspect of this variability is the interaction of GHI with the clouds. In general,
clouds absorb part of the solar irradiance, but under certain conditions, can enhance
the total shortwave irradiance reaching the ground. This effect, can locally increase
the observed GHI, even higher than the expected clear sky irradiance
[see references therein].

Some of the proposed underling mechanisms of those enhancements, have been summarized
by @Gueymard2017, and include scattering on the edge of cumulus clouds or through
thin cirrus.  Further investigation with radiation transfer modeling methods and
observations have pointed as the prevailing mechanism, the forward Mie scattering
[@Pecenak2016; @Thuillier2013; @Yordanov2013; @Yordanov2015], through the clouds.
The overall phenomenon depends on different interactive factors, which includes cloud
thickness, constitution and type; and the relative position of the sun and the clouds
[@Gueymard2017; @Veerman2022].  As such, there are multiple contributing mechanisms
that are responsible for the observed irradiance enhancements.

On multiple sites, cloud enhancements have been reported, to be able to briefly exceed the
value of the solar constant, resulting of clear indexes above unit. A summary of
extreme enchantment cases (ECE) have been compiled by @Martins2022. There are also, some
practical implication of the cloud enchantments.  The intensity and duration of
enhancements can effect the efficiency and stability of photovoltaic power production
[@Lappalainen2020; @Jaervelae2020], and extreme enhancements cases, have the
potential to compromise the integrity of photovoltaic plants infrastructure
[@DoNascimento2019].  It has also been demonstrated, that these events can interfere
in the comparison of ground data and satellite observations [@Damiani2018]

<!--
Study of the mechanism has also been done with modeling of different cloud types and
atmospheric conditions.  @Gueymard2017; @Veerman2022
-->
<!--
Instruments respose ....
@Martins2022
-->

Methods of identification cloud enchantment events in the literature, usually include
the use of a simulated clear sky radiation as baseline, that is combined with an
appropriate threshold or some other statistical characteristics, and in some cases,
with visual methods with a sky cam [@Vamvakas2020; @Mol2023 and references therein].

<!-- our work -->
In this study, we evaluated the effects CE on the GHI by studding the occurrences, their intensity, and their duration in a
thirty-year period at the city of Thessaloniki.  We used modeled clear sky
irradiance, as a baseline to identify cloud enhancements.  We were able to determine
some trends of the phenomenon, it's Climatology, and some of their  general
characteristics. We weren't able to find a comparable study that provides trends for
similar long term dataset, as ours.  The recording of the GHI signal is the
mean of one-minute. Thus, the minimum resolution of CE events, in
this study is one minute.

<!-- Define CE and ECE -->
In the relative bibliography different definition have been used for these events,
some of the are summarized by @Gueymard2017.  Here, we defined as CE
events, the cases when the measured GHI at ground
level, exceeds the expected value under clear-sky conditions.  Similar, we define as
extreme cloud enhancement events (ECE), the cases when GHI, exceeds the Total Solar
Irradiance (TSI) on horizontal plane at ground level.  Although the duration of these
bursts varies, from instantaneous to several minutes, here we are constrained by the
recorded data, to one-minute steps.

<!--
Identification methods from bib

- clearness index
- modeled approaches
- algorithmic id
- AI
- Visual

Constrains of the site

- data
- instrument
- location
-->

# Data and methodology

## GHI data

The monitoring site is operating in the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m\ a.s.l.).
In this study we present data from the period
`r strftime(min(ST_daily$Date), "%d\\ %B %Y")` to
`r strftime(max(ST_daily$Date), "%d\\ %B %Y")`.
<!-- Instrument -->
The GHI data were measured with a Kipp\ & Zonen CM-21 pyranometer.  During the study
period, the pyranometer was independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, verifying that the stability of the
instrument’s sensitivity was better than $0.7\,\%$ relative to the initial
calibration by the manufacturer.
<!-- Measurements -->
For the acquisition of radiometric data, the signal of the pyranometer was sampled at
a rate of $1\,\text{Hz}$.  The mean and the standard deviation of these samples were
calculated and recorded for every minute.  The measurements were corrected for the
zero offset ("dark signal" in volts), which was calculated by averaging all
measurements recorded for a period of $3\,\text{h}$, before (morning) or after
(evening) the Sun reaches an elevation angle of $-10^\circ$.  The signal was
converted to irradiance using the ramped value of the instrument’s sensitivity
between subsequent calibrations.

<!-- Radiation data quality check -->
To further improve the quality of the irradiance data, a manual screening was
performed, in order to remove inconsistent and erroneous recordings that can occur
stochastically or systematically during the long operation of the instrument.
The manual screening was aided by a radiation data quality assurance procedure,
adjusted for the site, which was based on the methods of Long and Shi\ [@Long2006;
@Long2008a].  Thus, problematic recordings have been excluded from further
processing.  Although it is impossible to detect all false data, the large number of
available data, and the aggregation scheme we used, ensures the quality of the
radiation measurements used in this study.
<!-- Data selection -->
To preserve an unbiased representation of the data we applied a constraint, similar
the one used by @CastillejoCuberos2020. Where, for each valid hour of day, there must
exist at least 45 minutes of valid measurements, including nighttime, when near sunrise and
sunset. Days with less than 5 valid hours were rejected completely.
<!--# CastillejoCuberos2020 There must be at least 5 h of valid data (Tests 1–7),
otherwise all data for this day is flagged as invalid. There must be at least This
work valid data 45 min of valid data (Tests 1–7) for an hour to be flagged as valid,
otherwise all minute data for this hour is flagged as invalid. -->
Furthermore, due to the significant measurement uncertainty, when the Sun is near the
horizon, and due to some systematic obstructions by nearby buildings, we have
excluded all measurements with solar zenith angle (SZA) greater than 
$`r 90 - BIO_ELEVA`^\circ$.

## Cloud enhancement detection

To be able to detect the CE cases, we had to establish a baseline, above which we 
characterized each data point as an enhancement event, by estimating the occurring
over irradiance (OIR). The OIR, is defined as the irradiance difference of the
measured one-minute $\text{GHI}_i$ from the CE identification criterion
($\text{OIR}_i = \text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}.
To have an estimation, and a first insight on the phenomenon, we experimented with
two simple approaches for the reference.  The Haurwitz's model [@Haurwitz1945], which
is as simple clear sky model, and we had already adjusted and had good fit with our
data [@Natsis2023], and the TSI at the top of the atmosphere.  We have tested both
cases by using an appropriate relative threshold and/or an additional constant
offset.  The initial results, showed that we can detect a big portion of the CE
events. These results were helpful, as they are independent from unknown factors, and
helped us to establish some criteria to further improve the CE identification. It was
evident, by inspecting the daily plot of irradiance, that changes on the atmospheric
conditions introduced numerous false positive and false negative results.  To produce
a more accurate reference, we had to take into account more factors that effect the
clear sky radiation. So we used a radiation transfer
model in order to include the effects of ADO and the water vapors.

## Modeled clear Sky Irradiance

### Climatology of clear sky irradiance

We approximated the expected clear sky GHI by using the Libradtran radiation transfer
model [@Emde2016], a similar approach, was also used by @Vamvakas2020 for creating a
clear sky reference for studding cloud enhancement events.  Main factors responsible
for the attenuation of the broadband downward solar radiation in the atmosphere are
the aerosols and the water vapors.  Because of the lack of observational data for the
whole period, we used some long term climatological data to recreate the whole
period.  Fortunately, our site participates in the Aerosol Robotic Network (AERONET)
[@Giles2019; @Buis1998], as there is in operation a Cimel photometer since 2003,
collocated with the CM-21 pyranometer.  The mean monthly AOD on different wavelengths
are provided by AERONET, along with the equivalent water column height in the
atmosphere.

<!-- Libradtran options -->
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated on combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].

<!-- Lookup table -->
To create a look-up table, that aligns with our dataset, we applied some adjustments.
To account for the Sun's variability, in our one-minute GHI measurements, we adapted
each modeled value by scaling the model's input spectrum integral, to the
corresponding TSI, provided by NOAA [@Coddington2005].  Also,
we account for the effect of the Earth -- Sun distance on the irradiance, by using
the distance calculated by the Astropy [@AstropyCollaboration2022] software library.
As needed, we interpolate the resulting irradiances to the SZA of our
measurements.  For each period of the year, we used the appropriate atmospheric
profile (afglms or afglmw).  Finally, we calculated the clear sky irradiance value at
the horizontal plane.  Thus, we were able to emulate different atmospheric condition
and levels of atmospheric clearness for the climatological conditions of the site.
With this method, the modeled clear sky irradiances can be directly compared to each
measured one-minute value of GHI, for different conditions of atmospheric clearness.

### Long term change of clear sky irradiance

<!-- Aerosol change -->
The above clear sky reference values are not enough to describe the long term
variation, due to the changes of the atmospheric constituents, mainly AOD. As observed by
@Natsis2023, there is a long term brightening effect, over Thessaloniki for the studied
period. In order to better incorporate this change, we used, estimations of AOD change
derived from two sources. For 1993 -- 2005, @Kazadzis2007 have found a change of
$-3.8\,\%$ per year at AOD on $340\,\text{nm}$, using a Brewer spectrophotometer. For
2005 to 2023, we derived the trend from the available monthly climatological data of AERONET.

To create a unified trend for the long term change of the clear sky irradiance, we
simulated the values of AOD derived from those sources with Libradtran. For both
inputs, we
used the AOD at $500\,\text{nm}$, which was inferred by the available Ångström
coefficients. We choose the SZA of $55^\circ$ as a representative value for all the
runs.

According to @Kazadzis2007, in the period 1997 -- 2005, the mean AOD of
$340\,\text{nm}$ was calculated to $0.403$ with the use of a Brewer
spectrophotometer, with a linear trend of $-3.8±0.93\,\%$ per year.  This corresponds
to a change of $0.0153$ per year for the AOD of $340\,\text{nm}$ .  Using the alpha Ångström
coefficient of $\alpha = 1.6$, this translates to a change in Ångström beta
coefficient of $\beta=0.00272$ per year (or $\beta=0.084$ in 1997 and $\beta=0.059$
in 2005).  By interpolating the resulting clear sky irradiance, we have a trend of
$+0.21\,\%$ per year for the clear sky irradiance.  For the period 2005 -- 2005 we
used the mean monthly values from AERONET, with a similar simulation scheme. We
computed the trend of clear sky radiation for each month of the year, which gave us a
mean clear sky radiation trend of $+0.14\,\%$ per year. We applied this long term
change to the clear sky irradiance, in order to create a more realistic
representation of the irradiance for the whole study period
(Figure\nobreakspace{}\ref{fig:CS-change}).

```{r CS-change, echo=FALSE, fig.cap="Simulated long term change of the clear sky irradiance relevant to its climatological values, due to changed of AOD."}
knitr::include_graphics("../images/P_CS_change-1.png")
```

<!-- 1997 0.0839901
<!-- 2005 0.05496013
<!-- # Kazadzis2007 - Nine years of UV aerosol optical depth measurements at Thessaloniki, Greece
<!-- # From Stelios' paper 2007:
<!-- # AOD @340 from Brewer 086
<!-- # Period: 1997-2005
<!-- # The mean AOD for this period was calculated to 0.403 for the Brewer and 0.422 for the CIMEL.
<!-- # A linear regression on the Brewer deseasonalized data reveals a change of -3.8±0.93% per year.
<!-- # From the above data AOD @340 changes by 0.0153 per year (or 0.138 for the 9 years).
<!-- # Using Angstrom a= 1.6  this translates to a change in Angstrom β=0.00272 per year (or β=0.084 in 1997 and β=0.059 in 2005).
-->

## CE Criteria investigation

Our main focus was to positively identify over irradiance events related to CEs.  For
this reason we had to select an appropriate set of atmospheric parameterization, in
Libradtran, and define an relevant threshold for the CE identification.  The
selection of the atmospheric parameterization will provide us with a reference, that
has a good agreement with the clear sky conditions. Additional, with the
implementation of the long term change of AOD, as described above, it will be able to align
well with most of the natural variation of the physical quantity.  The selection of
adequate threshold, will allow us to have a robust method to distinguish the CE
cases, which will compensate for the limited accuracy of the input data, and the
natural variability of the atmosphere.

First, we evaluated the performance of the modelled clear radiation, for each of
the atmospheric level of clearness we have created as reference, in relation to the
measured GHI data, and choose as a representative of the clear sky radiation, the
case where
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
These values represent a typical atmosphere in Thessaloniki, with low load of
aerosols and humidity, which are the main factor that attenuate the GHI, excluding
clouds.  To create the upper limit of CE identification, we used a value relative to
clear sky irradiance
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}).
\begin{equation}
\text{CE} : \text{E}_\text{i} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.

In the process of defining the above details, we implemented an empirical method, by
manual inspection of the CE identification results, on selected days from the whole
dataset.  We used seven sets of days, with characteristics relevant to the efficiency
of the identification threshold.  These sets were random groups of about 20 to 30
days with the following characteristics:
(a) the strongest over irradiance CE events,
(b) the largest daily total over irradiance,
(c) absence of clouds (by implementing a clear sky identification algorithm as discussed in @Natsis2023),
(d) absence of clouds and absence of EC events,
(e) with at least $60\,\%$ of the day length without clouds and presence of EC events,
(h) random selected days and
(i) some manual selected days that were included during the manual inspection process.
Where it was needed, for some of the edge cases, we also used images from a sky-cam,
to further aid the decisions of the manual inspection.

<!--
\begin{equation}
\text{CE}_\text{Threshold} = \begin{cases}
`r C4_cs_ref_ratio`, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) + `r C4_lowcut_ratio` , & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4a}
\end{equation}
where: $\theta$ solar zenith angle.
-->
<!--
\begin{equation}
\text{CE} : \begin{cases}
 \text{GHI}_\text{i} > `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{mCSi}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\
\text{GHI}_\text{i} > \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{GHI}_\text{mCSi}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\
\text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ
\end{cases}\label{eq:CE4b}
\end{equation}
where: $\theta$ is the solar zenith angle, $\text{GHI}_\text{i}$ the measured
irradiance, and $\text{GHI}_\text{mCSi}$ the selected modelled clear sky irradiance.
-->
<!-- \begin{equation} -->
<!-- \text{CE} : \text{E}_\text{i} > \text{E}_\text{CSlim,i}, \text{where} \begin{cases} -->
<!--  \text{E}_\text{CSlim,i} = `r C4_cs_ref_ratio` \cdot \text{E}_\text{CSm,i}, & \text{$\theta \leq `r C4_lowcut_sza`^\circ$}\\ -->
<!-- \text{E}_\text{CSlim,i} = \left ({ `r C4_lowcut_ratio` + \frac{`r C4_cs_ref_ratio` - `r C4_lowcut_ratio`}{`r C4_lowcut_sza` - `r 90 - BIO_ELEVA`} \cdot (\theta- `r 90 - BIO_ELEVA`) } \right ) \cdot \text{E}_\text{CSm,i}, & \text{$ `r 90 - BIO_ELEVA`^\circ > \theta > `r C4_lowcut_sza`^\circ$}\\ -->
<!-- \text{Excluded measurements}, & \theta > `r 90 - BIO_ELEVA`^\circ -->
<!-- \end{cases}\label{eq:CE4} -->
<!-- \end{equation} -->
<!-- where: $\theta$ is the solar zenith angle, $\text{E}$ the measured -->
<!-- irradiance, $\text{E}_\text{CSm}$ the selected modelled clear sky irradiance, and $i$ each of the one-minute observation. -->

<!-- reasoning for cretria -->
<!-- We have to note, that the differentiation of the threshold factor was needed, because
<!-- of the high irradiance values we observed, early in the morning and late in the
<!-- afternoon. 
<!-- We have confirmed, by inspecting images from the sky cam, that we have
<!-- duration of elevated irradiance, either due to the clearness of the atmosphere, or some
<!-- interferences by reflections on nearby bright surfaces. Although, these cases may
<!-- mask some of the CE events. The actual CE events, produce higher irradiance and thus
<!-- are identified as such. As a side effect, the reported OIR will be slight
<!-- underestimated for those SZAs, but the overall contribution of those cases to the
<!-- total daily energy, is minimal due the low occurrences and the lower irradiances.
-->

<!--
- can follow for clear days
- this selection has some subjectivity
- the most important are the thresholds
- model have to be uniform on clear days
- the level is adjusted by the threshold 
- the effect of the sza uniform
-->

The definition of the CE events with this method, has a level of subjectivity. The
actual clear sky irradiance, is not known, and can only be approximated. Although,
because our main focus is to be able to identify the OIR created by the clouds, we
expect that we were able to detect all the major events of OIR due to CE.  Where,
some CE events, with very low OIR, may be not detected, these are few and their magnitude
will not effect our results.

<!-- extreme cases criterion -->
Another aspect of the CE events, that is often reported in the relative bibliography,
are cases of extreme cloud enhancement (ECE). These are cases of CE, where the
measured intensity of the irradiance, exceeds the equivalent TSI on the top of the
atmosphere, and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant, $r$ the Sun -- Earth distance, and $i$ each of the one-minute observation.

To give a better understanding of the CE identification procedure, we provided an
example of the CE identification for a selected day, in the Appendix
(Figure\nobreakspace{}\ref{fig:example-day}), where the use of the clear sky
reference and the identification threshold are shown, along with the relevant
measurements.  Similar, we provided an example plot of data for one year, where the
relation of the GHI and clear sky reference is shown, along with the OIR intensity
(Appendix Figure\nobreakspace{}\ref{fig:example-year}).

<!-- ```{r echo=F, include=T, warning=FALSE, message=FALSE} -->
<!-- ST_total -->
<!-- cat("\n - - - \n") -->
<!-- ST_daily[which.max(wattGLB.max), wattGLB.max] -->
<!-- round(ST_E_daily[which.max(GLB_diff.max), GLB_diff.max], 2) -->
<!-- round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1) -->
<!-- plot(ST_daily$Date, ST_daily$GLB_diff.max) -->
<!-- ``` -->

<!-- \FloatBarrier -->

# Results

Our dataset, after the data selection processing, consists of
`r ST_total$wattGLB.N` records of GHI measurements, of which
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ were identified as CE events and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ as ECE events.
The highest recorded GHI due to CE was
$`r round(ST_daily[which.max(wattGLB.max), wattGLB.max],1)`\,W/m^2$
on `r strftime(ST_daily[which.max(wattGLB.max), Date], "%d\\ %B %Y")`, and had
the maximum value of OIR at $`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 1)`\,W/m^2$.
<!-- The absolute strongest CE event had an OIR of 
<!-- $`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 2)`\,W/m^2$ on
<!-- `r strftime(ST_daily[which.max(GLB_diff.max), Date], "%d\\ %B %Y")`. -->
The relative stronger CE event was 
$`r round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1)`\,\%$ above the 
clear sky threshold, on 
`r strftime(ST_daily[which.max(GLB_ench.max), Date], "%d\\ %B %Y")`.

## Long term trends

We computed the daily trend of the mean OIR of CE <!--
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), --> using a first-order autoregressive
model with a lag of one day, with the 'maximum likelihood' fitting method
[@Gardner1980; @Jones1980] by implementing the function 'arima' from the library
'stats' of the R programming language [@RCT2023]. The trends are reported together
with the $2\sigma$ error.  We observed a negligible change of
$`r round(dailyCE[var == "GLB_diff.mean", Tmod_Estimate], 3)`\pm
`r round(2 * dailyCE[var == "GLB_diff.mean", Tmod_Std..Error], 3)`\,\text{W}/\text{m}^2/\text{year}$
on the mean OIR.
<!-- ```{r CEmeanDaily, echo=FALSE, fig.cap="Daily mean values and trend of the CE over irradiance."} -->
<!-- knitr::include_graphics("../images/P_daily_trend-1.png") -->
<!-- ``` -->
Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements.  Despite the
very weak statistical significance of the mean OIR trend, the annual number of CE
occurrences shows a steady increase of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$
(Figure\nobreakspace{}\ref{fig:P-energy-N}).
Subsequently, the mean annual energy of the CE events is increasing with a rate of
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kj}/\text{year}$
(Figure\ \ref{fig:P-energy-sum}), witch follows the trend of the number of CE
occurrences.  We have to note that the energy related to the CE events can not be
directly linked with the total energy balance of the atmosphere. The net solar
radiation of the region is not increased, but rather redistributed through the CE
events.  This also is depicted by the "instantaneous" values of irradiance (average
over one minute), that can exceed the equivalent clear sky irradiance to a
considerable level.

<!-- $`r round(yearlyCE[var == "GLB_diff.mean", slope], 2)`\pm -->
<!-- `r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 2)`\,W/m^2/y$ -->
<!-- (Figure\nobreakspace{}\ref{fig:P-energy-mean}), -->
<!-- but this value alone is not very useful due the intrinsic high variability of this -->
<!-- metric. -->

<!-- ```{r CEmeanDaily2, echo=FALSE, fig.cap="Trend of the daily excess energy due to CE"} -->
<!-- knitr::include_graphics("../images/daily-12.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Treds of the mean over irradiance per CE"} -->
<!-- knitr::include_graphics("../images/energy-3.png") -->
<!-- ``` -->

<!-- ```{r P-energy-mean, echo=FALSE, fig.cap="Trends of the mean OIR per CE."} -->
<!-- knitr::include_graphics("../images/P_energy-7.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Trend of number of CE cases"} -->
<!-- knitr::include_graphics("../images/energy-2.png") -->
<!-- ``` -->

```{r P-energy-N, echo=FALSE, fig.cap="Trend of yearly CE number of occurancies."}
knitr::include_graphics("../images/P_energy-6.png")
```

<!-- ```{r echo=FALSE, fig.cap="Trend of the yearly excess energy due to CE over irradiance"} -->
<!-- knitr::include_graphics("../images/energy-1.png") -->
<!-- ``` -->

```{r P-energy-sum, echo=FALSE, fig.cap="Trend of the yearly excess energy due to CE over irradiance"}
knitr::include_graphics("../images/P_energy-5.png")
```

<!-- ```{r P-energy-median, echo=FALSE, fig.cap="Trend of the yearly median over irradiance due to CE over irradiance"} -->
<!-- knitr::include_graphics("../images/P_energy-8.png") -->
<!-- ``` -->

<!-- \FloatBarrier -->

## Cloud enhancements climatology

Another interesting aspect of the CE cases occurrences, is their seasonal cycle. In
Figure\nobreakspace{}\ref{fig:relative-month-occurancies}, we have the monthly box plot
(whisker plot), where the values have been normalized by the highest median value,
that occurs on June.  Although the number of occurrences has a wide spread throughout
the study period, the most active period of CE occurrences is during May and June.
During the Winter (December -- February) the CE cases are about $25\,\%$ of the maximum.
For the rest months, the occurrences gradually ramp between the maximum and minimum.
This seasonality is a combinations of the clouds occurrences, and the relevant
position of the sun during each season of the year. Unfortunately, the lack of detailed data on cloud
formation, type and location is not allowing further analysis.

```{r relative-month-occurancies, echo=FALSE, fig.cap="Seasonal statistics of the number of CE occurancies for each month, normalized to the maximum occurancies on June. The box represents the values of the low $25\\,\\%$ percentile to $75\\,\\%$ percentile, where the thick horizontal line inside is the mean, the verical lines extend to the macimum and minimum vales, the dots are outlier values, and the rhombus is the mean."}
knitr::include_graphics("../images/clim_CE_month_norm_MAX_median_N-2.png")
```
<!--
```{r echo=FALSE, fig.cap="Distribution of the number of CE cases"}
knitr::include_graphics("../images/climCEmonth-1.png")
```
--->

<!-- ```{r echo=FALSE, fig.cap="Distribution of CE above reference"} -->
<!-- knitr::include_graphics("../images/relative_distributions-2.png") -->
<!-- ``` -->
<!--
```{r echo=FALSE, fig.cap="Distribution of CE above reference"}
knitr::include_graphics("../images/relative_distributions-4.png")
```
-->

The distribution of the OIR intensity, follows an exponential decline
Figure\nobreakspace{}\ref{fig:ovir-distribution}, where there is an inverse relation
between the CE events frequency and their intensity. This is expected, as the stronger
the CE events become, the rarer the particular atmospheric and sun conditions can
occur. This fact, is indicative to the magnitude and the probability of the expected OIRs
events over Thessaloniki. Similar distribution of CE occurrences have been reported
by @Vamvakas2020, where the magnitude of OIR were
higher due to the location of the city of Patras, $2.5^\circ$ closer to the equator.

```{r ovir-distribution, echo=FALSE, fig.cap="Distribution of CE over irradiance magnitude."}
knitr::include_graphics("../images/P-relative-distribution-diff-1.png")
```

<!-- whisky plots relative to GLB -->

<!-- set of yearly plots -->

<!-- \FloatBarrier -->

## Groups of cloud enhancement

In order to study the total duration of the CE events, we grouped the one
minute CE events to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can represent cloud enhancement conditions spanning a
duration of multiple minutes.  We have identified `r ST_G0[, .N]` groups of CE in the
whole period of study, where the
group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude, that
although we detected some CEG with durations longer than an hour, about $80\,\%$ of
the CEG have a duration of less than 5 minutes.

```{r ceg-duration-distribution, echo=FALSE, fig.cap="Distribution of CE groups duration in minutes."}
knitr::include_graphics("../images/groups-1.png")
```

The relation between the duration and the mean OIR intensity of the groups have also been
studied (Figure\nobreakspace{}\ref{fig:group-2d}). We observed that the GCE events
tend to have either long duration, or large intensity. Events with strong enhancement
and long duration are very rare. Similar results on this relation, have been
reported by @Zhang2018, on a study using a far higher sampling rate than ours.

```{r group-2d, echo=FALSE, fig.cap="Relation of the mean over irradiance and duration of GCE, where the color scale donotes the frequency of the respected events."}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

<!-- ```{r echo=FALSE, fig.cap="Relation of mean over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-4.png") -->
<!-- ``` -->

<!-- ```{r echo=FALSE, fig.cap="Relation of maximun over irradiance and CE group duration"} -->
<!-- knitr::include_graphics("../images/groups-8.png") -->
<!-- ``` -->

<!-- \FloatBarrier -->

## Extreme cloud enhancement events

An aspect of the CE events that is commonly reported and has some significance
on the solar energy production infrastructure are the extreme CE events (ECE). Where
solar irradiance exceeds the expected irradiance on top of the atmosphere
(Equation\nobreakspace{}\ref{eq:ECE}).
Analogous to Figure\nobreakspace{}\ref{fig:relative-month-occurancies} we have
computed the distribution of the number of occurrences of ECE events by month in
Figure\nobreakspace{}\ref{fig:relative-month-occurancies-ECE}. The most active period
for ECE events is in the spring and the start of the summer (March -- June), followed
by a period in the late fall (September and October).  This is probably related to 
the weather characteristics of these periods, where there are continuous alternations
between clear sky periods and clouds.

```{r relative-month-occurancies-ECE, echo=FALSE, fig.cap="Seasonal statistics of the number of ECE occurancies for each month, normalized to the maximum occurancies on March. The box represents the values of the low $25\\,\\%$ percentile to $75\\,\\%$ percentile, where the thick horizontal line inside is the mean, the verical lines extend to the macimum and minimum vales, the dots are outlier values, and the rhombus is the mean."}
knitr::include_graphics("../images/clim_ECE_month_norm_MAX_median_N-2.png")
```

<!-- ```{r echo=FALSE, fig.cap="Distribution of the number of ECE cases"} -->
<!-- knitr::include_graphics("../images/climECEmonth-1.png") -->
<!-- ``` -->

The distribution of the ECE events (Figure\nobreakspace{}\ref{fig:P-extreme-distribution}),
shows that there are rare cases where the OIR can exceed the TSI even more than
$400\,\text{W}/\text{m}^2$, with the $75\,\%$ of the cases to be below
$200\,\text{W}/\text{m}^2$. 
Those finds are in accordance with results form @Vamvakas2020, 
with the difference that the OIR values are lower for Thessaloniki.

```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE above clear sky threshold, for cases that are exceeding the TSI."}
knitr::include_graphics("../images/P-extreme-distribution-1.png")
```

<!-- ```{r echo=FALSE, fig.cap="Distribution of ECE Irradiance above 'TSI'"} -->
<!-- knitr::include_graphics("../images/extremedistributions-4.png") -->
<!-- ``` -->

<!-- -   SZA -->

<!-- ### Yearly plots relative to all GHI data. -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-1.png") -->
<!-- ``` -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-2.png") -->
<!-- ``` -->
<!-- ```{r echo=FALSE, fig.cap=""} -->
<!-- knitr::include_graphics("../images/rel_energy-3.png") -->
<!-- ``` -->

<!-- \FloatBarrier -->

# Discussion and conclusions

By creating a clear sky approximation of the GHI, which represents the long- and
short-term variation of the expected clear sky GHI, we were able to identify cases of 
CE events.
After analyzing the CE cases, we found an increase of 
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
with the
mean annual total energy of the CE events increasing with a rate of
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kj}/\text{year}$.
The most active season of CE events over Thessaloniki is concentrated on early
summer, on May and June.

The magnitude of the ECE events doesn't exceeds the values reported from sites with
more favourable conditions for the phenomenon. Although, the climatological
characteristic of the ECE events, showed that the most active months are spread on
half of the year (March -- June and September and October).
We found that CE conditions, can have a duration of more than an hour in rare
cases, with the bulk of the cases having a duration under 5 minutes.
Some of the characteristics of CE and ECE events we analysed,
have strong similarities with results
by @Vamvakas2020, for a city southern of Thessaloniki, with the analogues differences on the
 intensity of the solar radiation.

An interpretation of the CE trends, shows that the interaction of GHI with the
clouds,
through this 30 year period, is a dynamic phenomenon that's needs further investigation.
Although, to approach it, we need more long term observations of AOD and clouds
characteristics.

# Appendix {.unnumbered}

```{r example-day, echo=FALSE, fig.cap="Example of cloud indentification data for 2019-07-11. Where are denoted with red circles all the detected CE events, the cloud intearaction with the GHI (blue crosses), and the thesholds of CE and the threshold of ECE events with red and black line respectively."}
knitr::include_graphics("../images/example_days-18.png")
```

```{r example-year, echo=FALSE, fig.cap="Example of the relation of the measured GHI versus the reference clear sky irradiance for the year 2005, where the intesity of the detected over irradianse (OIR) is color coded."}
knitr::include_graphics("../images/example_years-13.png")
```

<!-- # References {.unnumbered} -->

<!--
Elsevier template

Please make sure that your manuscript follows the guidelines in the 
Guide for Authors of the relevant journal. It is not necessary to 
typeset your manuscript in exactly the same way as an article, 
unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see   <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references: @Feynman1963118 [@Dirac1953888].

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML and with `elsearticle-harv.bst` which is among provided style by `elsarticle` documentclass. Other available style are `elsarticle-num.bst` and `elsarticle-num-names.bst` — the first one can be used for the numbered scheme, second one for numbered with new options of natbib.sty.

You can sets extra options with `natbiboptions` variable in YAML header. Example 
```yaml
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. 
To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL

If `citation_package` is set to `default` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. Alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded
and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation:
$$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Here is another:
\begin{align}
  a^2+b^2=c^2.
\end{align}

Inline equations: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

Figure \ref{fig2} is generated using an R chunk.

```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig2}A meaningless scatterplot.", echo = FALSE}
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in Table \ref{tab1} for example.

```{r tab1, echo = TRUE}
knitr::kable(head(mtcars)[,1:4], 
    caption = "\\label{tab1}Caption centered above table"
)
```

-->