### 2024-06-06 GHI_enh_04_investigate.tex
Commit 2024-06-06 13:11

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{float}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-06-06}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
ST_daily ST_E_daily ST_E_daily_seas ST_extreme_daily
\end{verbatim}

\newpage
\FloatBarrier

\hypertarget{daily-trends}{%
\subsubsection{Daily Trends}\label{daily-trends}}

\newpage
\FloatBarrier

\hypertarget{group-stats}{%
\subsubsection{Group stats}\label{group-stats}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-1}

}

\caption{ - empty caption - }\label{fig:groups-1}
\end{figure}

{[}1{]} 0.8092946

\begin{verbatim}
Warning in scale_y_log10(): log-10 transformation introduced infinite values.
\end{verbatim}

\begin{verbatim}
Warning: Removed 23 rows containing missing values or values outside the scale
range (`geom_bar()`).
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-2}

}

\caption{ - empty caption - }\label{fig:groups-2}
\end{figure}

\begin{verbatim}
Warning: A numeric `legend.position` argument in `theme()` was deprecated in ggplot2 3.5.0.
i Please use the `legend.position.inside` argument of `theme()` instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-3}

}

\caption{ - empty caption - }\label{fig:groups-3}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-4}

}

\caption{ - empty caption - }\label{fig:groups-4}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-5}

}

\caption{ - empty caption - }\label{fig:groups-5}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-6}

}

\caption{ - empty caption - }\label{fig:groups-6}
\end{figure}

\hypertarget{zhang2018}{%
\subsection{@Zhang2018}\label{zhang2018}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-7}

}

\caption{ - empty caption - }\label{fig:groups-7}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/groups-8}

}

\caption{ - empty caption - }\label{fig:groups-8}
\end{figure}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-1}

}

\caption{ - empty caption - }\label{fig:P-groups-bin2d-1}
\end{figure}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{verbatim}
Warning: Removed 1 row containing missing values or values outside the scale
range (`geom_tile()`).
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-2}

}

\caption{ - empty caption - }\label{fig:P-groups-bin2d-2}
\end{figure}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-3}

}

\caption{ - empty caption - }\label{fig:P-groups-bin2d-3}
\end{figure}

\begin{verbatim}
Scale for x is already present.
Adding another scale for x, which will replace the existing scale.
\end{verbatim}

\begin{verbatim}
Warning: Removed 16 rows containing non-finite outside the scale range
(`stat_bin2d()`).
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-groups-bin2d-4}

}

\caption{ - empty caption - }\label{fig:P-groups-bin2d-4}
\end{figure}

\hypertarget{climatoogy}{%
\section{Climatoogy}\label{climatoogy}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-CE-N-doy-1}

}

\caption{ - empty caption - }\label{fig:clim-CE-N-doy}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCEmonth-1}

}

\caption{ - empty caption - }\label{fig:climCEmonth}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climCEmonth-MW-1}

}

\caption{ - empty caption - }\label{fig:climCEmonth-MW}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-CE-month-norm-MAX-N-1}

}

\caption{ - empty caption - }\label{fig:clim-CE-month-norm-MAX-N}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-CE-month-norm-MAX-N-MW-1}

}

\caption{ - empty caption - }\label{fig:clim-CE-month-norm-MAX-N-MW}
\end{figure}

\hypertarget{monthly-ce}{%
\subsection{Monthly CE}\label{monthly-ce}}

\begin{verbatim}
Warning: The `fun.y` argument of `stat_summary()` is deprecated as of ggplot2 3.3.0.
i Please use the `fun` argument instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-CE-month-norm-MAX-median-N-1}

}

\caption{ - empty caption - }\label{fig:clim-CE-month-norm-MAX-median-N}
\end{figure}

\hypertarget{monthly-weighted-ce}{%
\subsection{Monthly weighted CE}\label{monthly-weighted-ce}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-CE-month-norm-MAX-median-N-MW-1}

}

\caption{ - empty caption - }\label{fig:clim-CE-month-norm-MAX-median-N-MW}
\end{figure}

\hypertarget{monthly-extreme-ce-raw}{%
\subsection{Monthly extreme CE raw}\label{monthly-extreme-ce-raw}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climECEmonth-1}

}

\caption{ - empty caption - }\label{fig:climECEmonth}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/climECEmonth-MW-1}

}

\caption{ - empty caption - }\label{fig:climECEmonth-MW}
\end{figure}

\hypertarget{monthly-extreme-ce}{%
\subsection{Monthly extreme CE}\label{monthly-extreme-ce}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-ECE-month-norm-MAX-median-N-1}

}

\caption{ - empty caption - }\label{fig:clim-ECE-month-norm-MAX-median-N-1}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-ECE-month-norm-MAX-median-N-2}

}

\caption{ - empty caption - }\label{fig:clim-ECE-month-norm-MAX-median-N-2}
\end{figure}

\hypertarget{monthly-weighted-extreme-ce}{%
\subsection{Monthly weighted extreme CE}\label{monthly-weighted-extreme-ce}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/clim-ECE-month-norm-MAX-median-N-MW-1}

}

\caption{ - empty caption - }\label{fig:clim-ECE-month-norm-MAX-median-N-MW}
\end{figure}

\newpage
\FloatBarrier

\hypertarget{energy-contribution-of-enhancements}{%
\subsubsection{Energy contribution of enhancements}\label{energy-contribution-of-enhancements}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-1}

}

\caption{ - empty caption - }\label{fig:energy-1}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-2}

}

\caption{ - empty caption - }\label{fig:energy-2}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-3}

}

\caption{ - empty caption - }\label{fig:energy-3}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-4}

}

\caption{ - empty caption - }\label{fig:energy-4}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-5}

}

\caption{ - empty caption - }\label{fig:energy-5}
\end{figure}

\textbf{@Martins2022}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/energy-6}

}

\caption{ - empty caption - }\label{fig:energy-6}
\end{figure}

\newpage

\hypertarget{yearly-trends}{%
\subsection{Yearly trends}\label{yearly-trends}}

lmY: 116.8672 +/- 33.89196

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-energy-1}

}

\caption{ - empty caption - }\label{fig:P-energy-1}
\end{figure}

lmY: 45.58799 +/- 10.96835

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-energy-2}

}

\caption{ - empty caption - }\label{fig:P-energy-2}
\end{figure}

lmY: 0.062157 +/- 0.054182

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-energy-3}

}

\caption{ - empty caption - }\label{fig:P-energy-3}
\end{figure}

lmY: 0.053059 +/- 0.041079

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-energy-4}

}

\caption{ - empty caption - }\label{fig:P-energy-4}
\end{figure}

\newpage

\hypertarget{yearly-trends-month-weighted}{%
\subsection{Yearly trends month weighted}\label{yearly-trends-month-weighted}}

lmY: 95.37747 +/- 25.54612

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-MW-energy-1}

}

\caption{ - empty caption - }\label{fig:P-MW-energy-1}
\end{figure}

lmY: 37.11864 +/- 8.427722

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/P-MW-energy-2}

}

\caption{ - empty caption - }\label{fig:P-MW-energy-2}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel-energy-1}

}

\caption{ - empty caption - }\label{fig:rel-energy-1}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel-energy-2}

}

\caption{ - empty caption - }\label{fig:rel-energy-2}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel-energy-3}

}

\caption{ - empty caption - }\label{fig:rel-energy-3}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel-energy-4}

}

\caption{ - empty caption - }\label{fig:rel-energy-4}
\end{figure}

@Martins2022

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_04_investigate_files/figure-latex/rel-energy-5}

}

\caption{ - empty caption - }\label{fig:rel-energy-5}
\end{figure}

\textbf{END}

\begin{verbatim}
2024-06-06 09:19:49.0 athan@tyler GHI_enh_04_investigate.R 2.764330 mins
\end{verbatim}

\end{document}