### 2024-06-06 GHI_enh_06_sza.tex
Commit 2024-06-06 18:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{float}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-06-06}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:data.table':

    yearmon, yearqtr
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\newpage
\FloatBarrier

\hypertarget{sza-enhancements}{%
\subsubsection{SZA enhancements}\label{sza-enhancements}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza-1}

}

\caption{ - empty caption - }(\#fig:sza-1)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza-2}

}

\caption{ - empty caption - }(\#fig:sza-2)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza-3}

}

\caption{ - empty caption - }(\#fig:sza-3)
\end{figure}

\newpage
\FloatBarrier

\hypertarget{sza-enhancements-by-month}{%
\subsubsection{SZA enhancements by month}\label{sza-enhancements-by-month}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-1}

}

\caption{ - empty caption - }(\#fig:sza_month-1)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-2}

}

\caption{ - empty caption - }(\#fig:sza_month-2)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-3}

}

\caption{ - empty caption - }(\#fig:sza_month-3)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-4}

}

\caption{ - empty caption - }(\#fig:sza_month-4)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-5}

}

\caption{ - empty caption - }(\#fig:sza_month-5)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-6}

}

\caption{ - empty caption - }(\#fig:sza_month-6)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-7}

}

\caption{ - empty caption - }(\#fig:sza_month-7)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-8}

}

\caption{ - empty caption - }(\#fig:sza_month-8)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-9}

}

\caption{ - empty caption - }(\#fig:sza_month-9)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-10}

}

\caption{ - empty caption - }(\#fig:sza_month-10)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-11}

}

\caption{ - empty caption - }(\#fig:sza_month-11)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-12}

}

\caption{ - empty caption - }(\#fig:sza_month-12)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-13}

}

\caption{ - empty caption - }(\#fig:sza_month-13)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-14}

}

\caption{ - empty caption - }(\#fig:sza_month-14)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-15}

}

\caption{ - empty caption - }(\#fig:sza_month-15)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-16}

}

\caption{ - empty caption - }(\#fig:sza_month-16)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-17}

}

\caption{ - empty caption - }(\#fig:sza_month-17)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-18}

}

\caption{ - empty caption - }(\#fig:sza_month-18)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-19}

}

\caption{ - empty caption - }(\#fig:sza_month-19)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-20}

}

\caption{ - empty caption - }(\#fig:sza_month-20)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-21}

}

\caption{ - empty caption - }(\#fig:sza_month-21)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-22}

}

\caption{ - empty caption - }(\#fig:sza_month-22)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-23}

}

\caption{ - empty caption - }(\#fig:sza_month-23)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-24}

}

\caption{ - empty caption - }(\#fig:sza_month-24)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-25}

}

\caption{ - empty caption - }(\#fig:sza_month-25)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-26}

}

\caption{ - empty caption - }(\#fig:sza_month-26)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-27}

}

\caption{ - empty caption - }(\#fig:sza_month-27)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-28}

}

\caption{ - empty caption - }(\#fig:sza_month-28)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-29}

}

\caption{ - empty caption - }(\#fig:sza_month-29)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-30}

}

\caption{ - empty caption - }(\#fig:sza_month-30)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-31}

}

\caption{ - empty caption - }(\#fig:sza_month-31)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-32}

}

\caption{ - empty caption - }(\#fig:sza_month-32)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-33}

}

\caption{ - empty caption - }(\#fig:sza_month-33)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-34}

}

\caption{ - empty caption - }(\#fig:sza_month-34)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-35}

}

\caption{ - empty caption - }(\#fig:sza_month-35)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_month-36}

}

\caption{ - empty caption - }(\#fig:sza_month-36)
\end{figure}

\newpage
\FloatBarrier

\hypertarget{sza-heatmaps-of-doy-and-sza}{%
\subsubsection{SZA heatmaps of DOY and SZA}\label{sza-heatmaps-of-doy-and-sza}}

\begin{verbatim}
Warning: A numeric `legend.position` argument in `theme()` was deprecated in ggplot2 3.5.0.
i Please use the `legend.position.inside` argument of `theme()` instead.
This warning is displayed once every 8 hours.
Call `lifecycle::last_lifecycle_warnings()` to see where this warning was generated.
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_doy-1}

}

\caption{ - empty caption - }(\#fig:sza_doy-1)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_doy-2}

}

\caption{ - empty caption - }(\#fig:sza_doy-2)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_doy-3}

}

\caption{ - empty caption - }(\#fig:sza_doy-3)
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_06_sza_files/figure-latex/sza_doy-4}

}

\caption{ - empty caption - }(\#fig:sza_doy-4)
\end{figure}

\textbf{END}

\begin{verbatim}
2024-06-06 10:30:43.6 athan@tyler GHI_enh_06_sza.R 0.359110 mins
\end{verbatim}

\end{document}