### 2024-08-31 GHI_enh_read_aod_brewer.tex
Commit 2024-08-31 12:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{float}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-08-31}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
dinfo
στήλες: Year, DOY, cloudiness flag (0: no clouds), total ozone flag (1: measured, 3: climatological)

daod
AOD at wavelengths 290 - 363 nm with a step of 0.5 nm (each row corresponds to a row from dinfo)

dmin
Minute corresponding to each AOD values

dsza
SZA corresponding to each AOD values

dqflg
Εδώ έχω ένα quality flag που έβγαλα από το schic και τελικά δε χρησιμοποίησα

\begin{verbatim}
Ignore time of day
\end{verbatim}

\begin{verbatim}
Ignore SZA
\end{verbatim}

\begin{verbatim}
Drop empy columns
\end{verbatim}

\begin{verbatim}
Ignore negative AOD
\end{verbatim}

\begin{verbatim}
Ignore infinite AOD
\end{verbatim}

\begin{verbatim}
Drop 1997
\end{verbatim}

\begin{verbatim}
AOD_292 
      1 
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-1}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-1}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-2}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-2}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-3}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-3}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-4}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-4}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-5}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-5}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-6}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-6}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-7}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-7}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-8}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-8}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-9}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-9}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-10}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-10}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-11}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-11}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-12}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-12}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-13}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-13}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-14}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-14}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-15}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-15}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-16}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-16}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-17}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-17}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-18}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-18}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-19}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-19}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-20}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-20}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-21}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-21}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-22}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-22}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-23}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-23}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-24}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-24}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-25}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-25}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-26}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-26}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-27}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-27}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-28}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-28}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-29}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-29}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-30}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-30}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-31}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-31}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-32}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-32}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-33}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-33}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-34}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-34}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-35}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-35}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-36}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-36}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-37}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-37}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-38}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-38}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-39}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-39}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-40}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-40}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-41}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-41}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-42}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-42}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-43}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-43}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-44}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-44}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-45}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-45}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-46}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-46}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-47}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-47}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-48}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-48}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-49}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-49}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-50}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-50}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-51}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-51}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-52}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-52}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-53}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-53}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-54}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-54}
\end{figure}

\textbf{END}

\end{document}