### 2024-09-02 GHI_enh_read_aod_brewer.tex
Commit 2024-09-02 21:09

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{float}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-09-02}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
dinfo
στήλες: Year, DOY, cloudiness flag (0: no clouds), total ozone flag (1: measured, 3: climatological)

daod
AOD at wavelengths 290 - 363 nm with a step of 0.5 nm (each row corresponds to a row from dinfo)

dmin
Minute corresponding to each AOD values

dsza
SZA corresponding to each AOD values

dqflg
Εδώ έχω ένα quality flag που έβγαλα από το schic και τελικά δε χρησιμοποίησα

\begin{verbatim}
Ignore time of day
\end{verbatim}

\begin{verbatim}
Ignore SZA
\end{verbatim}

\begin{verbatim}
Drop empty columns
\end{verbatim}

\begin{verbatim}
Ignore negative AOD
\end{verbatim}

\begin{verbatim}
Ignore infinite AOD
\end{verbatim}

\begin{verbatim}
Drop year 1997
\end{verbatim}

\begin{verbatim}
Max N data for 1 
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-1}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-1}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-2}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-2}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-3}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-3}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-4}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-4}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-5}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-5}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-6}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-6}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-7}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-7}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-8}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-8}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-9}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-9}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-10}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-10}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-11}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-11}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-12}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-12}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-13}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-13}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-14}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-14}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-15}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-15}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-16}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-16}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-17}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-17}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-18}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-18}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-19}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-19}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-20}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-20}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-21}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-21}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-22}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-22}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-23}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-23}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-24}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-24}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-25}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-25}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-26}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-26}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-27}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-27}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-28}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-28}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-29}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-29}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-30}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-30}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-31}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-31}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-32}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-32}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-33}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-33}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-34}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-34}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-35}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-35}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-36}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-36}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-37}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-37}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-38}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-38}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-39}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-39}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-40}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-40}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-41}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-41}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-42}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-42}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-43}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-43}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-44}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-44}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-45}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-45}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-46}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-46}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-47}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-47}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-48}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-48}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-49}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-49}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-50}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-50}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-51}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-51}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-52}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-52}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-53}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-53}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-54}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-54}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-55}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-55}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-56}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-56}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-57}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-57}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-58}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-58}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-59}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-59}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-60}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-60}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-61}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-61}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-62}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-62}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-63}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-63}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-64}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-64}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-65}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-65}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-66}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-66}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-67}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-67}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-68}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-68}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-69}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-69}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-70}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-70}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-71}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-71}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-72}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-72}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-73}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-73}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-74}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-74}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-75}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-75}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-76}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-76}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-77}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-77}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-78}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-78}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-79}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-79}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-80}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-80}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-81}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-81}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-82}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-82}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-83}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-83}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-84}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-84}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-85}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-85}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-86}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-86}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-87}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-87}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-88}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-88}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-89}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-89}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-90}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-90}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-91}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-91}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-92}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-92}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-93}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-93}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-94}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-94}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-95}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-95}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-96}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-96}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-97}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-97}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-98}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-98}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-99}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-99}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-100}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-100}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-101}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-101}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-102}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-102}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-103}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-103}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-104}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-104}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-105}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-105}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-106}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-106}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-107}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-107}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-108}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-108}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-109}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-109}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-110}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-110}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-111}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-111}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-112}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-112}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-113}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-113}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-114}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-114}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-115}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-115}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-116}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-116}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-117}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-117}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-118}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-118}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-119}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-119}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-120}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-120}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-121}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-121}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-122}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-122}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-123}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-123}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-124}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-124}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-125}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-125}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-126}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-126}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-127}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-127}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-128}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-128}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-129}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-129}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-130}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-130}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-131}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-131}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-132}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-132}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-133}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-133}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-134}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-134}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-135}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-135}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-136}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-136}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-137}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-137}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-138}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-138}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-139}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-139}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-140}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-140}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-141}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-141}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-142}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-142}
\end{figure}
\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-143}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-143}
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1806}}@{}}
\caption{AOD change \%/y}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
slope\_pCpY
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
AOD
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
slope\_pCpY
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
AOD
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.2558 & AOD\_292 \\
0.0385 & AOD\_292.5 \\
0.04534 & AOD\_293 \\
0.06775 & AOD\_293.5 \\
0.05734 & AOD\_294 \\
0.0763 & AOD\_294.5 \\
0.0831 & AOD\_295 \\
0.08266 & AOD\_295.5 \\
0.1019 & AOD\_296 \\
0.09853 & AOD\_296.5 \\
0.09384 & AOD\_297 \\
0.08095 & AOD\_297.5 \\
0.09393 & AOD\_298 \\
0.08817 & AOD\_298.5 \\
0.0804 & AOD\_299 \\
0.07093 & AOD\_299.5 \\
0.06308 & AOD\_300 \\
0.05071 & AOD\_300.5 \\
0.05359 & AOD\_301 \\
0.04396 & AOD\_301.5 \\
0.03347 & AOD\_302 \\
0.02571 & AOD\_302.5 \\
0.02131 & AOD\_303 \\
0.01803 & AOD\_303.5 \\
0.01171 & AOD\_304 \\
0.01022 & AOD\_304.5 \\
0.005416 & AOD\_305 \\
0.0005074 & AOD\_305.5 \\
0.002836 & AOD\_306 \\
0.0001914 & AOD\_306.5 \\
0.001028 & AOD\_307 \\
0.0002915 & AOD\_307.5 \\
0.01176 & AOD\_308 \\
0.02345 & AOD\_308.5 \\
0.02512 & AOD\_309 \\
0.02872 & AOD\_309.5 \\
0.03807 & AOD\_310 \\
0.04033 & AOD\_310.5 \\
0.04404 & AOD\_311 \\
0.04732 & AOD\_311.5 \\
0.04988 & AOD\_312 \\
0.05461 & AOD\_312.5 \\
0.05661 & AOD\_313 \\
0.06141 & AOD\_313.5 \\
0.05901 & AOD\_314 \\
0.05361 & AOD\_314.5 \\
0.0447 & AOD\_315 \\
0.03122 & AOD\_315.5 \\
0.02879 & AOD\_316 \\
0.02784 & AOD\_316.5 \\
0.02667 & AOD\_317 \\
0.02556 & AOD\_317.5 \\
0.02617 & AOD\_318 \\
0.02414 & AOD\_318.5 \\
0.01576 & AOD\_319 \\
0.007456 & AOD\_319.5 \\
0.002617 & AOD\_320 \\
0.00727 & AOD\_320.5 \\
0.0097 & AOD\_321 \\
0.01868 & AOD\_321.5 \\
0.02508 & AOD\_322 \\
0.02554 & AOD\_322.5 \\
0.02652 & AOD\_323 \\
0.02711 & AOD\_323.5 \\
0.03271 & AOD\_324 \\
0.03674 & AOD\_324.5 \\
0.03564 & AOD\_325 \\
0.04207 & AOD\_325.5 \\
0.04356 & AOD\_326 \\
0.04334 & AOD\_326.5 \\
0.04177 & AOD\_327 \\
0.04679 & AOD\_327.5 \\
0.04267 & AOD\_328 \\
0.03719 & AOD\_328.5 \\
0.03491 & AOD\_329 \\
0.03328 & AOD\_329.5 \\
0.03067 & AOD\_330 \\
0.03602 & AOD\_330.5 \\
0.03564 & AOD\_331 \\
0.03208 & AOD\_331.5 \\
0.02064 & AOD\_332 \\
0.01897 & AOD\_332.5 \\
0.02294 & AOD\_333 \\
0.02197 & AOD\_333.5 \\
0.02225 & AOD\_334 \\
0.02282 & AOD\_334.5 \\
0.02284 & AOD\_335 \\
0.02381 & AOD\_335.5 \\
0.02606 & AOD\_336 \\
0.02517 & AOD\_336.5 \\
0.02722 & AOD\_337 \\
0.02877 & AOD\_337.5 \\
0.03161 & AOD\_338 \\
0.03294 & AOD\_338.5 \\
0.03393 & AOD\_339 \\
0.03063 & AOD\_339.5 \\
0.02624 & AOD\_340 \\
0.02005 & AOD\_340.5 \\
0.02407 & AOD\_341 \\
0.02232 & AOD\_341.5 \\
0.02444 & AOD\_342 \\
0.02502 & AOD\_342.5 \\
0.02645 & AOD\_343 \\
0.0247 & AOD\_343.5 \\
0.02531 & AOD\_344 \\
0.02526 & AOD\_344.5 \\
0.02374 & AOD\_345 \\
0.02159 & AOD\_345.5 \\
0.02034 & AOD\_346 \\
0.02048 & AOD\_346.5 \\
0.01923 & AOD\_347 \\
0.01956 & AOD\_347.5 \\
0.0214 & AOD\_348 \\
0.01945 & AOD\_348.5 \\
0.01227 & AOD\_349 \\
0.01165 & AOD\_349.5 \\
0.01124 & AOD\_350 \\
0.009548 & AOD\_350.5 \\
0.008288 & AOD\_351 \\
0.006796 & AOD\_351.5 \\
0.004561 & AOD\_352 \\
0.004843 & AOD\_352.5 \\
0.003653 & AOD\_353 \\
0.0007953 & AOD\_353.5 \\
0.001056 & AOD\_354 \\
0.004497 & AOD\_354.5 \\
0.003201 & AOD\_355 \\
0.009072 & AOD\_355.5 \\
0.00913 & AOD\_356 \\
0.00913 & AOD\_356.5 \\
0.009986 & AOD\_357 \\
0.01327 & AOD\_357.5 \\
0.01149 & AOD\_358 \\
0.008446 & AOD\_358.5 \\
0.004012 & AOD\_359 \\
0.00145 & AOD\_359.5 \\
0.001221 & AOD\_360 \\
0.00208 & AOD\_360.5 \\
0.003965 & AOD\_361 \\
0.01953 & AOD\_361.5 \\
0.06372 & AOD\_362 \\
0.066 & AOD\_362.5 \\
0.05968 & AOD\_363 \\
\end{longtable}

\begin{verbatim}
range 0.0001914191 0.2557606 
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_read_aod_brewer_files/figure-latex/unnamed-chunk-4-144}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4-144}
\end{figure}

\textbf{END}

\end{document}