### 2024-09-06 GHI_enh_daily_test_plots.tex
Commit 2024-09-06 10:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{float}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-09-06}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
Warning:

 ** Test is active!! ** 
\end{verbatim}

\FloatBarrier

\section{Plot some days with strong enhancement cases}\label{plot-some-days-with-strong-enhancement-cases}

\FloatBarrier

\subsection{Days with extreme cases day}\label{days-with-extreme-cases-day}

\FloatBarrier

\subsection{Days with strong enhancement day}\label{days-with-strong-enhancement-day}

\FloatBarrier

\subsection{Days with sunny day}\label{days-with-sunny-day}

\FloatBarrier

\subsection{Days with sunny enhancement day}\label{days-with-sunny-enhancement-day}

\FloatBarrier

\subsection{Days with cloudy day}\label{days-with-cloudy-day}

\FloatBarrier

\subsection{Days with manual test days}\label{days-with-manual-test-days}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-1}

}

\caption{ - empty caption - }\label{fig:example-days-1}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-2}

}

\caption{ - empty caption - }\label{fig:example-days-2}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-3}

}

\caption{ - empty caption - }\label{fig:example-days-3}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-4}

}

\caption{ - empty caption - }\label{fig:example-days-4}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-5}

}

\caption{ - empty caption - }\label{fig:example-days-5}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-6}

}

\caption{ - empty caption - }\label{fig:example-days-6}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-7}

}

\caption{ - empty caption - }\label{fig:example-days-7}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-8}

}

\caption{ - empty caption - }\label{fig:example-days-8}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-9}

}

\caption{ - empty caption - }\label{fig:example-days-9}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-10}

}

\caption{ - empty caption - }\label{fig:example-days-10}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-11}

}

\caption{ - empty caption - }\label{fig:example-days-11}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-12}

}

\caption{ - empty caption - }\label{fig:example-days-12}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-13}

}

\caption{ - empty caption - }\label{fig:example-days-13}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-14}

}

\caption{ - empty caption - }\label{fig:example-days-14}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-15}

}

\caption{ - empty caption - }\label{fig:example-days-15}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-16}

}

\caption{ - empty caption - }\label{fig:example-days-16}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-17}

}

\caption{ - empty caption - }\label{fig:example-days-17}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-18}

}

\caption{ - empty caption - }\label{fig:example-days-18}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-19}

}

\caption{ - empty caption - }\label{fig:example-days-19}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-20}

}

\caption{ - empty caption - }\label{fig:example-days-20}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-21}

}

\caption{ - empty caption - }\label{fig:example-days-21}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-22}

}

\caption{ - empty caption - }\label{fig:example-days-22}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-23}

}

\caption{ - empty caption - }\label{fig:example-days-23}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-24}

}

\caption{ - empty caption - }\label{fig:example-days-24}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-25}

}

\caption{ - empty caption - }\label{fig:example-days-25}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-26}

}

\caption{ - empty caption - }\label{fig:example-days-26}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-27}

}

\caption{ - empty caption - }\label{fig:example-days-27}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-28}

}

\caption{ - empty caption - }\label{fig:example-days-28}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-29}

}

\caption{ - empty caption - }\label{fig:example-days-29}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-30}

}

\caption{ - empty caption - }\label{fig:example-days-30}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-31}

}

\caption{ - empty caption - }\label{fig:example-days-31}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-32}

}

\caption{ - empty caption - }\label{fig:example-days-32}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-33}

}

\caption{ - empty caption - }\label{fig:example-days-33}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-34}

}

\caption{ - empty caption - }\label{fig:example-days-34}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-35}

}

\caption{ - empty caption - }\label{fig:example-days-35}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-36}

}

\caption{ - empty caption - }\label{fig:example-days-36}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-37}

}

\caption{ - empty caption - }\label{fig:example-days-37}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-38}

}

\caption{ - empty caption - }\label{fig:example-days-38}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-39}

}

\caption{ - empty caption - }\label{fig:example-days-39}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-40}

}

\caption{ - empty caption - }\label{fig:example-days-40}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-41}

}

\caption{ - empty caption - }\label{fig:example-days-41}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-42}

}

\caption{ - empty caption - }\label{fig:example-days-42}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-43}

}

\caption{ - empty caption - }\label{fig:example-days-43}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-44}

}

\caption{ - empty caption - }\label{fig:example-days-44}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-45}

}

\caption{ - empty caption - }\label{fig:example-days-45}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-46}

}

\caption{ - empty caption - }\label{fig:example-days-46}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-47}

}

\caption{ - empty caption - }\label{fig:example-days-47}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-48}

}

\caption{ - empty caption - }\label{fig:example-days-48}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-49}

}

\caption{ - empty caption - }\label{fig:example-days-49}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-50}

}

\caption{ - empty caption - }\label{fig:example-days-50}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-51}

}

\caption{ - empty caption - }\label{fig:example-days-51}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-52}

}

\caption{ - empty caption - }\label{fig:example-days-52}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-53}

}

\caption{ - empty caption - }\label{fig:example-days-53}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-54}

}

\caption{ - empty caption - }\label{fig:example-days-54}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-55}

}

\caption{ - empty caption - }\label{fig:example-days-55}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-56}

}

\caption{ - empty caption - }\label{fig:example-days-56}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-57}

}

\caption{ - empty caption - }\label{fig:example-days-57}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-58}

}

\caption{ - empty caption - }\label{fig:example-days-58}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-59}

}

\caption{ - empty caption - }\label{fig:example-days-59}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-60}

}

\caption{ - empty caption - }\label{fig:example-days-60}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-61}

}

\caption{ - empty caption - }\label{fig:example-days-61}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-62}

}

\caption{ - empty caption - }\label{fig:example-days-62}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-63}

}

\caption{ - empty caption - }\label{fig:example-days-63}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-64}

}

\caption{ - empty caption - }\label{fig:example-days-64}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-65}

}

\caption{ - empty caption - }\label{fig:example-days-65}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-66}

}

\caption{ - empty caption - }\label{fig:example-days-66}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-67}

}

\caption{ - empty caption - }\label{fig:example-days-67}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-68}

}

\caption{ - empty caption - }\label{fig:example-days-68}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-69}

}

\caption{ - empty caption - }\label{fig:example-days-69}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-70}

}

\caption{ - empty caption - }\label{fig:example-days-70}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-71}

}

\caption{ - empty caption - }\label{fig:example-days-71}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-72}

}

\caption{ - empty caption - }\label{fig:example-days-72}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-73}

}

\caption{ - empty caption - }\label{fig:example-days-73}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-74}

}

\caption{ - empty caption - }\label{fig:example-days-74}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-75}

}

\caption{ - empty caption - }\label{fig:example-days-75}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-76}

}

\caption{ - empty caption - }\label{fig:example-days-76}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-77}

}

\caption{ - empty caption - }\label{fig:example-days-77}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-78}

}

\caption{ - empty caption - }\label{fig:example-days-78}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-79}

}

\caption{ - empty caption - }\label{fig:example-days-79}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-80}

}

\caption{ - empty caption - }\label{fig:example-days-80}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-81}

}

\caption{ - empty caption - }\label{fig:example-days-81}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-82}

}

\caption{ - empty caption - }\label{fig:example-days-82}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-83}

}

\caption{ - empty caption - }\label{fig:example-days-83}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-84}

}

\caption{ - empty caption - }\label{fig:example-days-84}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-85}

}

\caption{ - empty caption - }\label{fig:example-days-85}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-86}

}

\caption{ - empty caption - }\label{fig:example-days-86}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-87}

}

\caption{ - empty caption - }\label{fig:example-days-87}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-88}

}

\caption{ - empty caption - }\label{fig:example-days-88}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-89}

}

\caption{ - empty caption - }\label{fig:example-days-89}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-90}

}

\caption{ - empty caption - }\label{fig:example-days-90}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-91}

}

\caption{ - empty caption - }\label{fig:example-days-91}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-92}

}

\caption{ - empty caption - }\label{fig:example-days-92}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-93}

}

\caption{ - empty caption - }\label{fig:example-days-93}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-94}

}

\caption{ - empty caption - }\label{fig:example-days-94}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-95}

}

\caption{ - empty caption - }\label{fig:example-days-95}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-96}

}

\caption{ - empty caption - }\label{fig:example-days-96}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-97}

}

\caption{ - empty caption - }\label{fig:example-days-97}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-98}

}

\caption{ - empty caption - }\label{fig:example-days-98}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-99}

}

\caption{ - empty caption - }\label{fig:example-days-99}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-100}

}

\caption{ - empty caption - }\label{fig:example-days-100}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-101}

}

\caption{ - empty caption - }\label{fig:example-days-101}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-102}

}

\caption{ - empty caption - }\label{fig:example-days-102}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-103}

}

\caption{ - empty caption - }\label{fig:example-days-103}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-104}

}

\caption{ - empty caption - }\label{fig:example-days-104}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-105}

}

\caption{ - empty caption - }\label{fig:example-days-105}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-106}

}

\caption{ - empty caption - }\label{fig:example-days-106}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-107}

}

\caption{ - empty caption - }\label{fig:example-days-107}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-108}

}

\caption{ - empty caption - }\label{fig:example-days-108}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-109}

}

\caption{ - empty caption - }\label{fig:example-days-109}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-110}

}

\caption{ - empty caption - }\label{fig:example-days-110}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-111}

}

\caption{ - empty caption - }\label{fig:example-days-111}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-112}

}

\caption{ - empty caption - }\label{fig:example-days-112}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-113}

}

\caption{ - empty caption - }\label{fig:example-days-113}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-114}

}

\caption{ - empty caption - }\label{fig:example-days-114}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-115}

}

\caption{ - empty caption - }\label{fig:example-days-115}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-116}

}

\caption{ - empty caption - }\label{fig:example-days-116}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-117}

}

\caption{ - empty caption - }\label{fig:example-days-117}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-118}

}

\caption{ - empty caption - }\label{fig:example-days-118}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-119}

}

\caption{ - empty caption - }\label{fig:example-days-119}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-120}

}

\caption{ - empty caption - }\label{fig:example-days-120}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-121}

}

\caption{ - empty caption - }\label{fig:example-days-121}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-122}

}

\caption{ - empty caption - }\label{fig:example-days-122}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-123}

}

\caption{ - empty caption - }\label{fig:example-days-123}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-124}

}

\caption{ - empty caption - }\label{fig:example-days-124}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-125}

}

\caption{ - empty caption - }\label{fig:example-days-125}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-126}

}

\caption{ - empty caption - }\label{fig:example-days-126}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-127}

}

\caption{ - empty caption - }\label{fig:example-days-127}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-128}

}

\caption{ - empty caption - }\label{fig:example-days-128}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-129}

}

\caption{ - empty caption - }\label{fig:example-days-129}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-130}

}

\caption{ - empty caption - }\label{fig:example-days-130}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-131}

}

\caption{ - empty caption - }\label{fig:example-days-131}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-132}

}

\caption{ - empty caption - }\label{fig:example-days-132}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-133}

}

\caption{ - empty caption - }\label{fig:example-days-133}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-134}

}

\caption{ - empty caption - }\label{fig:example-days-134}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-135}

}

\caption{ - empty caption - }\label{fig:example-days-135}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-136}

}

\caption{ - empty caption - }\label{fig:example-days-136}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-137}

}

\caption{ - empty caption - }\label{fig:example-days-137}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-138}

}

\caption{ - empty caption - }\label{fig:example-days-138}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-139}

}

\caption{ - empty caption - }\label{fig:example-days-139}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-140}

}

\caption{ - empty caption - }\label{fig:example-days-140}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-141}

}

\caption{ - empty caption - }\label{fig:example-days-141}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-142}

}

\caption{ - empty caption - }\label{fig:example-days-142}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-143}

}

\caption{ - empty caption - }\label{fig:example-days-143}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-144}

}

\caption{ - empty caption - }\label{fig:example-days-144}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-145}

}

\caption{ - empty caption - }\label{fig:example-days-145}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-146}

}

\caption{ - empty caption - }\label{fig:example-days-146}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-147}

}

\caption{ - empty caption - }\label{fig:example-days-147}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-148}

}

\caption{ - empty caption - }\label{fig:example-days-148}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-149}

}

\caption{ - empty caption - }\label{fig:example-days-149}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-150}

}

\caption{ - empty caption - }\label{fig:example-days-150}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-151}

}

\caption{ - empty caption - }\label{fig:example-days-151}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-152}

}

\caption{ - empty caption - }\label{fig:example-days-152}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-153}

}

\caption{ - empty caption - }\label{fig:example-days-153}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-154}

}

\caption{ - empty caption - }\label{fig:example-days-154}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-155}

}

\caption{ - empty caption - }\label{fig:example-days-155}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-156}

}

\caption{ - empty caption - }\label{fig:example-days-156}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-157}

}

\caption{ - empty caption - }\label{fig:example-days-157}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-158}

}

\caption{ - empty caption - }\label{fig:example-days-158}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-159}

}

\caption{ - empty caption - }\label{fig:example-days-159}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-160}

}

\caption{ - empty caption - }\label{fig:example-days-160}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-161}

}

\caption{ - empty caption - }\label{fig:example-days-161}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-162}

}

\caption{ - empty caption - }\label{fig:example-days-162}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-163}

}

\caption{ - empty caption - }\label{fig:example-days-163}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-164}

}

\caption{ - empty caption - }\label{fig:example-days-164}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-165}

}

\caption{ - empty caption - }\label{fig:example-days-165}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-166}

}

\caption{ - empty caption - }\label{fig:example-days-166}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-167}

}

\caption{ - empty caption - }\label{fig:example-days-167}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-168}

}

\caption{ - empty caption - }\label{fig:example-days-168}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-169}

}

\caption{ - empty caption - }\label{fig:example-days-169}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-170}

}

\caption{ - empty caption - }\label{fig:example-days-170}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-171}

}

\caption{ - empty caption - }\label{fig:example-days-171}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-172}

}

\caption{ - empty caption - }\label{fig:example-days-172}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-173}

}

\caption{ - empty caption - }\label{fig:example-days-173}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-174}

}

\caption{ - empty caption - }\label{fig:example-days-174}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-175}

}

\caption{ - empty caption - }\label{fig:example-days-175}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-176}

}

\caption{ - empty caption - }\label{fig:example-days-176}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-177}

}

\caption{ - empty caption - }\label{fig:example-days-177}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-178}

}

\caption{ - empty caption - }\label{fig:example-days-178}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-179}

}

\caption{ - empty caption - }\label{fig:example-days-179}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-180}

}

\caption{ - empty caption - }\label{fig:example-days-180}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-181}

}

\caption{ - empty caption - }\label{fig:example-days-181}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-182}

}

\caption{ - empty caption - }\label{fig:example-days-182}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-183}

}

\caption{ - empty caption - }\label{fig:example-days-183}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-184}

}

\caption{ - empty caption - }\label{fig:example-days-184}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-185}

}

\caption{ - empty caption - }\label{fig:example-days-185}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-186}

}

\caption{ - empty caption - }\label{fig:example-days-186}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-187}

}

\caption{ - empty caption - }\label{fig:example-days-187}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-188}

}

\caption{ - empty caption - }\label{fig:example-days-188}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-189}

}

\caption{ - empty caption - }\label{fig:example-days-189}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-190}

}

\caption{ - empty caption - }\label{fig:example-days-190}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-191}

}

\caption{ - empty caption - }\label{fig:example-days-191}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-192}

}

\caption{ - empty caption - }\label{fig:example-days-192}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-193}

}

\caption{ - empty caption - }\label{fig:example-days-193}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-194}

}

\caption{ - empty caption - }\label{fig:example-days-194}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-195}

}

\caption{ - empty caption - }\label{fig:example-days-195}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-196}

}

\caption{ - empty caption - }\label{fig:example-days-196}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-197}

}

\caption{ - empty caption - }\label{fig:example-days-197}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-198}

}

\caption{ - empty caption - }\label{fig:example-days-198}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-199}

}

\caption{ - empty caption - }\label{fig:example-days-199}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-200}

}

\caption{ - empty caption - }\label{fig:example-days-200}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-201}

}

\caption{ - empty caption - }\label{fig:example-days-201}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-202}

}

\caption{ - empty caption - }\label{fig:example-days-202}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-203}

}

\caption{ - empty caption - }\label{fig:example-days-203}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-204}

}

\caption{ - empty caption - }\label{fig:example-days-204}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-205}

}

\caption{ - empty caption - }\label{fig:example-days-205}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-206}

}

\caption{ - empty caption - }\label{fig:example-days-206}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-207}

}

\caption{ - empty caption - }\label{fig:example-days-207}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-208}

}

\caption{ - empty caption - }\label{fig:example-days-208}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-209}

}

\caption{ - empty caption - }\label{fig:example-days-209}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-210}

}

\caption{ - empty caption - }\label{fig:example-days-210}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-211}

}

\caption{ - empty caption - }\label{fig:example-days-211}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-212}

}

\caption{ - empty caption - }\label{fig:example-days-212}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-213}

}

\caption{ - empty caption - }\label{fig:example-days-213}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-214}

}

\caption{ - empty caption - }\label{fig:example-days-214}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-215}

}

\caption{ - empty caption - }\label{fig:example-days-215}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-216}

}

\caption{ - empty caption - }\label{fig:example-days-216}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-217}

}

\caption{ - empty caption - }\label{fig:example-days-217}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-218}

}

\caption{ - empty caption - }\label{fig:example-days-218}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-219}

}

\caption{ - empty caption - }\label{fig:example-days-219}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-220}

}

\caption{ - empty caption - }\label{fig:example-days-220}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-221}

}

\caption{ - empty caption - }\label{fig:example-days-221}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-222}

}

\caption{ - empty caption - }\label{fig:example-days-222}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-223}

}

\caption{ - empty caption - }\label{fig:example-days-223}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-224}

}

\caption{ - empty caption - }\label{fig:example-days-224}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-225}

}

\caption{ - empty caption - }\label{fig:example-days-225}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-226}

}

\caption{ - empty caption - }\label{fig:example-days-226}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-227}

}

\caption{ - empty caption - }\label{fig:example-days-227}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-228}

}

\caption{ - empty caption - }\label{fig:example-days-228}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-229}

}

\caption{ - empty caption - }\label{fig:example-days-229}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-230}

}

\caption{ - empty caption - }\label{fig:example-days-230}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-231}

}

\caption{ - empty caption - }\label{fig:example-days-231}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-232}

}

\caption{ - empty caption - }\label{fig:example-days-232}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-233}

}

\caption{ - empty caption - }\label{fig:example-days-233}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-234}

}

\caption{ - empty caption - }\label{fig:example-days-234}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-235}

}

\caption{ - empty caption - }\label{fig:example-days-235}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-236}

}

\caption{ - empty caption - }\label{fig:example-days-236}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-237}

}

\caption{ - empty caption - }\label{fig:example-days-237}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-238}

}

\caption{ - empty caption - }\label{fig:example-days-238}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-239}

}

\caption{ - empty caption - }\label{fig:example-days-239}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-240}

}

\caption{ - empty caption - }\label{fig:example-days-240}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-241}

}

\caption{ - empty caption - }\label{fig:example-days-241}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-242}

}

\caption{ - empty caption - }\label{fig:example-days-242}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-243}

}

\caption{ - empty caption - }\label{fig:example-days-243}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-244}

}

\caption{ - empty caption - }\label{fig:example-days-244}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-245}

}

\caption{ - empty caption - }\label{fig:example-days-245}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-246}

}

\caption{ - empty caption - }\label{fig:example-days-246}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-247}

}

\caption{ - empty caption - }\label{fig:example-days-247}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-248}

}

\caption{ - empty caption - }\label{fig:example-days-248}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-249}

}

\caption{ - empty caption - }\label{fig:example-days-249}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-250}

}

\caption{ - empty caption - }\label{fig:example-days-250}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-251}

}

\caption{ - empty caption - }\label{fig:example-days-251}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-252}

}

\caption{ - empty caption - }\label{fig:example-days-252}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-253}

}

\caption{ - empty caption - }\label{fig:example-days-253}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-254}

}

\caption{ - empty caption - }\label{fig:example-days-254}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-255}

}

\caption{ - empty caption - }\label{fig:example-days-255}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-256}

}

\caption{ - empty caption - }\label{fig:example-days-256}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-257}

}

\caption{ - empty caption - }\label{fig:example-days-257}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-258}

}

\caption{ - empty caption - }\label{fig:example-days-258}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-259}

}

\caption{ - empty caption - }\label{fig:example-days-259}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-260}

}

\caption{ - empty caption - }\label{fig:example-days-260}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-261}

}

\caption{ - empty caption - }\label{fig:example-days-261}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-262}

}

\caption{ - empty caption - }\label{fig:example-days-262}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-263}

}

\caption{ - empty caption - }\label{fig:example-days-263}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-264}

}

\caption{ - empty caption - }\label{fig:example-days-264}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-265}

}

\caption{ - empty caption - }\label{fig:example-days-265}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-266}

}

\caption{ - empty caption - }\label{fig:example-days-266}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-267}

}

\caption{ - empty caption - }\label{fig:example-days-267}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-268}

}

\caption{ - empty caption - }\label{fig:example-days-268}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-269}

}

\caption{ - empty caption - }\label{fig:example-days-269}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-270}

}

\caption{ - empty caption - }\label{fig:example-days-270}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-271}

}

\caption{ - empty caption - }\label{fig:example-days-271}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-272}

}

\caption{ - empty caption - }\label{fig:example-days-272}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-273}

}

\caption{ - empty caption - }\label{fig:example-days-273}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-274}

}

\caption{ - empty caption - }\label{fig:example-days-274}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-275}

}

\caption{ - empty caption - }\label{fig:example-days-275}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-276}

}

\caption{ - empty caption - }\label{fig:example-days-276}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-277}

}

\caption{ - empty caption - }\label{fig:example-days-277}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-278}

}

\caption{ - empty caption - }\label{fig:example-days-278}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-279}

}

\caption{ - empty caption - }\label{fig:example-days-279}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-280}

}

\caption{ - empty caption - }\label{fig:example-days-280}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-281}

}

\caption{ - empty caption - }\label{fig:example-days-281}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-282}

}

\caption{ - empty caption - }\label{fig:example-days-282}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-283}

}

\caption{ - empty caption - }\label{fig:example-days-283}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-284}

}

\caption{ - empty caption - }\label{fig:example-days-284}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-285}

}

\caption{ - empty caption - }\label{fig:example-days-285}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-286}

}

\caption{ - empty caption - }\label{fig:example-days-286}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-287}

}

\caption{ - empty caption - }\label{fig:example-days-287}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-288}

}

\caption{ - empty caption - }\label{fig:example-days-288}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-289}

}

\caption{ - empty caption - }\label{fig:example-days-289}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-290}

}

\caption{ - empty caption - }\label{fig:example-days-290}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-291}

}

\caption{ - empty caption - }\label{fig:example-days-291}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-292}

}

\caption{ - empty caption - }\label{fig:example-days-292}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-293}

}

\caption{ - empty caption - }\label{fig:example-days-293}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-294}

}

\caption{ - empty caption - }\label{fig:example-days-294}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-295}

}

\caption{ - empty caption - }\label{fig:example-days-295}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-296}

}

\caption{ - empty caption - }\label{fig:example-days-296}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-297}

}

\caption{ - empty caption - }\label{fig:example-days-297}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-298}

}

\caption{ - empty caption - }\label{fig:example-days-298}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-299}

}

\caption{ - empty caption - }\label{fig:example-days-299}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-300}

}

\caption{ - empty caption - }\label{fig:example-days-300}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-301}

}

\caption{ - empty caption - }\label{fig:example-days-301}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-302}

}

\caption{ - empty caption - }\label{fig:example-days-302}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-303}

}

\caption{ - empty caption - }\label{fig:example-days-303}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-304}

}

\caption{ - empty caption - }\label{fig:example-days-304}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-305}

}

\caption{ - empty caption - }\label{fig:example-days-305}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-306}

}

\caption{ - empty caption - }\label{fig:example-days-306}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-307}

}

\caption{ - empty caption - }\label{fig:example-days-307}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-308}

}

\caption{ - empty caption - }\label{fig:example-days-308}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-309}

}

\caption{ - empty caption - }\label{fig:example-days-309}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-310}

}

\caption{ - empty caption - }\label{fig:example-days-310}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-311}

}

\caption{ - empty caption - }\label{fig:example-days-311}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-312}

}

\caption{ - empty caption - }\label{fig:example-days-312}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-313}

}

\caption{ - empty caption - }\label{fig:example-days-313}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-314}

}

\caption{ - empty caption - }\label{fig:example-days-314}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-315}

}

\caption{ - empty caption - }\label{fig:example-days-315}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-316}

}

\caption{ - empty caption - }\label{fig:example-days-316}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-317}

}

\caption{ - empty caption - }\label{fig:example-days-317}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-318}

}

\caption{ - empty caption - }\label{fig:example-days-318}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-319}

}

\caption{ - empty caption - }\label{fig:example-days-319}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-320}

}

\caption{ - empty caption - }\label{fig:example-days-320}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-321}

}

\caption{ - empty caption - }\label{fig:example-days-321}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-322}

}

\caption{ - empty caption - }\label{fig:example-days-322}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-323}

}

\caption{ - empty caption - }\label{fig:example-days-323}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-324}

}

\caption{ - empty caption - }\label{fig:example-days-324}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-325}

}

\caption{ - empty caption - }\label{fig:example-days-325}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-326}

}

\caption{ - empty caption - }\label{fig:example-days-326}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-327}

}

\caption{ - empty caption - }\label{fig:example-days-327}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-328}

}

\caption{ - empty caption - }\label{fig:example-days-328}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-329}

}

\caption{ - empty caption - }\label{fig:example-days-329}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-330}

}

\caption{ - empty caption - }\label{fig:example-days-330}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-331}

}

\caption{ - empty caption - }\label{fig:example-days-331}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-332}

}

\caption{ - empty caption - }\label{fig:example-days-332}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-333}

}

\caption{ - empty caption - }\label{fig:example-days-333}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-334}

}

\caption{ - empty caption - }\label{fig:example-days-334}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-335}

}

\caption{ - empty caption - }\label{fig:example-days-335}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-336}

}

\caption{ - empty caption - }\label{fig:example-days-336}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-337}

}

\caption{ - empty caption - }\label{fig:example-days-337}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-338}

}

\caption{ - empty caption - }\label{fig:example-days-338}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-339}

}

\caption{ - empty caption - }\label{fig:example-days-339}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-340}

}

\caption{ - empty caption - }\label{fig:example-days-340}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-341}

}

\caption{ - empty caption - }\label{fig:example-days-341}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-342}

}

\caption{ - empty caption - }\label{fig:example-days-342}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-343}

}

\caption{ - empty caption - }\label{fig:example-days-343}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-344}

}

\caption{ - empty caption - }\label{fig:example-days-344}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-345}

}

\caption{ - empty caption - }\label{fig:example-days-345}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-346}

}

\caption{ - empty caption - }\label{fig:example-days-346}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-347}

}

\caption{ - empty caption - }\label{fig:example-days-347}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-348}

}

\caption{ - empty caption - }\label{fig:example-days-348}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-349}

}

\caption{ - empty caption - }\label{fig:example-days-349}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-350}

}

\caption{ - empty caption - }\label{fig:example-days-350}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-351}

}

\caption{ - empty caption - }\label{fig:example-days-351}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-352}

}

\caption{ - empty caption - }\label{fig:example-days-352}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-353}

}

\caption{ - empty caption - }\label{fig:example-days-353}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-354}

}

\caption{ - empty caption - }\label{fig:example-days-354}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-355}

}

\caption{ - empty caption - }\label{fig:example-days-355}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-356}

}

\caption{ - empty caption - }\label{fig:example-days-356}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-357}

}

\caption{ - empty caption - }\label{fig:example-days-357}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-358}

}

\caption{ - empty caption - }\label{fig:example-days-358}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-359}

}

\caption{ - empty caption - }\label{fig:example-days-359}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-360}

}

\caption{ - empty caption - }\label{fig:example-days-360}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-361}

}

\caption{ - empty caption - }\label{fig:example-days-361}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-362}

}

\caption{ - empty caption - }\label{fig:example-days-362}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-363}

}

\caption{ - empty caption - }\label{fig:example-days-363}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-364}

}

\caption{ - empty caption - }\label{fig:example-days-364}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-365}

}

\caption{ - empty caption - }\label{fig:example-days-365}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-366}

}

\caption{ - empty caption - }\label{fig:example-days-366}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-367}

}

\caption{ - empty caption - }\label{fig:example-days-367}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-368}

}

\caption{ - empty caption - }\label{fig:example-days-368}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-369}

}

\caption{ - empty caption - }\label{fig:example-days-369}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-370}

}

\caption{ - empty caption - }\label{fig:example-days-370}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-371}

}

\caption{ - empty caption - }\label{fig:example-days-371}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-372}

}

\caption{ - empty caption - }\label{fig:example-days-372}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-373}

}

\caption{ - empty caption - }\label{fig:example-days-373}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-374}

}

\caption{ - empty caption - }\label{fig:example-days-374}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-375}

}

\caption{ - empty caption - }\label{fig:example-days-375}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-376}

}

\caption{ - empty caption - }\label{fig:example-days-376}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-377}

}

\caption{ - empty caption - }\label{fig:example-days-377}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-378}

}

\caption{ - empty caption - }\label{fig:example-days-378}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-379}

}

\caption{ - empty caption - }\label{fig:example-days-379}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-380}

}

\caption{ - empty caption - }\label{fig:example-days-380}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-381}

}

\caption{ - empty caption - }\label{fig:example-days-381}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-382}

}

\caption{ - empty caption - }\label{fig:example-days-382}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-383}

}

\caption{ - empty caption - }\label{fig:example-days-383}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-384}

}

\caption{ - empty caption - }\label{fig:example-days-384}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-385}

}

\caption{ - empty caption - }\label{fig:example-days-385}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-386}

}

\caption{ - empty caption - }\label{fig:example-days-386}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-387}

}

\caption{ - empty caption - }\label{fig:example-days-387}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-388}

}

\caption{ - empty caption - }\label{fig:example-days-388}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-389}

}

\caption{ - empty caption - }\label{fig:example-days-389}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-390}

}

\caption{ - empty caption - }\label{fig:example-days-390}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-391}

}

\caption{ - empty caption - }\label{fig:example-days-391}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-392}

}

\caption{ - empty caption - }\label{fig:example-days-392}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-393}

}

\caption{ - empty caption - }\label{fig:example-days-393}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-394}

}

\caption{ - empty caption - }\label{fig:example-days-394}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-395}

}

\caption{ - empty caption - }\label{fig:example-days-395}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-396}

}

\caption{ - empty caption - }\label{fig:example-days-396}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-397}

}

\caption{ - empty caption - }\label{fig:example-days-397}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-398}

}

\caption{ - empty caption - }\label{fig:example-days-398}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-399}

}

\caption{ - empty caption - }\label{fig:example-days-399}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-400}

}

\caption{ - empty caption - }\label{fig:example-days-400}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-401}

}

\caption{ - empty caption - }\label{fig:example-days-401}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-402}

}

\caption{ - empty caption - }\label{fig:example-days-402}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-403}

}

\caption{ - empty caption - }\label{fig:example-days-403}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-404}

}

\caption{ - empty caption - }\label{fig:example-days-404}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-405}

}

\caption{ - empty caption - }\label{fig:example-days-405}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-406}

}

\caption{ - empty caption - }\label{fig:example-days-406}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-407}

}

\caption{ - empty caption - }\label{fig:example-days-407}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-408}

}

\caption{ - empty caption - }\label{fig:example-days-408}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-409}

}

\caption{ - empty caption - }\label{fig:example-days-409}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-410}

}

\caption{ - empty caption - }\label{fig:example-days-410}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-411}

}

\caption{ - empty caption - }\label{fig:example-days-411}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-412}

}

\caption{ - empty caption - }\label{fig:example-days-412}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-413}

}

\caption{ - empty caption - }\label{fig:example-days-413}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-414}

}

\caption{ - empty caption - }\label{fig:example-days-414}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-415}

}

\caption{ - empty caption - }\label{fig:example-days-415}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-416}

}

\caption{ - empty caption - }\label{fig:example-days-416}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-417}

}

\caption{ - empty caption - }\label{fig:example-days-417}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-418}

}

\caption{ - empty caption - }\label{fig:example-days-418}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-419}

}

\caption{ - empty caption - }\label{fig:example-days-419}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-420}

}

\caption{ - empty caption - }\label{fig:example-days-420}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-421}

}

\caption{ - empty caption - }\label{fig:example-days-421}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-422}

}

\caption{ - empty caption - }\label{fig:example-days-422}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-423}

}

\caption{ - empty caption - }\label{fig:example-days-423}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-424}

}

\caption{ - empty caption - }\label{fig:example-days-424}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-425}

}

\caption{ - empty caption - }\label{fig:example-days-425}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-426}

}

\caption{ - empty caption - }\label{fig:example-days-426}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-427}

}

\caption{ - empty caption - }\label{fig:example-days-427}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-428}

}

\caption{ - empty caption - }\label{fig:example-days-428}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-429}

}

\caption{ - empty caption - }\label{fig:example-days-429}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-430}

}

\caption{ - empty caption - }\label{fig:example-days-430}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-431}

}

\caption{ - empty caption - }\label{fig:example-days-431}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-432}

}

\caption{ - empty caption - }\label{fig:example-days-432}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-433}

}

\caption{ - empty caption - }\label{fig:example-days-433}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-434}

}

\caption{ - empty caption - }\label{fig:example-days-434}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-435}

}

\caption{ - empty caption - }\label{fig:example-days-435}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-436}

}

\caption{ - empty caption - }\label{fig:example-days-436}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-437}

}

\caption{ - empty caption - }\label{fig:example-days-437}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-438}

}

\caption{ - empty caption - }\label{fig:example-days-438}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-439}

}

\caption{ - empty caption - }\label{fig:example-days-439}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-440}

}

\caption{ - empty caption - }\label{fig:example-days-440}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-441}

}

\caption{ - empty caption - }\label{fig:example-days-441}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-442}

}

\caption{ - empty caption - }\label{fig:example-days-442}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-443}

}

\caption{ - empty caption - }\label{fig:example-days-443}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-444}

}

\caption{ - empty caption - }\label{fig:example-days-444}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-445}

}

\caption{ - empty caption - }\label{fig:example-days-445}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-446}

}

\caption{ - empty caption - }\label{fig:example-days-446}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-447}

}

\caption{ - empty caption - }\label{fig:example-days-447}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-448}

}

\caption{ - empty caption - }\label{fig:example-days-448}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-449}

}

\caption{ - empty caption - }\label{fig:example-days-449}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-450}

}

\caption{ - empty caption - }\label{fig:example-days-450}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-451}

}

\caption{ - empty caption - }\label{fig:example-days-451}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-452}

}

\caption{ - empty caption - }\label{fig:example-days-452}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-453}

}

\caption{ - empty caption - }\label{fig:example-days-453}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-454}

}

\caption{ - empty caption - }\label{fig:example-days-454}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-455}

}

\caption{ - empty caption - }\label{fig:example-days-455}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-456}

}

\caption{ - empty caption - }\label{fig:example-days-456}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-457}

}

\caption{ - empty caption - }\label{fig:example-days-457}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-458}

}

\caption{ - empty caption - }\label{fig:example-days-458}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-459}

}

\caption{ - empty caption - }\label{fig:example-days-459}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-460}

}

\caption{ - empty caption - }\label{fig:example-days-460}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-461}

}

\caption{ - empty caption - }\label{fig:example-days-461}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-462}

}

\caption{ - empty caption - }\label{fig:example-days-462}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-463}

}

\caption{ - empty caption - }\label{fig:example-days-463}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-464}

}

\caption{ - empty caption - }\label{fig:example-days-464}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-465}

}

\caption{ - empty caption - }\label{fig:example-days-465}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-466}

}

\caption{ - empty caption - }\label{fig:example-days-466}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-467}

}

\caption{ - empty caption - }\label{fig:example-days-467}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-468}

}

\caption{ - empty caption - }\label{fig:example-days-468}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-469}

}

\caption{ - empty caption - }\label{fig:example-days-469}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-470}

}

\caption{ - empty caption - }\label{fig:example-days-470}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-471}

}

\caption{ - empty caption - }\label{fig:example-days-471}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-472}

}

\caption{ - empty caption - }\label{fig:example-days-472}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-473}

}

\caption{ - empty caption - }\label{fig:example-days-473}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-474}

}

\caption{ - empty caption - }\label{fig:example-days-474}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-475}

}

\caption{ - empty caption - }\label{fig:example-days-475}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-476}

}

\caption{ - empty caption - }\label{fig:example-days-476}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-477}

}

\caption{ - empty caption - }\label{fig:example-days-477}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-478}

}

\caption{ - empty caption - }\label{fig:example-days-478}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-479}

}

\caption{ - empty caption - }\label{fig:example-days-479}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-480}

}

\caption{ - empty caption - }\label{fig:example-days-480}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-481}

}

\caption{ - empty caption - }\label{fig:example-days-481}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-482}

}

\caption{ - empty caption - }\label{fig:example-days-482}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-483}

}

\caption{ - empty caption - }\label{fig:example-days-483}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-484}

}

\caption{ - empty caption - }\label{fig:example-days-484}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-485}

}

\caption{ - empty caption - }\label{fig:example-days-485}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-486}

}

\caption{ - empty caption - }\label{fig:example-days-486}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-487}

}

\caption{ - empty caption - }\label{fig:example-days-487}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-488}

}

\caption{ - empty caption - }\label{fig:example-days-488}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-489}

}

\caption{ - empty caption - }\label{fig:example-days-489}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-490}

}

\caption{ - empty caption - }\label{fig:example-days-490}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-491}

}

\caption{ - empty caption - }\label{fig:example-days-491}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-492}

}

\caption{ - empty caption - }\label{fig:example-days-492}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-493}

}

\caption{ - empty caption - }\label{fig:example-days-493}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-494}

}

\caption{ - empty caption - }\label{fig:example-days-494}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-495}

}

\caption{ - empty caption - }\label{fig:example-days-495}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-496}

}

\caption{ - empty caption - }\label{fig:example-days-496}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-497}

}

\caption{ - empty caption - }\label{fig:example-days-497}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-498}

}

\caption{ - empty caption - }\label{fig:example-days-498}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-499}

}

\caption{ - empty caption - }\label{fig:example-days-499}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-500}

}

\caption{ - empty caption - }\label{fig:example-days-500}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-501}

}

\caption{ - empty caption - }\label{fig:example-days-501}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-502}

}

\caption{ - empty caption - }\label{fig:example-days-502}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-503}

}

\caption{ - empty caption - }\label{fig:example-days-503}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-504}

}

\caption{ - empty caption - }\label{fig:example-days-504}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-505}

}

\caption{ - empty caption - }\label{fig:example-days-505}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-506}

}

\caption{ - empty caption - }\label{fig:example-days-506}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-507}

}

\caption{ - empty caption - }\label{fig:example-days-507}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-508}

}

\caption{ - empty caption - }\label{fig:example-days-508}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-509}

}

\caption{ - empty caption - }\label{fig:example-days-509}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-510}

}

\caption{ - empty caption - }\label{fig:example-days-510}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-511}

}

\caption{ - empty caption - }\label{fig:example-days-511}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-512}

}

\caption{ - empty caption - }\label{fig:example-days-512}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-513}

}

\caption{ - empty caption - }\label{fig:example-days-513}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-514}

}

\caption{ - empty caption - }\label{fig:example-days-514}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-515}

}

\caption{ - empty caption - }\label{fig:example-days-515}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-516}

}

\caption{ - empty caption - }\label{fig:example-days-516}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-517}

}

\caption{ - empty caption - }\label{fig:example-days-517}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-518}

}

\caption{ - empty caption - }\label{fig:example-days-518}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-519}

}

\caption{ - empty caption - }\label{fig:example-days-519}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-520}

}

\caption{ - empty caption - }\label{fig:example-days-520}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-521}

}

\caption{ - empty caption - }\label{fig:example-days-521}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-522}

}

\caption{ - empty caption - }\label{fig:example-days-522}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-523}

}

\caption{ - empty caption - }\label{fig:example-days-523}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-524}

}

\caption{ - empty caption - }\label{fig:example-days-524}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-525}

}

\caption{ - empty caption - }\label{fig:example-days-525}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-526}

}

\caption{ - empty caption - }\label{fig:example-days-526}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-527}

}

\caption{ - empty caption - }\label{fig:example-days-527}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-528}

}

\caption{ - empty caption - }\label{fig:example-days-528}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-529}

}

\caption{ - empty caption - }\label{fig:example-days-529}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-530}

}

\caption{ - empty caption - }\label{fig:example-days-530}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-531}

}

\caption{ - empty caption - }\label{fig:example-days-531}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-532}

}

\caption{ - empty caption - }\label{fig:example-days-532}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-533}

}

\caption{ - empty caption - }\label{fig:example-days-533}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-534}

}

\caption{ - empty caption - }\label{fig:example-days-534}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-535}

}

\caption{ - empty caption - }\label{fig:example-days-535}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-536}

}

\caption{ - empty caption - }\label{fig:example-days-536}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-537}

}

\caption{ - empty caption - }\label{fig:example-days-537}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-538}

}

\caption{ - empty caption - }\label{fig:example-days-538}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-539}

}

\caption{ - empty caption - }\label{fig:example-days-539}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-540}

}

\caption{ - empty caption - }\label{fig:example-days-540}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-541}

}

\caption{ - empty caption - }\label{fig:example-days-541}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-542}

}

\caption{ - empty caption - }\label{fig:example-days-542}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-543}

}

\caption{ - empty caption - }\label{fig:example-days-543}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-544}

}

\caption{ - empty caption - }\label{fig:example-days-544}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-545}

}

\caption{ - empty caption - }\label{fig:example-days-545}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-546}

}

\caption{ - empty caption - }\label{fig:example-days-546}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-547}

}

\caption{ - empty caption - }\label{fig:example-days-547}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-548}

}

\caption{ - empty caption - }\label{fig:example-days-548}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-549}

}

\caption{ - empty caption - }\label{fig:example-days-549}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-550}

}

\caption{ - empty caption - }\label{fig:example-days-550}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-551}

}

\caption{ - empty caption - }\label{fig:example-days-551}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-552}

}

\caption{ - empty caption - }\label{fig:example-days-552}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-553}

}

\caption{ - empty caption - }\label{fig:example-days-553}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-554}

}

\caption{ - empty caption - }\label{fig:example-days-554}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-555}

}

\caption{ - empty caption - }\label{fig:example-days-555}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-556}

}

\caption{ - empty caption - }\label{fig:example-days-556}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-557}

}

\caption{ - empty caption - }\label{fig:example-days-557}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-558}

}

\caption{ - empty caption - }\label{fig:example-days-558}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-559}

}

\caption{ - empty caption - }\label{fig:example-days-559}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-560}

}

\caption{ - empty caption - }\label{fig:example-days-560}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-561}

}

\caption{ - empty caption - }\label{fig:example-days-561}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-562}

}

\caption{ - empty caption - }\label{fig:example-days-562}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-563}

}

\caption{ - empty caption - }\label{fig:example-days-563}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-564}

}

\caption{ - empty caption - }\label{fig:example-days-564}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-565}

}

\caption{ - empty caption - }\label{fig:example-days-565}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-566}

}

\caption{ - empty caption - }\label{fig:example-days-566}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-567}

}

\caption{ - empty caption - }\label{fig:example-days-567}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-568}

}

\caption{ - empty caption - }\label{fig:example-days-568}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-569}

}

\caption{ - empty caption - }\label{fig:example-days-569}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-570}

}

\caption{ - empty caption - }\label{fig:example-days-570}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-571}

}

\caption{ - empty caption - }\label{fig:example-days-571}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-572}

}

\caption{ - empty caption - }\label{fig:example-days-572}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-573}

}

\caption{ - empty caption - }\label{fig:example-days-573}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-574}

}

\caption{ - empty caption - }\label{fig:example-days-574}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-575}

}

\caption{ - empty caption - }\label{fig:example-days-575}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-576}

}

\caption{ - empty caption - }\label{fig:example-days-576}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-577}

}

\caption{ - empty caption - }\label{fig:example-days-577}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-578}

}

\caption{ - empty caption - }\label{fig:example-days-578}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-579}

}

\caption{ - empty caption - }\label{fig:example-days-579}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-580}

}

\caption{ - empty caption - }\label{fig:example-days-580}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-581}

}

\caption{ - empty caption - }\label{fig:example-days-581}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-582}

}

\caption{ - empty caption - }\label{fig:example-days-582}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-583}

}

\caption{ - empty caption - }\label{fig:example-days-583}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-584}

}

\caption{ - empty caption - }\label{fig:example-days-584}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-585}

}

\caption{ - empty caption - }\label{fig:example-days-585}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-586}

}

\caption{ - empty caption - }\label{fig:example-days-586}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-587}

}

\caption{ - empty caption - }\label{fig:example-days-587}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-588}

}

\caption{ - empty caption - }\label{fig:example-days-588}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-589}

}

\caption{ - empty caption - }\label{fig:example-days-589}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-590}

}

\caption{ - empty caption - }\label{fig:example-days-590}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-591}

}

\caption{ - empty caption - }\label{fig:example-days-591}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-592}

}

\caption{ - empty caption - }\label{fig:example-days-592}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-593}

}

\caption{ - empty caption - }\label{fig:example-days-593}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-594}

}

\caption{ - empty caption - }\label{fig:example-days-594}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-595}

}

\caption{ - empty caption - }\label{fig:example-days-595}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-596}

}

\caption{ - empty caption - }\label{fig:example-days-596}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-597}

}

\caption{ - empty caption - }\label{fig:example-days-597}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-598}

}

\caption{ - empty caption - }\label{fig:example-days-598}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-599}

}

\caption{ - empty caption - }\label{fig:example-days-599}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-600}

}

\caption{ - empty caption - }\label{fig:example-days-600}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-601}

}

\caption{ - empty caption - }\label{fig:example-days-601}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-602}

}

\caption{ - empty caption - }\label{fig:example-days-602}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-603}

}

\caption{ - empty caption - }\label{fig:example-days-603}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-604}

}

\caption{ - empty caption - }\label{fig:example-days-604}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-605}

}

\caption{ - empty caption - }\label{fig:example-days-605}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-606}

}

\caption{ - empty caption - }\label{fig:example-days-606}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-607}

}

\caption{ - empty caption - }\label{fig:example-days-607}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-608}

}

\caption{ - empty caption - }\label{fig:example-days-608}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-609}

}

\caption{ - empty caption - }\label{fig:example-days-609}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-610}

}

\caption{ - empty caption - }\label{fig:example-days-610}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-611}

}

\caption{ - empty caption - }\label{fig:example-days-611}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-612}

}

\caption{ - empty caption - }\label{fig:example-days-612}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-613}

}

\caption{ - empty caption - }\label{fig:example-days-613}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-614}

}

\caption{ - empty caption - }\label{fig:example-days-614}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-615}

}

\caption{ - empty caption - }\label{fig:example-days-615}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-616}

}

\caption{ - empty caption - }\label{fig:example-days-616}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-617}

}

\caption{ - empty caption - }\label{fig:example-days-617}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-618}

}

\caption{ - empty caption - }\label{fig:example-days-618}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-619}

}

\caption{ - empty caption - }\label{fig:example-days-619}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-620}

}

\caption{ - empty caption - }\label{fig:example-days-620}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-621}

}

\caption{ - empty caption - }\label{fig:example-days-621}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-622}

}

\caption{ - empty caption - }\label{fig:example-days-622}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-623}

}

\caption{ - empty caption - }\label{fig:example-days-623}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-624}

}

\caption{ - empty caption - }\label{fig:example-days-624}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-625}

}

\caption{ - empty caption - }\label{fig:example-days-625}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-626}

}

\caption{ - empty caption - }\label{fig:example-days-626}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-627}

}

\caption{ - empty caption - }\label{fig:example-days-627}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-628}

}

\caption{ - empty caption - }\label{fig:example-days-628}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-629}

}

\caption{ - empty caption - }\label{fig:example-days-629}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-630}

}

\caption{ - empty caption - }\label{fig:example-days-630}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-631}

}

\caption{ - empty caption - }\label{fig:example-days-631}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-632}

}

\caption{ - empty caption - }\label{fig:example-days-632}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-633}

}

\caption{ - empty caption - }\label{fig:example-days-633}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-634}

}

\caption{ - empty caption - }\label{fig:example-days-634}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-635}

}

\caption{ - empty caption - }\label{fig:example-days-635}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-636}

}

\caption{ - empty caption - }\label{fig:example-days-636}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-637}

}

\caption{ - empty caption - }\label{fig:example-days-637}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-638}

}

\caption{ - empty caption - }\label{fig:example-days-638}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-639}

}

\caption{ - empty caption - }\label{fig:example-days-639}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-640}

}

\caption{ - empty caption - }\label{fig:example-days-640}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-641}

}

\caption{ - empty caption - }\label{fig:example-days-641}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-642}

}

\caption{ - empty caption - }\label{fig:example-days-642}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-643}

}

\caption{ - empty caption - }\label{fig:example-days-643}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-644}

}

\caption{ - empty caption - }\label{fig:example-days-644}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-645}

}

\caption{ - empty caption - }\label{fig:example-days-645}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-646}

}

\caption{ - empty caption - }\label{fig:example-days-646}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-647}

}

\caption{ - empty caption - }\label{fig:example-days-647}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-648}

}

\caption{ - empty caption - }\label{fig:example-days-648}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-649}

}

\caption{ - empty caption - }\label{fig:example-days-649}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-650}

}

\caption{ - empty caption - }\label{fig:example-days-650}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-651}

}

\caption{ - empty caption - }\label{fig:example-days-651}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-652}

}

\caption{ - empty caption - }\label{fig:example-days-652}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-653}

}

\caption{ - empty caption - }\label{fig:example-days-653}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-654}

}

\caption{ - empty caption - }\label{fig:example-days-654}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-655}

}

\caption{ - empty caption - }\label{fig:example-days-655}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-656}

}

\caption{ - empty caption - }\label{fig:example-days-656}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-657}

}

\caption{ - empty caption - }\label{fig:example-days-657}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-658}

}

\caption{ - empty caption - }\label{fig:example-days-658}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-659}

}

\caption{ - empty caption - }\label{fig:example-days-659}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-660}

}

\caption{ - empty caption - }\label{fig:example-days-660}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-661}

}

\caption{ - empty caption - }\label{fig:example-days-661}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-662}

}

\caption{ - empty caption - }\label{fig:example-days-662}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-663}

}

\caption{ - empty caption - }\label{fig:example-days-663}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-664}

}

\caption{ - empty caption - }\label{fig:example-days-664}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-665}

}

\caption{ - empty caption - }\label{fig:example-days-665}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-666}

}

\caption{ - empty caption - }\label{fig:example-days-666}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-667}

}

\caption{ - empty caption - }\label{fig:example-days-667}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-668}

}

\caption{ - empty caption - }\label{fig:example-days-668}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-669}

}

\caption{ - empty caption - }\label{fig:example-days-669}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-670}

}

\caption{ - empty caption - }\label{fig:example-days-670}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-671}

}

\caption{ - empty caption - }\label{fig:example-days-671}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-672}

}

\caption{ - empty caption - }\label{fig:example-days-672}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-673}

}

\caption{ - empty caption - }\label{fig:example-days-673}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-674}

}

\caption{ - empty caption - }\label{fig:example-days-674}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-675}

}

\caption{ - empty caption - }\label{fig:example-days-675}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-676}

}

\caption{ - empty caption - }\label{fig:example-days-676}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-677}

}

\caption{ - empty caption - }\label{fig:example-days-677}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-678}

}

\caption{ - empty caption - }\label{fig:example-days-678}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-679}

}

\caption{ - empty caption - }\label{fig:example-days-679}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-680}

}

\caption{ - empty caption - }\label{fig:example-days-680}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-681}

}

\caption{ - empty caption - }\label{fig:example-days-681}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-682}

}

\caption{ - empty caption - }\label{fig:example-days-682}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-683}

}

\caption{ - empty caption - }\label{fig:example-days-683}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-684}

}

\caption{ - empty caption - }\label{fig:example-days-684}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-685}

}

\caption{ - empty caption - }\label{fig:example-days-685}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-686}

}

\caption{ - empty caption - }\label{fig:example-days-686}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-687}

}

\caption{ - empty caption - }\label{fig:example-days-687}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-688}

}

\caption{ - empty caption - }\label{fig:example-days-688}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-689}

}

\caption{ - empty caption - }\label{fig:example-days-689}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-690}

}

\caption{ - empty caption - }\label{fig:example-days-690}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-691}

}

\caption{ - empty caption - }\label{fig:example-days-691}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-692}

}

\caption{ - empty caption - }\label{fig:example-days-692}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-693}

}

\caption{ - empty caption - }\label{fig:example-days-693}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-694}

}

\caption{ - empty caption - }\label{fig:example-days-694}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-695}

}

\caption{ - empty caption - }\label{fig:example-days-695}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-696}

}

\caption{ - empty caption - }\label{fig:example-days-696}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-697}

}

\caption{ - empty caption - }\label{fig:example-days-697}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-698}

}

\caption{ - empty caption - }\label{fig:example-days-698}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-699}

}

\caption{ - empty caption - }\label{fig:example-days-699}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-700}

}

\caption{ - empty caption - }\label{fig:example-days-700}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-701}

}

\caption{ - empty caption - }\label{fig:example-days-701}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-702}

}

\caption{ - empty caption - }\label{fig:example-days-702}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-703}

}

\caption{ - empty caption - }\label{fig:example-days-703}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-704}

}

\caption{ - empty caption - }\label{fig:example-days-704}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-705}

}

\caption{ - empty caption - }\label{fig:example-days-705}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-706}

}

\caption{ - empty caption - }\label{fig:example-days-706}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-707}

}

\caption{ - empty caption - }\label{fig:example-days-707}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-708}

}

\caption{ - empty caption - }\label{fig:example-days-708}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-709}

}

\caption{ - empty caption - }\label{fig:example-days-709}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-710}

}

\caption{ - empty caption - }\label{fig:example-days-710}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-711}

}

\caption{ - empty caption - }\label{fig:example-days-711}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-712}

}

\caption{ - empty caption - }\label{fig:example-days-712}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-713}

}

\caption{ - empty caption - }\label{fig:example-days-713}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-714}

}

\caption{ - empty caption - }\label{fig:example-days-714}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-715}

}

\caption{ - empty caption - }\label{fig:example-days-715}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-716}

}

\caption{ - empty caption - }\label{fig:example-days-716}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-717}

}

\caption{ - empty caption - }\label{fig:example-days-717}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-718}

}

\caption{ - empty caption - }\label{fig:example-days-718}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-719}

}

\caption{ - empty caption - }\label{fig:example-days-719}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-720}

}

\caption{ - empty caption - }\label{fig:example-days-720}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-721}

}

\caption{ - empty caption - }\label{fig:example-days-721}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-722}

}

\caption{ - empty caption - }\label{fig:example-days-722}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-723}

}

\caption{ - empty caption - }\label{fig:example-days-723}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-724}

}

\caption{ - empty caption - }\label{fig:example-days-724}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-725}

}

\caption{ - empty caption - }\label{fig:example-days-725}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-726}

}

\caption{ - empty caption - }\label{fig:example-days-726}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-727}

}

\caption{ - empty caption - }\label{fig:example-days-727}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-728}

}

\caption{ - empty caption - }\label{fig:example-days-728}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-729}

}

\caption{ - empty caption - }\label{fig:example-days-729}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-730}

}

\caption{ - empty caption - }\label{fig:example-days-730}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-731}

}

\caption{ - empty caption - }\label{fig:example-days-731}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-732}

}

\caption{ - empty caption - }\label{fig:example-days-732}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-733}

}

\caption{ - empty caption - }\label{fig:example-days-733}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-734}

}

\caption{ - empty caption - }\label{fig:example-days-734}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-735}

}

\caption{ - empty caption - }\label{fig:example-days-735}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-736}

}

\caption{ - empty caption - }\label{fig:example-days-736}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-737}

}

\caption{ - empty caption - }\label{fig:example-days-737}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-738}

}

\caption{ - empty caption - }\label{fig:example-days-738}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-739}

}

\caption{ - empty caption - }\label{fig:example-days-739}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-740}

}

\caption{ - empty caption - }\label{fig:example-days-740}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-741}

}

\caption{ - empty caption - }\label{fig:example-days-741}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-742}

}

\caption{ - empty caption - }\label{fig:example-days-742}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-743}

}

\caption{ - empty caption - }\label{fig:example-days-743}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-744}

}

\caption{ - empty caption - }\label{fig:example-days-744}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-745}

}

\caption{ - empty caption - }\label{fig:example-days-745}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-746}

}

\caption{ - empty caption - }\label{fig:example-days-746}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-747}

}

\caption{ - empty caption - }\label{fig:example-days-747}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-748}

}

\caption{ - empty caption - }\label{fig:example-days-748}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-749}

}

\caption{ - empty caption - }\label{fig:example-days-749}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-750}

}

\caption{ - empty caption - }\label{fig:example-days-750}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-751}

}

\caption{ - empty caption - }\label{fig:example-days-751}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-752}

}

\caption{ - empty caption - }\label{fig:example-days-752}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-753}

}

\caption{ - empty caption - }\label{fig:example-days-753}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-754}

}

\caption{ - empty caption - }\label{fig:example-days-754}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-755}

}

\caption{ - empty caption - }\label{fig:example-days-755}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-756}

}

\caption{ - empty caption - }\label{fig:example-days-756}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-757}

}

\caption{ - empty caption - }\label{fig:example-days-757}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-758}

}

\caption{ - empty caption - }\label{fig:example-days-758}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-759}

}

\caption{ - empty caption - }\label{fig:example-days-759}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-760}

}

\caption{ - empty caption - }\label{fig:example-days-760}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-761}

}

\caption{ - empty caption - }\label{fig:example-days-761}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-762}

}

\caption{ - empty caption - }\label{fig:example-days-762}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-763}

}

\caption{ - empty caption - }\label{fig:example-days-763}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-764}

}

\caption{ - empty caption - }\label{fig:example-days-764}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-765}

}

\caption{ - empty caption - }\label{fig:example-days-765}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-766}

}

\caption{ - empty caption - }\label{fig:example-days-766}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-767}

}

\caption{ - empty caption - }\label{fig:example-days-767}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-768}

}

\caption{ - empty caption - }\label{fig:example-days-768}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-769}

}

\caption{ - empty caption - }\label{fig:example-days-769}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-770}

}

\caption{ - empty caption - }\label{fig:example-days-770}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-771}

}

\caption{ - empty caption - }\label{fig:example-days-771}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-772}

}

\caption{ - empty caption - }\label{fig:example-days-772}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-773}

}

\caption{ - empty caption - }\label{fig:example-days-773}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-774}

}

\caption{ - empty caption - }\label{fig:example-days-774}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-775}

}

\caption{ - empty caption - }\label{fig:example-days-775}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-776}

}

\caption{ - empty caption - }\label{fig:example-days-776}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-777}

}

\caption{ - empty caption - }\label{fig:example-days-777}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-778}

}

\caption{ - empty caption - }\label{fig:example-days-778}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-779}

}

\caption{ - empty caption - }\label{fig:example-days-779}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-780}

}

\caption{ - empty caption - }\label{fig:example-days-780}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-781}

}

\caption{ - empty caption - }\label{fig:example-days-781}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-782}

}

\caption{ - empty caption - }\label{fig:example-days-782}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-783}

}

\caption{ - empty caption - }\label{fig:example-days-783}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-784}

}

\caption{ - empty caption - }\label{fig:example-days-784}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-785}

}

\caption{ - empty caption - }\label{fig:example-days-785}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-786}

}

\caption{ - empty caption - }\label{fig:example-days-786}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-787}

}

\caption{ - empty caption - }\label{fig:example-days-787}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-788}

}

\caption{ - empty caption - }\label{fig:example-days-788}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-789}

}

\caption{ - empty caption - }\label{fig:example-days-789}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-790}

}

\caption{ - empty caption - }\label{fig:example-days-790}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-791}

}

\caption{ - empty caption - }\label{fig:example-days-791}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-792}

}

\caption{ - empty caption - }\label{fig:example-days-792}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-793}

}

\caption{ - empty caption - }\label{fig:example-days-793}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-794}

}

\caption{ - empty caption - }\label{fig:example-days-794}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-795}

}

\caption{ - empty caption - }\label{fig:example-days-795}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-796}

}

\caption{ - empty caption - }\label{fig:example-days-796}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-797}

}

\caption{ - empty caption - }\label{fig:example-days-797}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-798}

}

\caption{ - empty caption - }\label{fig:example-days-798}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-799}

}

\caption{ - empty caption - }\label{fig:example-days-799}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-800}

}

\caption{ - empty caption - }\label{fig:example-days-800}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-801}

}

\caption{ - empty caption - }\label{fig:example-days-801}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-802}

}

\caption{ - empty caption - }\label{fig:example-days-802}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-803}

}

\caption{ - empty caption - }\label{fig:example-days-803}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-804}

}

\caption{ - empty caption - }\label{fig:example-days-804}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-805}

}

\caption{ - empty caption - }\label{fig:example-days-805}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-806}

}

\caption{ - empty caption - }\label{fig:example-days-806}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-807}

}

\caption{ - empty caption - }\label{fig:example-days-807}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-808}

}

\caption{ - empty caption - }\label{fig:example-days-808}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-809}

}

\caption{ - empty caption - }\label{fig:example-days-809}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-810}

}

\caption{ - empty caption - }\label{fig:example-days-810}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-811}

}

\caption{ - empty caption - }\label{fig:example-days-811}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-812}

}

\caption{ - empty caption - }\label{fig:example-days-812}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-813}

}

\caption{ - empty caption - }\label{fig:example-days-813}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-814}

}

\caption{ - empty caption - }\label{fig:example-days-814}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-815}

}

\caption{ - empty caption - }\label{fig:example-days-815}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-816}

}

\caption{ - empty caption - }\label{fig:example-days-816}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-817}

}

\caption{ - empty caption - }\label{fig:example-days-817}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-818}

}

\caption{ - empty caption - }\label{fig:example-days-818}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-819}

}

\caption{ - empty caption - }\label{fig:example-days-819}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-820}

}

\caption{ - empty caption - }\label{fig:example-days-820}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-821}

}

\caption{ - empty caption - }\label{fig:example-days-821}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-822}

}

\caption{ - empty caption - }\label{fig:example-days-822}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-823}

}

\caption{ - empty caption - }\label{fig:example-days-823}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-824}

}

\caption{ - empty caption - }\label{fig:example-days-824}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-825}

}

\caption{ - empty caption - }\label{fig:example-days-825}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-826}

}

\caption{ - empty caption - }\label{fig:example-days-826}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-827}

}

\caption{ - empty caption - }\label{fig:example-days-827}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-828}

}

\caption{ - empty caption - }\label{fig:example-days-828}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-829}

}

\caption{ - empty caption - }\label{fig:example-days-829}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-830}

}

\caption{ - empty caption - }\label{fig:example-days-830}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-831}

}

\caption{ - empty caption - }\label{fig:example-days-831}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-832}

}

\caption{ - empty caption - }\label{fig:example-days-832}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-833}

}

\caption{ - empty caption - }\label{fig:example-days-833}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-834}

}

\caption{ - empty caption - }\label{fig:example-days-834}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-835}

}

\caption{ - empty caption - }\label{fig:example-days-835}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_daily_test_plots_files/figure-latex/example-days-836}

}

\caption{ - empty caption - }\label{fig:example-days-836}
\end{figure}

\textbf{END}

\begin{verbatim}
2024-09-06 06:08:04.9 athan@sagan GHI_enh_02_ID_CE.R 2.698150 mins
\end{verbatim}

\end{document}