### 2024-09-06 GHI_enh_08_validation.tex
Commit 2024-09-06 14:14

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Enhancement of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{float}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Enhancement of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2024-09-06}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI enchantment.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
Drop all enhacement cases
\end{verbatim}

\begin{verbatim}
Drop all clouds
\end{verbatim}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/unnamed-chunk-4-1}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-4}
\end{figure}

\FloatBarrier

\hypertarget{plot-some-days-with-strong-enhancement-cases}{%
\section{Plot some days with strong enhancement cases}\label{plot-some-days-with-strong-enhancement-cases}}

\FloatBarrier

\hypertarget{days-with-sunny-day}{%
\subsection{Days with sunny day}\label{days-with-sunny-day}}

\FloatBarrier

\hypertarget{days-with-sunny-enhancement-day}{%
\subsection{Days with sunny enhancement day}\label{days-with-sunny-enhancement-day}}

\FloatBarrier

\hypertarget{days-with-cloudy-day}{%
\subsection{Days with cloudy day}\label{days-with-cloudy-day}}

\FloatBarrier

\hypertarget{days-with-random-day}{%
\subsection{Days with random day}\label{days-with-random-day}}

\FloatBarrier

\hypertarget{days-with-manual-test-days}{%
\subsection{Days with manual test days}\label{days-with-manual-test-days}}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/example-days-1}

}

\caption{ - empty caption - }\label{fig:example-days-1}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/example-days-2}

}

\caption{ - empty caption - }\label{fig:example-days-2}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/example-days-3}

}

\caption{ - empty caption - }\label{fig:example-days-3}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/example-days-4}

}

\caption{ - empty caption - }\label{fig:example-days-4}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/example-days-5}

}

\caption{ - empty caption - }\label{fig:example-days-5}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/unnamed-chunk-6-1}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-6}
\end{figure}

\begin{verbatim}
    Pearson's product-moment correlation

data:  KEEP$wattGLB and KEEP$Low_B.Low_W.glo
t = 7903.5, df = 345175, p-value < 2.2e-16
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
 0.9972301 0.9972667
sample estimates:
      cor 
0.9972485 
\end{verbatim}

\begin{verbatim}
Call:
lm(formula = KEEP[, wattGLB, Low_B.Low_W.glo])

Coefficients:
(Intercept)      wattGLB  
     18.357        1.038  
\end{verbatim}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{GHI_enh_08_validation_files/figure-latex/unnamed-chunk-7-1}

}

\caption{ - empty caption - }\label{fig:unnamed-chunk-7}
\end{figure}

\begin{verbatim}
    Pearson's product-moment correlation

data:  KEEP$wattGLB and KEEP$Enhanc_C_4_ref
t = 7938, df = 345175, p-value < 2.2e-16
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
 0.9972540 0.9972904
sample estimates:
      cor 
0.9972722 
\end{verbatim}

\begin{verbatim}
Call:
lm(formula = KEEP[, wattGLB, Enhanc_C_4_ref])

Coefficients:
(Intercept)      wattGLB  
     43.323        1.036  
\end{verbatim}

\textbf{END}

\begin{verbatim}
2024-09-06 10:32:08.3 athan@sagan GHI_enh_08_validation.R 1.913171 mins
\end{verbatim}

\end{document}